### 2016-03-28 derivations.tex
first commit

\subsection{Preservation of Angles}

For this term the energy depends on the difference of the angles between adjacent edges.

Energy\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
\begin{eqnarray}
 E_\varphi(\varphi) & = & \frac{1}{2}k_\varphi(\varphi - \varphi_0) \textmd{   *} \\
 \varphi(p_i,p_j,p_k) & = & \arccos \left( \dfrac{(p_j - p_i) \cdot (p_k - p_i)}{ \left\| p_j - p_i \right\| \left\|p_k - p_i \right\|} \right)^2 \nonumber
\end{eqnarray}

Where $\varphi$ is the angle between adjacent edges, $E_\varphi$ is the energy associated to changes in the angle, $k_\varphi$ is the corresponding stiffness constant and the $p_i$s are the Cartesian coordinates of the mass particles.  Contrary to the previous cases, it is not so evident in which direction the force will act.  Given that the force acts in the direction of the gradient, it will tend to restore the angles in the most efficient way, but it may produce tiny or very big triangles if it is not accompanied by some of the other terms proposed by Teschner, that tend to restore the original dimensions, apart from the angles.

% \begin{figure}[htp]
%  \centering
% \includegraphics{./Figures/springs/anglepres.png}
%  \caption{The angular force displaces the vertex of interest in the direction of maximum change of the angle in order to recover its original value.  This direction is a linear combination of the vectors that emerge from both edges forming the angle, and does not respect the original size of the triangle.}\label{fig:angular}
% \end{figure}

The force emerging from this term is a linear combination of the vectors along the edges that form the angle of interest, it pretends to restore the original angle, but does not take the original size into account. See \fref{fig:forces}(c). Therefore, it helps to recover a similar triangle, but if used alone can collapse or explode the triangle.  The derivation of this force is as follows:  %An additional line in the code also forces an inverted angle to recover its original orientation.

\begin{align}
 F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
 \dfrac{\partial \varphi}{\partial p_i} & = \dfrac{\partial}{\partial p_i} arccos \left(\frac{(\vec{p}_j-\vec{p}_i)\cdot(\vec{p}_k-\vec{p}_i)}{\left\|\vec{p}_j-\vec{p}_i\right\|\left\|\vec{p}_k-\vec{p}_i\right\|}\right) \displaybreak[0] \\
\end{align}

Summarising, the force is:
\begin{align}
 F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
 \dfrac{\partial \varphi}{\partial p_i}(p_i) & =
    \dfrac{1}{d_{ji}d_{ki}\sqrt{1-\left[\dfrac{pp}{(d_{ji}) (d_{ki})}\right]^2}} \nonumber \\
 & \left\{\left[1-\dfrac{pp}{d_{ki}^2}\right](p_k-p_i)+\left[1-\dfrac{pp}{d_{ji}^2}\right](p_j-p_i) \right\} \nonumber \\
 pp(p_i) & =(p_j-p_i)\cdot(p_k-p_i) \nonumber \\
 d_{ji}(p_i) & = \left\| p_j-p_i \right\| \nonumber \\
 d_{ki}(p_i) & = \left\| p_k-p_i \right\|
\end{align}

The terms for the preservation of length and angle model elastic deformations.  Even though the preservation of area by itself would allow some plasticity, it is still necessary to incorporate permanent deformations in the rest lengths of the springs, to stop the triangles of the mesh from trying to recover their original shape.

### 2016-03-28 twostagedo.tex
first commit

%% bare_jrnl.tex
%% V1.4
%% 2012/12/27
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex,
%%                    bare_jrnl_transmag.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}

\newcommand{\eref}[1]{(\ref{#1})}
\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\tref}[1]{Table~\ref{#1}}
\newcommand{\sref}[1]{Section~\ref{#1}}
\newcommand{\aref}[1]{Appendix~\ref{#1}}
\newcommand{\alref}[1]{Algorithm~\ref{#1}}
\newcommand{\unit}[1]{\ensuremath{\, \mathrm{#1}}}
\usepackage{tikz}

\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{xcolor}
% \definecolor{DarkGreen}{rgb}{0,0.5,0}
% \def\gtick{\textcolor{DarkGreen}{\checkmark}}
\def\gtick{\checkmark}
% \definecolor{DarkRed}{rgb}{0.5,0,0}
% \def\rcross{\textcolor{DarkRed}{x}}
\def\rcross{x}
\usepackage[pdftex]{rotating}

%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  \graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
  \DeclareGraphicsExtensions{.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
\usepackage{framed} % suggested in http://tex.stackexchange.com/questions/65219/how-to-make-a-code-fragment-in-ieeetran-classes
% My own definition:
\newcounter{algorithm}
\newenvironment{algorithmieee}[1]
{\refstepcounter{algorithm}{1} \begin{figure} \textbf{Algorithm \arabic{algorithm}:} \textit{#1} %
\begin{algorithmic}[1]}
{\end{algorithmic}
 \end{figure}}

% Adapted from http://tex.stackexchange.com/questions/40228/ieeetrans-template-requires-not-to-use-algorithm-package
\newsavebox{\ieeealgbox}
\newenvironment{algorithmicieee}[1]
  {\refstepcounter{algorithm}
   \begin{lrbox}{\ieeealgbox}
   \begin{minipage}{\dimexpr\columnwidth-2\fboxsep-2\fboxrule}
   \textbf{Algorithm \arabic{algorithm}:} \textit{#1} \par
   \begin{algorithmic}[1]}
  {\end{algorithmic}
   \end{minipage}
   \end{lrbox}\noindent\fbox{\usebox{\ieeealgbox}}}

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% ++++ Used temporarily to make comments between authors
\newcommand{\redNote}[1]{{\color{red} #1}}
\usepackage[normalem]{ulem}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Veronica E.~Arriola-Rios,
        Jeremy~L.~Wyatt}
%\author{Veronica E.~Arriola-Rios,~\IEEEmembership{Member,~IEEE,}
%        John~Doe,~\IEEEmembership{Fellow,~OSA,}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell is with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised December 27, 2012.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Systems, Man, and Cybernetics—Part B: Cybernetics,
          ~Vol.~11, No.~4, December~2012}%
{Arriola-Rios \MakeLowercase{\textit{et al.}}: A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2012 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper we present a new framework for learning generative models of object deformation under robotic action. The model is multi-modal in that it is based on integrating force and visual information. Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning model framework, and none has offered classification from a generative predictable model.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Deformable objects, prediction, classification, learning.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robotic manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, and these include objects that are elastic, plastic, breakable, tearable, and spreadable. This work is based on the premise that in order to plan manipulations of deformable objects, predictive models of object deformation are required. Without such models planners cannot reason about the effects of actions, or sequences of actions to achieve desired effects.

There is a wide range of models of deformation used in engineering, that are capable of producing predictions, such as finite element models, but these typically require careful optimisation by hand over many weeks or months. In graphics there are many models used for generating qualitatively plausible simulations of objects, but these are not necessarily easy to tune to real data. By contrast robots must perform system identification of their model parameters from data, on-line, in reasonable amounts of training time. This includes building models from the usually very limited sensing typically available, such as partial visual views of an object, and force-torque monitoring at contact points. Thus tackling this problem in robotics adds additional difficulties to an already challenging problem.

\begin{figure}[!t]
\centering\includegraphics[width=3.5in]{arrio17.png}
\caption{Force information acquired while working with a block of sponge (elastic) and a piece of plasticine (plastic).}
\label{fig:forceElasticPlastic}
\end{figure}

Previous work on prediction for deformable objects in robotics has been suprisingly limited. Table~\ref{t:advances} summarises the features of the major efforts to date. In each case critical abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g. elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?). In this paper we present a framework for solving all these problems simultaneously. The key novel contribution we make is to bring together a model of what forces the robot will feel at a contact point with a model of how an object will deform under such applied forces over time. This formulation allows the robot to make predictions about the type of object, and thus it's future behaviour using the resistive forces and the object deformations \fref{fig:forceElasticPlastic} and \fref{fig:sstrain}. A second critical contribution is that our model is both generative and generalisable with respect to contact point.  Finally the third major feature of the model is that it is adaptive: a set of model parameters is learned for each material presented from data, thus creating a model library of materials. This generative property enables both prediction suitable for use in planning, and also classification of materials. The overall model structure is summarised in \fref{fig:system}. Essentially the approach is to use visual tracking of shape deformations to create feedback that is used to train the parameters of a model (Model A) that predicts deformation when coupled with contact forces. The contact forces are themselves predicted using a second model (Model B). In the specific implementation of our system we use a mass spring system for Model A, and a regression model for Model B. These models are learned from the same data, and are both indexable by a known material type from the library. It is important to note, however, that the model structure is separate from the specific model forms used, and that other model types could be substituted for one or both stages.  This behaviour is described in \alref{alg:overall}.

\begin{figure}
 \begin{algorithmicieee}{Overall Behaviour}\label{alg:overall}
 {\small
 \STATE \textbf{Training Phase:}
 \FORALL {training materials:}
   \FORALL {finger commands (push down, push up)}
     \STATE The controler gives the command.
     \WHILE {the command is executed}
       \STATE The F/T sensor module records the reaction force.
       \FORALL {candidate regresion curves}
         \STATE The F/T sensor module adjusts the curve.
       \ENDFOR
       \STATE The visual system tracks and records the modifications of the shape from a 2D camera.
     \ENDWHILE
     \STATE The F/T sensor module records a discontinuity in the regresion curve.
   \ENDFOR
   \STATE The F/T sensor module returns the best regresion curve as Model B.
   \STATE The learning algorith for the mass-spring model proposes candidate parameters.
   \FORALL{frames of captured information}
     \FORALL{sets of candidate parameters}
       \STATE Evaluate how similar the mass-spring simulation is to the real object.
     \ENDFOR
   \ENDFOR
   \STATE The values of the best set of parameters is adopted as Model A.
 \ENDFOR
 \STATE \textbf{Simulation Phase:}
 \WHILE{the finger penetrates the material}
   \STATE Use Model B to predict the reaction force given the distance traversed after contact with the material.
   \STATE Use this force to run the simulation with Model A.
   \STATE Model A predicts the shape the material should have.
 \ENDWHILE
 \STATE \textbf{Classification Phase:}
 \FORALL{frames of captured information}
  \FORALL{previously trained materials}
    \STATE Run simulation and evaluate similitude
  \ENDFOR
 \ENDFOR
 \STATE Return material with the highest score.}
 \end{algorithmicieee}
\caption{Overall algorithmic structure of the system.}
\end{figure}

In summary the two stage model we present gives us the ability to solve {\em simultaneously} for the first time the following problems. Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Model Overview}

As has been mentioned above there are several choices made in our modelling approach. The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour. The specific approach taken here separates the predictive model out into components in two ways (see Figure~\ref{fig:system}). First we decompose models according to the modality of the information. In this case it is important to note that deformation has a particular causal structure related to the modes of sensing that are most easily employed. The forces and deformations can be measured separately, using force-torque and visual sensors respectively. The sensed forces are the cause of the observed deformation, and this can be exploited by sequencing the models by mode. Thus the input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces. The second predictor takes this sequence of forces as input and predicts a deformation sequence. Both models must be conditioned on the material and the object. This is where we decompose the model space in the second way, by specialising models to particular material and object combinations. This follows the principle of modular motor learning \cite{KopickiICRA11,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type. Thus we aim to acquire many specialist models of different deformable objects rather than a very few rather general models. The specific model forms we choose here could be replaced while retaining the overall scheme. In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.

In each case the models can be run in two modes, prediction and classification. In prediction mode the models make predictions based on a sequence of robot finger positions. To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive. This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging. In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models. Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.

The training of the model is performed offline using ground truthed data. The overall algorithm is given in \alref{alg:overall}. For clarity the algorithm is given in three phases: training, simulation (i.e. the prediction phase), and classification (i.e. matching prediction to new data). In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape. Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).

\section{Background}
\label{sec:background}
This work focuses on prediction of response forces and of the deformation of the shape over several frames of elastic and plastic objects when they are pushed and released by a robotic finger.  Acquired models are used to classify materials through recognition of trained samples, even when the interactions are novel.  Even though there is abundant literature in the industrial robotics area for 1D and 2D materials (e.g. strings, hair, metallic sheets and textiles), there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.

First, Howard and Bekey \cite{Howard2000} address robotic grasping and lifting of viscoelastic objects.  Their model of the material can predict a net amount of deformation given the applied forces, but does not evaluate details of the deformed shape.   This model is inspired in the crystalline structure of atoms in solids.  The space lattice of the crystal is approximated by a particle system where elements with mass are connected by springs and dampers, thus representing the attraction and repulsion between atoms \emph{in equilibrium} with a damped mass-spring system.  The masses are calculated by measuring the force required to lift the object without sliding.  The deformation and damping coefficients are estimated as functions of:
\begin{itemize}
 \item the force applied while pushing against the material with two robotic arms in opposite directions and,
 \item the global displacement and velocity vectors resulting from adding the displacements and velocities of all particles.
\end{itemize}
The estimated values of the mass, deformability and damping coefficients are used as the entry of a neural network that outputs the minimum necessary force to lift the object without sliding.  This function is used to compliantly adjust the force during an interaction.  However, they do not evaluate the quality of the overall shape prediction, and use only the final values of the prediction after the system has stabilised; that is, they do not evaluate the predictions during the dynamic phase.

\begin{figure}[!t]
\centering
\setlength\fboxsep{0pt}    % padding
\setlength\fboxrule{0.5pt} % line width
\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
\label{fig:system}
\end{figure}

%%%
%%% TODO: Text
%%%
% \begin{figure}[!t]
% \centering
% \setlength\fboxsep{0pt}    % padding
% \setlength\fboxrule{0.5pt} % line width
% \fbox{\includegraphics{arrio3}}
% \caption{The system learns to predict the object's shape and reaction forces from real world data.}
% \label{fig:system_old}
% \end{figure}

Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, whose constraints are expressed as potential energy terms.  These terms encode the tendency of springs to recover their length (with damping), of triangular faces of a mesh to recover their area and of tetrahedrons to recover their volume.  He sketches the inclusion of plastic deformation.  Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a FEM model.  The  search begins with a uniform random sampling of possible values for the elasticity constants, which are latter modified using adaptive simulated annealing to minimize the difference between the deformations in the FEM mesh and those in the mass-spring mesh.  However, they only compare the steady states and do not include plastic deformations.  We extend on this work, by changing the FEM model with force and 2D visual data gathered from real materials.  Furthermore, we analysed the quality of the predictions while the deformations take place.

Frank et al.\ \cite{Frank2010} use a force sensor and a bumblebee stereo camera to measure the Young modulus and the Poisson ratio of unknown deformable objects.  By assuming a homogeneous material, they propose candidate values for these constitutive parameters and simulate the expected behaviour of the object with a FEM method.  The volume of the prediction is compared with a 3D reconstruction of the 3D surface of the object.  A gradient descent search is used to minimize the difference.  Once the parameters are selected, a collision detection algorithm informs the model of the position of the actuator thus implying the amount of local deformation.  The FEM simulation provides estimates for the global deformation and response forces.  However, a quasi-static assumption is made and a detailed study of the dynamic process is not offered\footnote{With the quasi-static assumption a small force is applied and a state of equilibrium is reached before a new force is applied.  Therefore the deformation process is approximated 
by a succession of states of equilibrium.}.  The resulting model is applied to estimating costs of robotic navigation around deformable objects \cite{Frank2011}.

The advances of this work with respect to these related approaches is summarised in \tref{t:advances}.

\begin{table}[htp]
  %\renewcommand{\arraystretch}{1.3}
  \caption{Advances over previous work}\label{t:advances}
  \begin{tabular}{lccccccccc}
  %\hline\hline \\ [-1.5ex]
  \hline \\ [-1.5ex]
 & \begin{sideways}\textbf{Model type}\end{sideways} & \begin{sideways}\textbf{Learned from GT}\end{sideways} & \begin{sideways}\textbf{GT=Real data}\end{sideways} & \begin{sideways}\textbf{Prediction of Shape} \end{sideways} & \begin{sideways}\textbf{Prediction of Forces}\end{sideways} & \begin{sideways}\textbf{Actuator Retrieval} \end{sideways} & \begin{sideways}\textbf{Elastic}\end{sideways} & \begin{sideways}\textbf{Plastic}\end{sideways} & \begin{sideways}\textbf{Classification}\end{sideways} \\
 \hline\hline \\ [-1.5ex]
  Howard & MS & \gtick  & \gtick  & \rcross & \gtick & \rcross & \gtick & \rcross & \rcross\\
Teschner & MS & \rcross & \rcross & \gtick & \rcross & \gtick & \gtick & \gtick & \rcross \\
  Morris & MS & \gtick  & \rcross & \gtick & \rcross & \rcross & \gtick & \rcross & \rcross \\
  Frank & FEM & \gtick  & \gtick & \gtick & \rcross & QSA & \gtick & \rcross & \rcross \\
  This work & MS & \gtick  & \gtick & \gtick & \gtick & \gtick & \gtick & \gtick & \gtick \\
         & + &&&&&&&&\\
         & Regression &&&&&&&& \\
 %[-1.5ex] \\ \hline\hline \\ [-1.5ex]
 [-1.5ex] \\ \hline \\ [-1.5ex]
 \end{tabular}
 *MS = \textit{Mass-spring}\\
 *QSA = \textit{Quasi-Static Assumption}
\end{table}

Other interesting variants are: Conti et al.\ \cite{Conti2003} introduce six-degrees of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects. The open source project Chai3D includes an implementation of this model. The spheres are placed along the skeleton of the object and are connected together with elastic links which model elongation, flexion and torsion properties. Each vertex of the surface mesh is attached to the nearest sphere or link with a damped spring.  In \cite{Burion2008}, Burion and Baur, collaborated with Conti's group to automatically calibrate this model using particle filters.  Again, there is no analysis of the dynamics and no real objects are used.  Furthermore, this model is more adequate for objects that deform preferably around fixed joints.

Finally, Cretu et al., used growing neural gas networks to learn to predictively track the deformation of objects.  Even though they managed to make predictions about the deformation of the overall shape of the object, under previously unseen sets of forces, their predictions were evaluated only for the next frame.

Additionally, there is a vast amount of literature on modelling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.  However, these types of models have not been successfully applied to robotic tasks beyond what was presented above.

%\subsection{Predicting the Shape}
%Contribution: we analyse the dynamic behaviour of the predictions for several frames.
%\subsection{Predicting the Force}
%Physics based models are used to predict reaction forces (surgery simulators).
%Contribution: we learn to predict with a linear regression model.  Also calibrated automatically.

\section{A Mass-Spring Model of Deformation}
\label{sec:model}
\subsection{Elastic Deformation}

\begin{figure}[!t]
\centering
%\includegraphics[width=2.5in]{arrio1}
\includegraphics[width=88mm]{arrio1}
\caption{Spring Forces. Recovery of: (a) Length, forces act along the spring.  (b) Area, forces act along the heights of the triangles. (c) Angle, masses slide trying to restore the angle.}
\label{fig:forces}
\end{figure}

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004} was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  The potential energy terms for the preservation of length with damping and preservation of area were included in our system.  To compensate for the lack of a preservation of volume, which tends to avoid distortions of the graph, a new term for the preservation of the angles was included.  The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.

For any preservation term, the potential energy $E$ is proportionally dependent on a preservation constraint $C$, where $k$ is the proportionality constant \eref{eq:EConstraint}.  The force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 E(\vec{p}_0,...,\vec{p}_{n-1}) & = \frac{1}{2}kC^2, \label{eq:EConstraint} \\
 \vec{F}^i(\vec{p}_0,...,\vec{p}_{n-1}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Which with damping becomes:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_0,...,\vec{p}_{n-1},\vec{v}_0,...,\vec{v}_{n-1}) = \nonumber \\
    &\left( -kC-k_d\sum_{0\le j < n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j$ is the velocity of the $j^{th}$ particle connected to $i$.

\subsubsection{Preservation of Length}
The constraint for the potential energy is given by:
\begin{equation}
 C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0},
\end{equation}
where $D_0$ is the rest length of the spring.  This force pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].  Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.  Morris \cite{Morris2008} omitted that in his final equations for the force.  Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.
\begin{align}
 \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
 & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
\end{align}
where $k_L$ is the linear spring constant and $k_d$ the damping constant.

\subsubsection{Preservation of Area}
Since Teschner and Morris didn't find it helpful to add damping to the preservation of areas, it is avoided here.  Therefore the force $\vec{F}^i_A$ is derived as follows:
\begin{align}
 C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
 Area &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
 \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( Area-A_0 \right) \nonumber \\
 & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
\end{align}
where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.

It is possible to rewrite \eref{eq:area_force} in a more geometrically intuitive manner by rewriting the direction of the gradient according to Morris' geometric derivation \cite{Morris2008}, but keeping the area normalisation constant\footnote{This substitution is valid because both are expressions for the gradients of the area of the triangle formed by the three verticies.}.  It becomes evident that the forces pull along the heights of the triangles [\fref{fig:forces}(b)].  Then the magnitude of the force is given by:
\begin{equation}
 forcemag_A(\vec{p}_i)= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0^2}
\end{equation}

While the rest becomes:
\begin{align}
 \vec{F}_A(\vec{p}_i) & = k_A \cdot forcemag_A(\vec{p}_i) \cdot forcedir_A(\vec{p}_i) \nonumber \\
 forcedir_A(\vec{p}_i) & = \dfrac{\vec{F}_A(\vec{p}_i)}{|\vec{F}_A(\vec{p}_i)|} = \dfrac{areagradient(\vec{p}_i)}{|areagradient(\vec{p}_i)|} \nonumber \\
 areagradient(\vec{p}_i) & = (\vec{p}_i-\vec{p}_j)- \nonumber \\
 & \left( (\vec{p}_k-\vec{p}_j) \cdot \dfrac{(\vec{p}_k-\vec{p}_j) \cdot (\vec{p}_i-\vec{p}_j)}{(\vec{p}_k-\vec{p}_j) \cdot (\vec{p}_k-\vec{p}_j)} \right)
\end{align}

\subsubsection{Preservation of Angles}
The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.  For this reason we included a new term that enforces the preservation of the original angles of the triangles.  [\fref{fig:forces}(c)] gives an idea of how these forces look.  Here the energy depends on the difference of the angles between adjacent edges.

Energy\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
\begin{align}
 E_\varphi(\varphi) & = \frac{1}{2}k_\varphi(\varphi - \varphi_0)^2 \\
 \varphi(p_i,p_j,p_k) & = \arccos \left( \dfrac{(p_j - p_i) \cdot (p_k - p_i)}{ \left\| p_j - p_i \right\| \left\|p_k - p_i \right\|} \right)^2 \nonumber
\end{align}

Where $\varphi$ is the angle between adjacent edges, $E_\varphi$ is the energy associated to changes in the angle, $k_\varphi$ is the corresponding stiffness constant and the $p_i$s are the Cartesian coordinates of the mass particles.

The force emerging from this term is a linear combination of the vectors along the edges that form the angle of interest, acting in the direction of the gradient.  It tends to restore the angles in the most efficient way, but does not take the original size into account [\fref{fig:forces}(c)].  Therefore, it helps to recover a similar triangle, but it also may produce tiny or very big triangles if it is not accompanied by some of the other terms that tend to restore the original dimensions, in addition to the angles.
The force is derived from:  %An additional line in the code also forces an inverted angle to recover its original orientation.
\begin{align}
 F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
 \dfrac{\partial \varphi}{\partial p_i} & = \dfrac{\partial}{\partial p_i} arccos \left(\frac{(\vec{p}_j-\vec{p}_i)\cdot(\vec{p}_k-\vec{p}_i)}{\left\|\vec{p}_j-\vec{p}_i\right\|\left\|\vec{p}_k-\vec{p}_i\right\|}\right) \displaybreak[0] \\
\end{align}

The force is:
\begin{align}
 F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
 \dfrac{\partial \varphi}{\partial p_i}(p_i) & =
    \dfrac{1}{d_{ji}d_{ki}\sqrt{1-\left[\dfrac{pp}{(d_{ji}) (d_{ki})}\right]^2}} \nonumber \\
 & \left\{\left[1-\dfrac{pp}{d_{ki}^2}\right](p_k-p_i)+\left[1-\dfrac{pp}{d_{ji}^2}\right](p_j-p_i) \right\} \nonumber \\
 pp(p_i) & =(p_j-p_i)\cdot(p_k-p_i) \nonumber \\
 d_{ji}(p_i) & = \left\| p_j-p_i \right\| \nonumber \\
 d_{ki}(p_i) & = \left\| p_k-p_i \right\|
\end{align}

The terms for the preservation of length and angle model elastic deformations.  Even though the preservation of area by itself would allow some plasticity, it is still necessary to incorporate permanent deformations in the rest lengths of the springs, to stop the triangles of the mesh from trying to recover their original shape.

\subsection{Plastic Deformation}
\begin{figure}
 \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
 \STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \IF {$\alpha > max\_\alpha$}
  \STATE $l_0 \gets l$
 \ELSE
  \STATE $l_0 \gets (1.0 - \alpha) D_0$
 \ENDIF
 \STATE $elastic\_deformation \gets \frac{l_0 - l}{D_0}$
 \IF {$elastic\_deformation > yield$}
  \STATE $\alpha \gets \alpha + creep * elastic\_deformation$
  \IF {$\alpha > max\_\alpha$}
   \STATE $\alpha \gets max\_\alpha$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
\caption{Algorithm to calculate the permanent deformation of an edge.}
\end{figure}

The simplest method to model plastic deformation consists in changing the length at rest $D_0$ of the springs, in \eref{eq:FDlength}.  The effect of the permanent deformation can be either to permanently compress this length, or to permanently expand it (up to the breaking point).  However, arbitrarily changing the value of $D_0$ can produce geometrical oddities like collapsed edges.  For this reason, in this work, the permanent modification is expressed in terms of a constant of compression $\alpha$, with respect to the original length $D_0$, so that the effective value of the length at rest becomes $l_0 = (1.0 - \alpha) D_0$.  The value of $\alpha$ changes every time the spring is deformed beyond the threshold value \textit{yield}.  A \textit{maximum deformation} limit is introduced to avoid collapsing edges due to compression: if the new length $l$ is smaller than a given fraction of the original length at rest (e.g. $l=0.2 D_0$), the length at rest suffers no further modifications.  See \alref{alg:deformEdge}.

\subsection{Integration Scheme}
\label{sec:ischeme}
Rather than using the traditional Newton equations with a numeric integration scheme like Euler of Verlet, we made the velocities proportional to the forces:
\begin{equation}
 \vec{x}(t+h) = \vec{x}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{x}(t+h)$, the position of the vertex at time $t+h$.  The simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measuremente.

\subsection{Collision Detection}
In the scenario, rigid objects are treated as geometric obstacles.  Two types of obstacles are considered: circles and a floor line that can not be crossed.  For the floor obstacle, the collision is roughly solved by pushing vertices back to the border of the floor.  The springs will eventually propagate the effect to the neighbours.  For the circle the resolution is in two stages: vertices and edges.  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.  For edges, the common algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the circle and the edge.  Given this distance, the algorithm in \alref{alg:push_edge} indicates how and in which direction the edge must be displaced.

\begin{figure}
  \begin{algorithmicieee}{Closest Point on a Segment to Any Other Point}\label{alg:point_segment}
 \STATE Be $a$, $b$ the ends of the edge $\overline{AB}$, $c$ the external point, and $x$ the closest point to $c$ in the line.
 \STATE Be the vectors $\vec{AB} = b - a$ and $\vec{AC} = c - a$.
 \STATE The projection of $\vec{AC}$ over $\vec{AB}$ is $\|x\| = \dfrac{\vec{AC} \cdot \vec{AB}}{\|\vec{AB}\|}$
 \IF{$\|x\| < 0$}
  \STATE $closest \gets a$
 \ELSE
  \IF{$\|x\| > \|\vec{AB}\| $}
   \STATE $closest \gets b$
  \ELSE
   \STATE $closest \gets a + \|x\| \cdot \dfrac{\vec{AB}}{\|\vec{AB}\|}$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
 \caption{Algorithm to find the closest point on a segment to any other point.}
\end{figure}

\begin{figure}
  \begin{algorithmicieee}{Push Edge Out}\label{alg:push_edge}
 \STATE Be $a$, $b$ the ends of the edge $\overline{AB}$, $c$ the centre of the circle.  Use algorithm \ref{alg:point_segment} to find the closest point to the centre in the edge.
 \IF {$\|(closest - c)\| < radius$ (There is an intersection)}
  \STATE $displacement = (radius - \|(closest - c)\| + \epsilon) * \dfrac{(closest - c)}{\|(closest - c)\|}$
  \STATE $ a \gets a + displacement$
  \STATE $ b \gets b + displacement$
 \ENDIF
\end{algorithmicieee}
 \caption{Algorithm to push an edge out of a circular obstacle.}
\end{figure}

\subsection{Geometric Constraints}
Even with the addition of the term for preservation of angles, the discretized mass-spring forces can cause triangles of the mesh to overlap each other and be reversed or to be flattened.  The option of giving additional aid through small geometric subroutines was evaluated as well.  There are two geometric constraints that the mesh is enforced to maintain:

\begin{enumerate}
 \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed beyond half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.  This is enough for most cases, but triangles in the border of the mesh can still overlap each other.

 \item If a triangle becomes flat: the longest edge and its opposite vertex along the perpendicular to the edge are pushed in opposite directions, to form a tiny triangle\footnote{An area of $2.0$ units was good enough.}.
\end{enumerate}

\subsection{Many Steps Prediction}
\label{sec:manysteps}

Given the initial shape of the object and a set of parameters, the mass-spring system allows to make predictions of the deformation of an object for various interactions.  Given the procedures explained in the previous sections, there are three ways to calculate the subsequent deformations of the shape:
\begin{enumerate}
 \item \textbf{Collision detection only.} Collision detection is used to determine the compression of the springs around the actuator.  In this case, the neighbouring springs are compressed and their preservation forces propagate the deformation towards their neighbours, and so on.  There is no need to use the force sensor.  It can be equated with observing someone else pushing an object we have touched before.
 \item \textbf{Forces on vertices.} The forces to be applied on the  vertices around the actuator are approximated by the measured reaction forces.  Both, the series of positions of the actuator and the reaction forces are necessary.  These forces will be solely responsible for the deformation of the mesh.  If the deformation is insufficient, the actuator may trespass invalidly the border of the mesh.  The quality of the simulation is much more sensitive to the values of the elasticity-plasticity constants.  This corresponds to the robot itself pushing the object and feeling the response, while using vision solely to determine the position of the actuator and for evaluation purposes.
 \item \textbf{Both.} The previous two are combined.  Visual constraints (collision detection) and forces are used to determine the displacement of nodes around the actuator.  Again, the springs propagate the deformation to the rest of the mesh.
\end{enumerate}
The mesh as predicted by one step is fed as current shape for the next step and the cycle is repeated until the simulation is stopped.

The next stage of the learning process consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.

\section{Learning to Predict the Deformation}
\label{sec:learning}
\subsection{Evolutionary Algorithm}
The evolutionary algorithm listed in \alref{alg:genetic} was used to search the space of parameters of the mass-spring model presented in the previous section.
From the parameters that must be chosen for the simulation to work, some were fixed while others had to be searched for.  They are:
\begin{itemize}
 \item The mass per vertex.  \textit{(For the moment we consider a unitary mass for all vertices.)}
 \item The elasticity constants: for the preservation of length $k_L$, area $k_A$, angles $k_{\varphi}$ and linear damping $k_d$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well.)}
 \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
 \item The minimum area before a triangle is considered \textit{flat}.  (It has been set to $2.0$ units.)
\end{itemize}
To evaluate the suitability of each set of parameters, the simulation is compared with the ground truth tracked from real objects.

\begin{figure}
  \begin{algorithmicieee}{Evolutionary Search}\label{alg:genetic}
 \STATE Be $N$ the number of sets of parameters to be evaluated per generation and $M$ the number of generations to try.
 \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements.
 \STATE Be $\sigma_{max} > \sigma_{min}$ standard deviations for a Gaussian
 \STATE Let $\Delta\sigma \gets (\sigma_{max} - \sigma_{min})/M$.
 \FOR {$i=1 \to M$}
 \STATE Run the $N$ simulations and compare prediction with the ground truth (real object), by comparing the area occupied by the real object with the area occupied by the mesh of the simulation [\fref{fig:scene}(b)].
 \STATE Keep the best $\alpha$ of all simulations.
 \STATE $\sigma \gets \sigma - \Delta\sigma$
 \STATE Generate $\beta$ by adding random Gaussian noise, with standard deviation $\sigma$, to the best candidates.
 \STATE Generate $\gamma$ by selecting values of the parameters from two different sets at random, after eliminating the $\epsilon$ worst.
 \FOR {$\zeta$ elements}
   \STATE Generate completely new sample with:
   \FOR {every parameter}
     \STATE $n \leftarrow randomNumber \in [0,1)$
     \STATE $min \leftarrow \log(minValue)$
     \STATE $max \leftarrow \log(maxValue)$
     \STATE $param \leftarrow e^{(n (max - min) + min)}$
   \ENDFOR
 \ENDFOR
 \STATE $i \gets i + 1$
 \ENDFOR
\end{algorithmicieee}
 \caption{Evolutionary Search for Automatic Calibration}
\end{figure}

\subsection{Tracking to Provide Ground Truth}

\begin{figure}[!t]
\centering
\includegraphics[width=88mm]{arrio2}
\caption{(a) Experimental set up.  A robotic finger pushes a block of deformable material over a table.  A camera registers the events of the plane where the main deformations take place.  (b) Evaluation function.  Left top: Mesh over real block of material.  Left bottom: Tipical error of the model vs the ground truth (red) and tipical error of the tracker (blue).  Right top: tracked ground data.  Middle: TP in yellow, FP in blue, FN in red and TN in gray.  Bottom: simulated mesh. }
\label{fig:scene}
\end{figure}

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline}.  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains were it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

\begin{figure}
 \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
 \STATE Be $L$ a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
 \FORALL {$L_i$ in $L$}
 \IF {$dist(L_i,L_{i-1}) > max\_distance$}
  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
 \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
  \STATE erase $L_i$
 \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
  \STATE erase $L_i$
 \ENDIF
 \ENDFOR
\end{algorithmicieee}
\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
\end{figure}

\subsection{Evaluation Function}
\label{sec:evaluation}

To evaluate the quality of the predicted shapes, the pixels inside the tracked contour are used as ground truth.  These are  compared against those inside the mesh of the simulation.  In machine learning terms, these pixels can be divided among four classes: the \textit{true positives $TP$}, where the material is present, and the simulation predicted it would be present; the \textit{true negatives $TN$}, where the absence of material is in accordance with the prediction; the \textit{false positives $FP$}, where the prediction said there would be material, but there is not; and the \textit{false negatives $FN$}, where there is material, but the prediction said there wouldn't [\fref{fig:scene}(b)].  Given that maintaining a record of the true negatives (the empty/background space around the material) is mostly irrelevant and highly inefficient, the mark for a whole video was derived from the following quantities:
\begin{align}
 Precision &= \frac{TP}{TP+FP}, \\
 Recall &= \frac{TP}{TP+FN}
\end{align}
From which the harmonic average $F$ is derived:
\begin{align}
 F = \frac{2}{\frac{1}{Precision} + \frac{1}{Recall}} = \frac{1}{1 + \frac{FN + FP}{2TP}}
\end{align}

This measurement is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the mark for the video is the mean over all the frames so far, $\mu(F)$.  If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closest to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

\section{Models of Force Learning}
%We used two methods to predict the forces.  The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.  This force is feed as input for the following simulation step.
\label{sec:forces}

Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the training data set [\fref{fig:sstrain}].  The curve in this diagram is approximated by segmented regression curves.  Several candidates are proposed using the least squares technique and the best are chosen as follows.  Since there is only one independent variable the equations for adjusting a line are straightforward:
\begin{align}
 y & = mx + y_0 + Error \\
 m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
 y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
 Error & = \sum(y_i - (mx_i+y0))^2
\end{align}
Where $y$ is the dependent variable, $x$ the independent one, $m$ is the sloop of the line, and $y_0$ the $y$ intercept.  To adjust logarithmic curves a change of variable is used.  Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.  In this case, the final equation will be of the form: $y = m \ln x + y_0$.  It would be possible to add other types of curves, but these were enough for the materials covered.

For the automatic calibration of the pushing phase both types of curves are considered.  To determine the place of the discontinuities between the segmentes, several candidates are calculated.  Each candidate is obtained by adjusting a curve through different amounts of data points and calculating the error.  Beginning with a small number of points (e.g. 5), as new measurements are taken, new curves are adjusted until the corresponding normalized error increases beyond a threshold value.  The type of curve that can cover more points with a low fit error is adopted for the model.  This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.

\section{Experimental Results}
\label{sec:experiments}
\subsection{Scenario}
For the experimental scenario a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth, is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.  A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].  The force sensor was a DAQ-FT-Gamma, that can measure forces up to 200 Newtons in the $z$ direction (up to 65N in the others) and torques up to 5 Newton-metre.  The size of the photographs is $800 \times 600$ pixels.  The scale between the real world measurements and the pixels in the image were: $4600\frac{\unit{pixels}}{\unit{m}}$ \footnote{$\unit{m}=metre$} for the sponge and $4350\frac{\unit{pixels}}{\unit{m}}$ for the plasticine (they are different because the camera was placed at slightly different positions).

For both sets of experiments, each of the blocks was pushed in different positions:
\begin{enumerate}
 \item In the middle of its longest side.
 \item Close to the corner of its longest side.
 \item In the middle of its shortest side.
\end{enumerate}
The first set of data was used to train the model for the respective material.  The set of parameters that allowed the model to behave more like the real object, according to the criterion explained in \sref{sec:evaluation}, was selected.  The other two videos were used to test the quality of the predictions made by the model for the new interactions.

To take into account the cylindrical shape of the actuator, the radial component of the registered force is applied on any vertex that happens to be located within the radius of the robotic finger.  For collision detection routines the finger is considered a circular obstacle.

\begin{figure}[!t]
\centering
\setlength\fboxsep{0pt}    % padding
\setlength\fboxrule{0.5pt} % line width
\fbox{\includegraphics{arrio8}}
%\includegraphics[width=88mm]{arrio8}
\caption{The simulation time system with shape and force prediction.  The position of the actuator is extracted from the image, while the force is predicted by a piece-wise regression model. This information is used by the calibrated mass-spring system to predict the shape of the object for several frames. }\label{fig:diagram}
\end{figure}

\subsection{Integration Scheme}
Sets of 24 training sessions were run, using different combinations of elasticity and plasticity terms.  For example: using damped preservation of length only, using preservation of area and angle only or all of them together;  including plastic deformation or removing the collision resolution routines and the aid of geometric constraints.  For this first rounds of experiments three different integration schemes were used: Euler, Verlet and our Customised version.  The resulting marks are summarised in \tref{tab:ischemes}.

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Marks of Sets of Parameters Obtained With Different Integration Schemes}
\label{tab:ischemes}
\centering
\begin{tabular}{cccc}
\hline
\bfseries Scheme & \bfseries Best & \bfseries Worst & \bfseries Average \\
\hline\hline
Euler & 0.88 $\pm$ 0.04 & 0.44 $\pm$ 0.13 & 0.75 $\pm$ 0.08 \\
Verlet & 0.76 $\pm$ 0.19 & 0.53 $\pm$ 0.18 & 0.68 $\pm$ 0.19 \\
Customised & 0.921 $\pm$ 0.008 & 0.60 $\pm$ 0.13 & 0.86 $\pm$ 0.02 \\
\hline
\multicolumn{4}{l}{Where the mark is $\mu(F)$ as explained in \sref{sec:evaluation}.}
\end{tabular}
\end{table}

The traditional equations of movement, estimated with Euler or Verlet show big oscillations and frequently become unstable.  Verlet seems to be badly affected by the collision resolution routines and geometric aids, that affect drastically the values of the velocity and acceleration.  On the contrary, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that resemble more the ground truth, thus obtaining better marks.  From this group of experiments it was also found that the evolutive learning is as efficient when considering all elastic and plastic terms, as when using only a few of them.

\subsection{Parameters Calibrated Automatically}
\subsubsection{Performance of the Learning Algorithm for the Mass-spring Model}
The performance of the genetic learning algorithm was evaluated using the customised integration scheme and the mass-spring elastic-plastic model.  At this point no prediction of the force was included, rather, the reaction forces measured with the force sensor were used.  \fref{fig:performance} shows how the quality of the simulation throughout the whole video improved with each generation.  Both graphs show that even though the enforcing of geometric constraints with aid routines may produce simulations with lower marks at the begining of the training, they perform just as well after enough generations.  These marks reflect the degree at which areas covered by the ground truth and the mesh of the simulation match; however, when looking at the videos produced, it becomes evident that simulations with geometric aids succeed more in keeping the mesh flat. % Should put images? (if there is space left)

\begin{figure}[!t]
\centering
\includegraphics[width=88mm]{arrio6}
\includegraphics[width=88mm]{arrio7}
\caption{Final marks of the videos after each generation of the genetic search.  The filled areas correspond to the standard deviation $\sigma(\bar{\mu}(F))$, across 15 runs of the learning algorithm.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with and without the aid of geometric recovery routines.  The block of material was represented by a mesh of $20 \times 10$ elements.}\label{fig:performance}
\end{figure}

\subsubsection{Prediction of the Force}

\begin{figure}[!t]
\centering
\includegraphics[width=88mm]{arrio4}
\includegraphics[width=88mm]{arrio5}
\caption{The response force during the interaction can be characterised in three phases: pushing, transition and retrieval of the finger.  A different regression curve can be used to approximate the force during the pushing and retrieving phases.  Linear and logarithmic regressions were suggested for the pushing movement and the best was chosen autonomously.  Only linear regression was used while retrieving.}\label{fig:sstrain}
\end{figure}

Using the readings of the force sensor, and associating the deformation of the material with the displacement of the robotic finger, a regression curve was fitted to the stress-strain graph for the pushing and retrieving movements.  See \fref{fig:sstrain}.  As it was expected from typical stress-strain diagrams of elastic and plastic phases of materials, a linear regression is characteristic of an elastic material, while the logarithmic fits better to a plastic one.  For the retrieving movement only a linear model was considered, since the shape deformation of the plastic material makes it harder to measure the strain from the visual information.  The curves obtained are used to approximately predict the reaction force for unknown interactions.  These forces are used for the mass-spring model as in the diagram of \fref{fig:diagram}.

\subsection{Prediction}

One hundred photographs and force readings where taken every $0.17\pm0.007 s$ in average.  The mass spring model runs with an integration step of $0.1s$, where forces that were not measured directly are interpolated.  The push down movement lasts approximately 50 frames and the retrieval, the other 50.  The continuous lines in \fref{fig:sstrain} show the predicted forces, given the amount of compression of the material (strain), while the point-lines show the actual readings.  \fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the shapes predicted by the mesh.

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio9}
\includegraphics[width=178mm]{arrio10}
\includegraphics[width=178mm]{arrio11}
\caption{Simulation of the sponge obtained with parameters $k_L=1898.22, k_d=1.24705, k_A=18629.2 k_{\phi}=764.649, max\_Force=183.71, yield=0.558793, creep=0.433517, max\_\alpha=0.73364$.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements. Top: Training. Middle: Test 1. Bottom: Test 2}\label{fig:simulation_sponge}
\end{figure*}

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio12}
\includegraphics[width=178mm]{arrio13}
\includegraphics[width=178mm]{arrio14}
\caption{Simulation of the plasticine obtained with parameters $k_L=4800.43, k_d=0.803787, k_A=7475.74 k_{\phi}=72.7379, max\_Force=571.392, yield=0.848239, creep=0.151826, max\_\alpha=0.126444$.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements. Top: Training. Middle: Test 1. Bottom: Test 2}\label{fig:simulation_plasticine}
\end{figure*}

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio16}
\caption{Simulation of the plasticine obtained with parameters $k_L=0, k_d=0, k_A=28202.1 k_{\phi}=34.5491, max\_Force=780.825, yield=0.772727, creep=0.67466, max\_\alpha=0.656572$.  Using the customised integration scheme, the elastic-plastic model with some preservation terms, furtherly tuned by hand,  with the aid of geometric recovery routines with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements.}\label{fig:simulation_plasticine_hand}
\end{figure*}

\subsection{Classification}

The calibrated mass-spring models were used to classify objects through recognition of their behaviour, even for novel interactions.  The models make predictions as if the object was known.  The model that makes the best predictions indicates which object is being observed.  There are two criteria for selecting an object:

\begin{enumerate}
 \item Per frame: an object is selected for each frame of the interaction. See \tref{tab:classperframe}.

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Classification per Frame}
\label{tab:classperframe}
\centering
\begin{tabular}{ccccccccc}
\hline
 & & \multicolumn{3}{c}{\bfseries Sponge} & & \multicolumn{3}{c}{\bfseries Plasticine} \\
Data Set & & TP & FN & Precision & & TP & FN & Precision\\
\hline\hline
 \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
\hline
   training & & 167 & 132 & 0.56 & & 186 & 113 & 0.62 \\
   test1 & & 167 & 132 & 0.56 & & 271 & 28 & 0.91 \\
   test2 & & 242 & 57 & 0.81 & & 289 & 10 & 0.97 \\ 
\hline
%\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
\end{tabular}
\end{table}

 \item Global: The difference in the marks is added for all frames.  Frames where the distinction is more noticeable will receive a greater weight.
\end{enumerate}

\section{Discussion}
\label{sec:discussion}
We note the following points about the results and the method in general.
\subsection{The Learning Method}
Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to improve the quality of the simulations as much as it could.  Better simulations have been obtained by hand tunning some of the parameters.  \sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.  The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}

\subsection{Using Geometric Constraints}
Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.   Also, when attempting to model the plastic material, the geometric constraints provoqued a much more notorious effect comming from any plasticity term, even if its corresponding constant was small.  As a consequence the plastic material also recovered its original shape after some extra frames.  It has been seen that, if the constant for the linear preservation is set to zero, the problem disappears.  Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.  We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the begining, instant of greatest deformation and final shape.  Meanwhile, the evaluation function gives the same weight to all frames in the sequence.  It is possible to use the force learning algorithm and the commands given to the robot (push/retreive) to detect these points, since they are corelated to discontinuities in these other modalities.  Future work will evaluate the impact of these modifications on the evaluation function.

\subsection{Training With Collision Resolution with the Finger vs. Without}
It was found that the overall mark of the best simulations was similar in both cases, since only geometric similarities are under consideration.  However, there was a notorious difference on the rate of success for the classification problem.  If no collision detection with the finger is used during training, the springs react only to the forces and, if the spring coefficients are not adequate, the shape of the mesh will be very different from the shape of the ground truth.

On the other side, if collision detection is used, the geometric constraints strongly favour a correct configuration of the mesh in the vicinity of the actuator, even though the springs did not react properly.  For subsequent frames, it is more likely that the neighbours will succeed in propagating the deformation to the rest of the material, than if the positions of the springs had not been tampered by the collision resolution, thus producing better simulations.  However, this lessens the effectiveness of the mass-spring model in distinguishing the materals.

\subsection{Classification}
The best results for classification with the mass-spring model are obtained when the training of the models uses only forces to displace the nodes in the mass-spring model\footnote{Case 2 in \sref{sec:manysteps}}, but the combination of force readings and collision resolution is used during classification\footnote{Case 3 in \sref{sec:manysteps}}.  This finding agrees with the fact that the objects are more distinguishable from the data obtained from the force sensor, than from the visual data.  The fact that the spring constants must respond to the applied forces during the training and not only to the visual appearance is reflected in a better differentiation of the behaviour of the models.  When the calibration is more sensitive to the constants of the springs, it is more robust for novel interactions, therefore the difference in quality of the predictions is a better indicator of which material is beeing used.

\subsection{Conclusions}
In summary we have presented a framework, together with specific algorithms for learning the deformable behaviour of objects from robotic interaction. The main properties of the model are as follows. First, the parameters for the model were adjusted to data from real objects under robotic interaction. Second, we have attempted the automatic callibration of a simulation of plasticity. Third we use regression to predict the reaction forces. Fourth we include a new energy term for the preservation of angles of the mesh's triangles. Fifth we employ a regularization algorithm for a linear snake, that increases or decreases the number of its control points, according to the level of detail required to represent the deformed object. Sixth we provide a detailed evaluation of the dynamic simulation, instead of only during its stable states as in previous work. Finally we showed the application of the model to the problem of classifying deformable materials.

There is considerable future work to be done. The first extension is due to the fact that we restricted the deformation model to 2D, and this can easily revert to the 3D meshes as used in Teschner's original work. The second issue is to explore new methods for the learning phase. The current search procedure doesn't always find good solutions for the plastic deformation model. But a more basic problem is that mass-spring models restrict the range of deformations that can be modelled. While the framework presented here is promising, we regard the actual deformation model representation as an open research problem.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% \appendices

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
% \section{}
% Appendix two text goes here.

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank Rustam Stolkin for his help in setting up the data gathering experiments and for the construction of the robotic finger. The research leading to this results has received funding from the PhD Scholarships program of CONACYT, Mexico and from the European Community's Seventh Framework Program [FP7/2007-2013] under grant agreements 215181 CogX, and 600918 PaCMan.

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
\bibliographystyle{IEEEtranS}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{twostagedo}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \begin{thebibliography}{1}
% 
% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
% 
% \end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{arrio}}]{Veronica E. Arriola-Rios}
  has recently obtained her PhD in Computer Science from the University of Birmingham.
  She works as full time lecturer at the National Autonomous University of Mexico (UNAM).
  Veronica holds undergraduate degrees in Physics and Computer Science, and a masters degree in Computer Sciences, in the area of Computer Graphics and Virtual Environments.
  Her research interests include artificial inteligence, cognitive robotics and natural cognition.
\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{wyatt}}]{Jeremy L. Wyatt} is Professor of Robotics
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{wyatt}}]{Jeremy L. Wyatt}
%   is Professor of Robotics and Artificial Intelligence at the University of Birmingham. He
%   obtained a BA in Theology from the University of Bristol, an MSc in
%   Knowledge Based Systems from the University of Sussex, and his
%   Ph.D. in Artificial Intelligence from the University of Edinburgh in
%   1997. He has published more than 80 refereed articles and edited three books. His interests
%   include machine learning, robot manipulation, robot learning, robot vision, 
%   robotic architectures, and planning and learning for robots in uncertain 
%   and incomplete worlds.
  is Professor of Robotics
  and Artificial Intelligence at the University of Birmingham. He
  obtained a BA in Theology from the University of Bristol, an MSc in
  Knowledge Based Systems from the University of Sussex, and his
  Ph.D. in Artificial Intelligence from the University of Edinburgh in
  1997. He has published more than 80 refereed articles. His interests
  include machine learning, planning, architectures, vision, mobile robotics and robot manipulation.
\end{IEEEbiography}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{stolk}}]{Rustam Stolkin}
%%  is currently Senior Research Fellow in robotics, in the School of
%  Mechanical Engineering, University of Birmingham, UK. Prior to
%  coming to Birmingham, he was a Research Assistant Professor at
%  Stevens Institute of Technology, USA. Rustam holds undergraduate and
%  masters degrees in Engineering Science from University of Oxford,
%  and a PhD in Robot Vision, undertaken between University College
%  London and the UK imaging industry. His research interests include
%  robotics, computer vision, sensor systems, and science and
%  engineering education
%\end{IEEEbiography}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2016-04-12 twostagedo.tex
Plasticine images updated.

@@
 \centering\includegraphics[width=3.5in]{arrio17.png}
-\caption{Force information acquired while working with a block of sponge (elastic) and a piece of plasticine (plastic).}
+\caption{Force information acquired while working with a block of sponge (elastic) and a piece of plasticine (plastic).
+Notice the difference in magnitude between the force used with the sponge and the plasticine.}
 \label{fig:forceElasticPlastic}
@@
 \begin{align}
- E(\vec{p}_0,...,\vec{p}_{n-1}) & = \frac{1}{2}kC^2, \label{eq:EConstraint} \\
- \vec{F}^i(\vec{p}_0,...,\vec{p}_{n-1}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
+ E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint} \\
+ \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
 \end{align}
@@
  \label{eq:Fdamping}
- \vec{F}^i&(\vec{p}_0,...,\vec{p}_{n-1},\vec{v}_0,...,\vec{v}_{n-1}) = \nonumber \\
-    &\left( -kC-k_d\sum_{0\le j < n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
+ \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
+    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
  \dfrac{\partial C}{\partial \vec{p}_i},
 \end{align}
-where $\vec{v}_j$ is the velocity of the $j^{th}$ particle connected to $i$.
+where $\vec{v}_j$ is the velocity of the $j^{th}$ particle connected to $i$, in a system with $n$ particles.
 
@@
 \begin{equation}
- C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0},
+ C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0}, \label{eq:clength}
 \end{equation}
-where $D_0$ is the rest length of the spring.
-This force pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].
+where $\vec{p}_i$ and $\vec{p}_j$ are connected by a spring and $D_0$ is the rest length of that spring.
+The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].
 Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.
@@
 Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.
+Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping} we get:
 \begin{align}
@@
 \subsubsection{Preservation of Area}
-Since Teschner and Morris didn't find it helpful to add damping to the preservation of areas, it is avoided here.  Therefore the force $\vec{F}^i_A$ is derived as follows:
+This constraint is applied per triangle in the mesh.
+Since Teschner and Morris didn't find it helpful to add damping to the preservation of areas, it is avoided here.
+Therefore the force $\vec{F}^i_A$ is derived as follows:
 \begin{align}
@@
 \includegraphics[width=178mm]{arrio14}
-\caption{Simulation of the plasticine obtained with parameters $k_L=4800.43, k_d=0.803787, k_A=7475.74 k_{\phi}=72.7379, max\_Force=571.392, yield=0.848239, creep=0.151826, max\_\alpha=0.126444$.
-Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines with the finger.
+\caption{Simulation of the plasticine obtained with parameters $k_L=642.328, k_d=0.562201, k_A=29.2776 k_{\phi}=72.6202, max\_Force=765.912, yield=0.0790105, creep=0.664469, max\_\alpha=0.48842$.
+Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.
 The block of material was represented by a mesh of $20 \times 10$ elements.

### 2016-04-13 twostagedo.tex
small changes to emphasise role of sensorimotor contingencies, and to remove overview figure.

@@
 % The paper headers
-\markboth{IEEE Transactions on Systems, Man, and Cybernetics—Part B: Cybernetics,
-          ~Vol.~11, No.~4, December~2012}%
+\markboth{IEEE CDS}%
 {Arriola-Rios \MakeLowercase{\textit{et al.}}: A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
@@
 Most objects that are manipulated by animals in natural environments are deformable, and these include objects that are elastic, plastic, breakable, tearable, and spreadable.
-This work is based on the premise that in order to plan manipulations of deformable objects, predictive models of object deformation are required.
-Without such models planners cannot reason about the effects of actions, or sequences of actions to achieve desired effects.
+This work is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required.
+These sensorimotor contingencies can then be used to feedback on themselves to predict over long timescales.
+Without such models planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.
 
@@
 
-Previous work on prediction for deformable objects in robotics has been suprisingly limited.
+Previous work on learning sensorimotor contingencies for deformable objects in robotics has been suprisingly limited.
 Table~\ref{t:advances} summarises the features of the major efforts to date.
@@
 It is important to note, however, that the model structure is separate from the specific model forms used, and that other model types could be substituted for one or both stages.
-This behaviour is described in \alref{alg:overall}.
-
-\begin{figure}
- \begin{algorithmicieee}{Overall Behaviour}\label{alg:overall}
- {\small
- \STATE \textbf{Training Phase:}
- \FORALL {training materials:}
-   \FORALL {finger commands (push down, push up)}
-     \STATE The controler gives the command.
-     \WHILE {the command is executed}
-       \STATE The F/T sensor module records the reaction force.
-       \FORALL {candidate regresion curves}
-         \STATE The F/T sensor module adjusts the curve.
-       \ENDFOR
-       \STATE The visual system tracks and records the modifications of the shape from a 2D camera.
-     \ENDWHILE
-     \STATE The F/T sensor module records a discontinuity in the regresion curve.
-   \ENDFOR
-   \STATE The F/T sensor module returns the best regresion curve as Model B.
-   \STATE The learning algorith for the mass-spring model proposes candidate parameters.
-   \FORALL{frames of captured information}
-     \FORALL{sets of candidate parameters}
-       \STATE Evaluate how similar the mass-spring simulation is to the real object.
-     \ENDFOR
-   \ENDFOR
-   \STATE The values of the best set of parameters is adopted as Model A.
- \ENDFOR
- \STATE \textbf{Simulation Phase:}
- \WHILE{the finger penetrates the material}
-   \STATE Use Model B to predict the reaction force given the distance traversed after contact with the material.
-   \STATE Use this force to run the simulation with Model A.
-   \STATE Model A predicts the shape the material should have.
- \ENDWHILE
- \STATE \textbf{Classification Phase:}
- \FORALL{frames of captured information}
-  \FORALL{previously trained materials}
-    \STATE Run simulation and evaluate similitude
-  \ENDFOR
- \ENDFOR
- \STATE Return material with the highest score.}
- \end{algorithmicieee}
-\caption{Overall algorithmic structure of the system.}
-\end{figure}
+%This behaviour is described in \alref{alg:overall}.
+
+%\begin{figure}
+% \begin{algorithmicieee}{Overall Behaviour}\label{alg:overall}
+% {\small
+% \STATE \textbf{Training Phase:}
+% \FORALL {training materials:}
+%   \FORALL {finger commands (push down, push up)}
+%     \STATE The controler gives the command.
+%     \WHILE {the command is executed}
+%       \STATE The F/T sensor module records the reaction force.
+%       \FORALL {candidate regresion curves}
+%         \STATE The F/T sensor module adjusts the curve.
+%       \ENDFOR
+%       \STATE The visual system tracks and records the modifications of the shape from a 2D camera.
+%     \ENDWHILE
+%     \STATE The F/T sensor module records a discontinuity in the regresion curve.
+%   \ENDFOR
+%   \STATE The F/T sensor module returns the best regresion curve as Model B.
+%   \STATE The learning algorith for the mass-spring model proposes candidate parameters.
+%   \FORALL{frames of captured information}
+%     \FORALL{sets of candidate parameters}
+%       \STATE Evaluate how similar the mass-spring simulation is to the real object.
+%     \ENDFOR
+%   \ENDFOR
+%   \STATE The values of the best set of parameters is adopted as Model A.
+% \ENDFOR
+% \STATE \textbf{Simulation Phase:}
+% \WHILE{the finger penetrates the material}
+%   \STATE Use Model B to predict the reaction force given the distance traversed after contact with the material.
+%   \STATE Use this force to run the simulation with Model A.
+%   \STATE Model A predicts the shape the material should have.
+% \ENDWHILE
+% \STATE \textbf{Classification Phase:}
+% \FORALL{frames of captured information}
+%  \FORALL{previously trained materials}
+%    \STATE Run simulation and evaluate similitude
+%  \ENDFOR
+% \ENDFOR
+% \STATE Return material with the highest score.}
+% \end{algorithmicieee}
+%\caption{Overall algorithmic structure of the system.}
+%\end{figure}
 

### 2016-04-13 twostagedo.tex
Changed illustrations for simulation of sponge.

@@
 
-The training of the model is performed offline using ground truthed data.
-The overall algorithm is given in \alref{alg:overall}.
+The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
 For clarity the algorithm is given in three phases: training, simulation (i.e.
@@
 
-For any preservation term, the potential energy $E$ is proportionally dependent on a preservation constraint $C$, where $k$ is the proportionality constant \eref{eq:EConstraint}.
-The force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
+For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.
+When the value of the constraint is zero, the potential energy is zero as well.
+When the constraint is not satisfied, the potential energy increases.
+This behaviour is obtained through the following definition for the energy:
 \begin{align}
- E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint} \\
+ E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint}
+\end{align}
+where $k$ is a proportionality constant \eref{eq:EConstraint} and $p_i$ is the position of particle $i$.
+
+In order to reduce the energy of the system when a constraint is not satisfied, the force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
+\begin{align}
  \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
 \end{align}
-Which with damping becomes:
+Given the oscilatory nature of these equations, damping can be introduced to stop the oscillations.
+This is modelled with a force acting in the opposite direction to the velocity of the affected particles:
 \begin{align}
@@
 \end{align}
-where $\vec{v}_j$ is the velocity of the $j^{th}$ particle connected to $i$, in a system with $n$ particles.
+where $\vec{v}_j = \frac{\partial \vec{p}_j}{\partial t}$ is the velocity of the $j^{th}$ particle connected to $i$, when the constraint acts on $n$ particles.
 
@@
 Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.
-Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping} we get:
+Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping}, after some manipulation we get:
 \begin{align}
@@
 \includegraphics[width=178mm]{arrio11}
-\caption{Simulation of the sponge obtained with parameters $k_L=1898.22, k_d=1.24705, k_A=18629.2 k_{\phi}=764.649, max\_Force=183.71, yield=0.558793, creep=0.433517, max\_\alpha=0.73364$.
+\caption{Simulation of the sponge obtained with parameters $k_L=3034.48, k_d=0.000237, k_A=103.568 k_{\phi}=66.774, max\_Force=1218.81, yield=0.000434973, creep=0.00890343, max\_\alpha=0.45204$.
 Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.
@@
 \includegraphics[width=178mm]{arrio13}
-\includegraphics[width=178mm]{arrio14}
+%\includegraphics[width=178mm]{arrio14}
 \caption{Simulation of the plasticine obtained with parameters $k_L=642.328, k_d=0.562201, k_A=29.2776 k_{\phi}=72.6202, max\_Force=765.912, yield=0.0790105, creep=0.664469, max\_\alpha=0.48842$.

### 2016-04-14 twostagedo.tex
added some more text to link to journal theme

@@
 The sensed forces are the cause of the observed deformation, and this can be exploited by sequencing the models by mode.
-Thus the input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces.
+Thus our approach can be seen as a sensorimotor contingency in which two contigencies are sequenced to produce an overall prediction.
+The input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces.
 The second predictor takes this sequence of forces as input and predicts a deformation sequence.
@@
 Better simulations have been obtained by hand tunning some of the parameters.
-\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
+%\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
 The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}
@@
 \subsection{Conclusions}
-In summary we have presented a framework, together with specific algorithms for learning the deformable behaviour of objects from robotic interaction.
+In summary we have presented a framework, together with specific algorithms for learning the sensorimotor contingencies for the deformable behaviour of objects under robot manipulation.
+We have developed a learning model that learns the parameters of two contingencies that are sequenced, one for kinematic motion to force and the other for force to deformation.
+By sequencing these we obtained a compound contingency that predicts deformation given motion.
 The main properties of the model are as follows.
-First, the parameters for the model were adjusted to data from real objects under robotic interaction.
-Second, we have attempted the automatic callibration of a simulation of plasticity.
+First, the parameters for the model were adjusted to data from real objects under robotic manipulation.
+Second, we have attempted the automatic calibration of a simulation of plasticity.
 Third we use regression to predict the reaction forces.
@@
 
-The authors would like to thank Rustam Stolkin for his help in setting up the data gathering experiments and for the construction of the robotic finger.
-The research leading to this results has received funding from the PhD Scholarships program of CONACYT, Mexico and from the European Community's Seventh Framework Program [FP7/2007-2013] under grant agreements 215181 CogX, and 600918 PaCMan.
+We gratefully acknowledge funding from the PhD Scholarship program of CONACYT, Mexico and from the European Community's FP7 program, under grant agreements 215181 CogX, and 600918 PaCMan.
 

### 2016-04-14 CoverLetter.tex
Added cover letter.

\documentclass[a4paper,10pt]{letter}
\usepackage[utf8]{inputenc}

%% --- Font
\usepackage{lmodern}
\renewcommand{\familydefault}{\sfdefault}

%% --- Margins

\setlength\topmargin{-50pt}
\setlength\textheight{10in}
\longindentation=0pt

%% --- Logo
\usepackage{graphicx}
%\usepackage{fancyhdr}
%\usepackage{geometry}

\atbegindocument{\includegraphics[width=.98\textwidth]{figures/unam.jpg}}

%\geometry{headheight = 1.1in}
%\fancypagestyle{firstpage}{\fancyhf{}\fancyhead[L]{\includegraphics[height=1in, keepaspectratio=true]{figures/unam.jpg}}}
%\fancypagestyle{plain}{\fancyhf{}\fancyhead[L]{\includegraphics[height=0.5in, keepaspectratio=true]{figures/unam.jpg}}}
%\fancypagestyle{empty}{\fancyhf{}\fancyhead[R]{\includegraphics[height=0.5in, keepaspectratio=true]{figures/unam.jpg}}}
%\pagestyle{plain}

%\vspace*{-2cm}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{IEEE Transactions on Cognitive and Developmental Systems\newline
Special Issue on Sensorimotor Contingencies for Cognitive Robotics}
\address{Facultad de Ciencias, \\Ciudad Universitaria,\\ UNAM, México, D.F.}

\opening{Dear Dr. Ricardo Téllez,}

Please find attached our submission:
\begin{center}
 ``A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation'' \\
 Veronica E.~Arriola-Rios (UNAM, Mexico), Jeremy~L.~Wyatt (University of Birmingham, UK)
\end{center}

This work is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. These sensorimotor contingencies can then be used to feedback on themselves to predict over long timescales. Without such models planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.
Here we present a framework, together with specific algorithms for learning the sensorimotor contingencies for the deformable behaviour of objects under robot manipulation.
The model is multi-modal in that it is based on integrating force and visual information.

I look forward to hearing from you.

\signature{Dra. Verónica Esther Arriola Ríos\\
Profesora Asociada C de T.C.\newline
Departamento de Matemáticas, Cub 119.\newline
Facultad de Ciencias, UNAM \newline
v.arriola@ciencias.unam.mx \newline
+(52)55 5622 5426}

\closing{Yours sincerely,}

%enclosure listing
%\encl{}

\end{letter}
\end{document}

### 2016-08-01 AnswersToComments.tex
Addressed first comments, added AswersToComments

\documentclass[a4paper,12pt]{letter}
\usepackage[utf8]{inputenc}

%% --- Font
\usepackage{lmodern}
\renewcommand{\familydefault}{\sfdefault}

%% --- Margins
\usepackage[top=1in, left=1.2in,right=1.2in]{geometry}
\longindentation=0pt

%% --- Logo
\usepackage{graphicx}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{IEEE Transactions on Cognitive and Developmental Systems\newline
Special Issue on Sensorimotor Contingencies for Cognitive Robotics}
\includegraphics[height=3cm, keepaspectratio=true]{figures/unam.jpg}\vspace*{-3cm}
\address{Facultad de Ciencias, \\Ciudad Universitaria,\\ UNAM, México, D.F.}

\opening{Dear Dr. Ricardo Téllez,}

Please find attached our submission:
\begin{center}
 ``A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation'' \\
 Veronica E.~Arriola-Rios (UNAM, Mexico), Jeremy~L.~Wyatt (University of Birmingham, UK)
\end{center}

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author

This article aims at tracking three-dimensional deformable objects under robot motion and predicting the force causing the deformation.

\comment{There is an omission here: the article aims at predicting the force and deformation on the visual plane of the three-dimensional object.  The tracking of the shape is used only during training and for evaluation purposes.}

For these purposes, the authors propose two models, including an extension of the mass-spring model and a regression model, where the parameters of the mass-spring model are learned with an evolutionary searching algorithm on visual data and force data, and the parameters of the regression model are learned with least-square methods on stress-strain graph. These models are tested on a sponge and a plasticine, and evaluated based on the balanced F-score. The results are presented both quantitatively and qualitatively.

However, there are several structural problems, as listed below.

1: in the abstract, the authors claim that five features are simultaneously enabled by the proposed single learning model framework, but the article actually presents a collection of different models and methods to solve different problems, thus it is inappropriate to claim them as a single unified model framework. Besides, it seems that only robot pushing is modeled, not general "robot actions".

2: section II is not included in the final paragraph of section I, and this section should be included in section I or section IV.

\comment{It was included in the final paragraph of section I.  It was not included in section I to keep the general idea of the model isolated from other aspects in the introduction.}

In addition, this section is really difficult to follow. For example, in line 27 of page 1, “the specific approach taken here separates the predictive model out into components in two ways". It will be easier to read if the authors clearly indicate the second way in line 44. And the third paragraph, on model training, should come before the second graph, which is on model testing.

3: section IV-C, "Integration Scheme", seems to be part of algorithm 1 or section IV-B, thus it should be introduced.

\comment{It is neither part of algorithm 1, nor of section IV-B.  It is part of a numerical algorithm for solving differential equations. \newline TODO: Add general algorithm which makes use of all the pieces explained in section IV at the beginning of the section.}

4, section IV-D is not explained: what is the relation between "collision detection" and the proposed model?

5, section IV-E, "Geometric Constraints", should be integrated into section IV-A, where the energy function is built on different geometric constraints.

\comment{TODO: Move section IV-E after IV-A but must emphasize that the geometric constraints in IV-A are enforced through energy preservation, while those in IV-E have immediate effects on the triangulations without directly affecting the equations.}

6, section IV-F, "Many Step Prediction", explains how to recursively train and test the model. It should be included in section V, where the model training is given.

\comment{There is a missunderstanding here.  Section IV-F is independent of any training.  As a physics based model, it depends only on proper entries and parameters, which could be provided by a human expert after manual calibration or by a computer algorithm.  The last paragraph of this section was modified accordingly and the algorithm introduced for comment 3 should help here.}

7, section V-A explains how to search for the optimal parameter with evolutionary algorithm based on the ground truth; section V-B explains how to get the ground truth; section V-C explains how to test the learned parameter. It seems that the correct order is B-A-C.

\comment{Changed order as suggested.}

Besides, there are several technical ambiguities in the manuscript, as listed below.

1, in line 37 of page 1, "sensorimotor contingencies" needs to be defined or referenced.

2, in line 35 of page 1, what does "sequencing the models by mode" mean?

3, in line 58 of page 1, the authors claim "In each case the models can be run in two modes, prediction and classification". But the classification mode is not explained later, but a filtering mode is included instead. Besides, please clearly illustrate the mentioned modes and processes in Figure 2.

4, in section IV-A, the author needs to explain how they extend the model proposed in [16].

5, in section IV-A, the author needs to briefly introduce the mass-spring model.

\comment{Added an introductory paragraph to section IV.}

6, in equation (1), please explain what "particle i" is. Is it vertex i?

7, in line 27 of page 4, please explain what "mass particle" is.

8, in line 33 of page 4, please explain why the equations are oscillatory.

9, please explain how equation (3) is derived.

10, in line 54 of page 4, please add reference to "Teschner normalised the value dividing by D0 to make the elasticity constants scale independent."

11, in line 13, please explain "Since Teschner and Morris didn’t find it helpful to add damping to the preservation of areas, it is avoided here".

12, in line 50 of page 4, please explain "the angles between adjacent edges". As adjacent edges form one angle, do the authors mean adjacent angles? Similar ambiguity exists in line 2 of page 5.

13, in Algorithm 1, please explain how "an integration step" leads to new length of the edge.

14, in line 25 of page 4, please explain what "Euler of Verlet" means. Euler or Verlet?

15, in equation (17), please explain if $x(t+h)$ is a position, how does $h*\frac{F(t)}{m}$ generate a position vector, since $h*\frac{F(t)}{m}$ seems to be the production of $h$ and acceleration? And please explain "we made the velocities proportional to the forces". Which variable is the velocity? What does "m" mean? The mass?

16, in section IV-D, please explain what "collision detection" means, the collision between robot and objects, or the collision between objects?

17, is algorithm 2 designed by the authors?

18, in section V-A, please explain why evolutionary algorithm is used. In algorithm 4, please explain what $\alpha, \beta,\gamma,\zeta$ represent.

19, in line 60 of page 7, please explain how the object is segmented from the environment, which "simple colour segmentation is used", and how to define "edges that not likely to belong to the object of interest".

20, in line 23 of page 7, please explain what "linear snake" indicates. Is it from active contour model? Also, please add reference.

21, in section V-C, please note the measurement has a name, "F-score", in machine learning community. Also, please define $\mu(f)$ based on "the mean over all the frames so far".

22, in line 30 of page 8, please explain how the stress-strain diagram is obtained, and how the training data set is collected.

23, in the experiments, please compare with other existing methods.

24, in Table II, please explain the meaning of the values. Are they mean values and standard deviations? If so, what does "best, worst, average" mean?

25, in Figure 12, please clarify the legends. For example, why are there two lines with the same marker?

26, in line 56 of page 10, please explain "integration step". Is it the "h" in equation 17?

27, in section VII-E, please explain how the model is used in object classification. For example, what does "even for novel interactions" mean? Object classification typically tries to identify an object in a new environment. Besides, how is "Global" criteria evaluated in object classification?

28, in Table III, what does "test 1, test 2" mean? Please explain how these numbers are computed. Besides, object classification is typically evaluated with precision-recall curve.

29, please include more related works, since there are tons of articles on deformable object tracking.

Associate Editor's Comments to Author:

Associate Editor 
Comments to the Author: 
(There are no comments. Please check to see if comments were included as a file attachment with this e-mail or as an attachment in your Author Center.)

I look forward to hearing from you.

\signature{Dra. Verónica Esther Arriola Ríos\\
Profesora Asociada C de T.C.\newline
Departamento de Matemáticas, Cub 119.\newline
Facultad de Ciencias, UNAM \newline
v.arriola@ciencias.unam.mx \newline
+(52)55 5622 5426}

\closing{Yours sincerely,}

%enclosure listing
%\encl{}

\end{letter}
\end{document}

### 2016-08-01 twostagedo.tex
Balanced columns in last page.
Addressed first comments, added AswersToComments

@@
 % ++++ Used temporarily to make comments between authors
-\newcommand{\redNote}[1]{{\color{red} #1}}
+\newcommand{\comment}[1]{{\color{red} #1}}
 \usepackage[normalem]{ulem}
+\newcommand{\commentErrase}[1]{\comment{\sout{#1}}}
 
@@
 Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
-In the rest of the paper we present related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).
+In the rest of the paper we present \comment{an overview of the model in more detail (Section~\ref{sec:model_overview});} related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).
 
@@
 \section{Model Overview}
-
-As has been mentioned above there are several choices made in our modelling approach.
+\label{sec:model_overview}
+
+As has been mentioned above there are several choices made in our modeling approach.
 The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour.
@@
 
-Additionally, there is a vast amount of literature on modelling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.
+Additionally, there is a vast amount of literature on modeling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.
 However, these types of models have not been successfully applied to robotic tasks beyond what was presented above.
@@
 \section{A Mass-Spring Model of Deformation}
+
+\comment{This work makes use of the ability of a mass-spring model to predict the shape of a deformable object given its shape at time $t$ and the forces acting on it.
+The parameters, that attune this generic model to a particular object, will be obtained through machine learning, and that procedure will be explained in \sref{sec:learning}, since it is independent of the model itself.
+Also the forces required as inputs by this mass-spring model will be learned by another module, so that the entire system will require only the initial shape of the object, and the command to be executed, to be able to predict the future behaviour of the object and sensed feedback force.
+This section explains the details of the mass-spring model we used.
+
+Mass-spring models are a very well established method to approximately simulate the behaviour of solid deformable objects.
+Here, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).
+At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located.
+The edges between vertices roughly model the interactions between particles that attract each other and at the same time repel each other, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.
+The simplest model associates a linear spring with each edge, as if those forces would occur only between connected neighbours.
+The resting state of the solid corresponds to a configuration of the masses and particles where the sum of all forces acting over each particle is zero and they are said to be in equilibrium.
+
+When one or more particles are displaced from their positions of equilibrium, the forces that tend to restore the body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).
+These forces act upon the particles modifying their positions in time.
+Traditionally, the relationship between the position of the particle $p_i$ and those forces $F$ obeys the second law of Newton:
+\begin{align}
+ F = m \dfrac{\partial^2 p}{{\partial t}^2}
+\end{align}
+where $t$ is time.  However, it is possible to modify this equation according to need if what we are interested in is numerical accuracy, as it will be done here.
+
+Furthermore, the previous equation is a second order differential equation, which is continuous by definition.
+In order to obtain a computer simulation from this equation some numerical method will be selected \cite{Nealen2006review}.
+Time is discretized in intervals of length $h$, positions and forces at discrete past times are used to estimate the new positions at time $t + h$ according to an integration schemed use to approximately solve the differential equation.
+The fact that the value of the forces remain constant for the whole interval $h$ (also know as \textit{time step}), introduces errors which can be severe, depending on their magnitude and the speed at which the force would change during that interval according to the original continuous equation.
+For this reason, computer simulations add procedures that pretend to lessen these effects.
+
+Additionally, the mass spring model of the solid object does not specify how the rest of the world interacts with it.
+Usually, any external force is introduced as a vector of forces acting on a vertex or as a deformation of the mesh, which breaks the state of equilibrium, thus causing a reaction of the object due to internal forces produced by the springs.
+Therefore, to simulate the effects of interacting with the object of interest the model must include the definition of the interactions with surrounding objects (like the effector of the robot).
+Those details are shown below.
+}
+
 \label{sec:model}
@@
 
-The next stage of the learning process consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.
+The next stage \commentErrase{of} \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.
+This problem is addressed in the following section.
 
@@
 \label{sec:learning}
+
+\subsection{Tracking to Provide Ground Truth}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=88mm]{arrio2}
+\caption{(a) Experimental set up.
+A robotic finger pushes a block of deformable material over a table.
+A camera registers the events of the plane where the main deformations take place.
+(b) Evaluation function.
+Left top: Mesh over real block of material.
+Left bottom: Tipical error of the model vs the ground truth (red) and tipical error of the tracker (blue).
+Right top: tracked ground data.
+Middle: TP in yellow, FP in blue, FN in red and TN in gray.
+Bottom: simulated mesh.
+}
+\label{fig:scene}
+\end{figure}
+
+The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
+Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline}.
+The tracking of the deformable target in the ground truth has two stages:
+\begin{enumerate}
+ \item The target object is segmented from the environment.
+The Canny algorithm is used to extract the edges of the image.
+Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.
+ \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
+\end{enumerate}
+For the first frame, the snake is initialised as a rectangle around the target.
+This rectangle fragments itself into smaller pieces.
+Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].
+If there is no edge within that area, the control point remains were it is.
+This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.
+Once the edge reappears, the control point adapts itself again.
+Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.
+See \alref{alg:regularise}.
+This allows the snake to adapt very quickly to the complexity of the contour it represents.
+
+\begin{figure}
+ \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
+ \STATE Be $L$ a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
+ \FORALL {$L_i$ in $L$}
+ \IF {$dist(L_i,L_{i-1}) > max\_distance$}
+  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
+ \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
+  \STATE erase $L_i$
+ \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
+  \STATE erase $L_i$
+ \ENDIF
+ \ENDFOR
+\end{algorithmicieee}
+\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
+\end{figure}
+
 \subsection{Evolutionary Algorithm}
@@
 
-\subsection{Tracking to Provide Ground Truth}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=88mm]{arrio2}
-\caption{(a) Experimental set up.
-A robotic finger pushes a block of deformable material over a table.
-A camera registers the events of the plane where the main deformations take place.
-(b) Evaluation function.
-Left top: Mesh over real block of material.
-Left bottom: Tipical error of the model vs the ground truth (red) and tipical error of the tracker (blue).
-Right top: tracked ground data.
-Middle: TP in yellow, FP in blue, FN in red and TN in gray.
-Bottom: simulated mesh.
-}
-\label{fig:scene}
-\end{figure}
-
-The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
-Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline}.
-The tracking of the deformable target in the ground truth has two stages:
-\begin{enumerate}
- \item The target object is segmented from the environment.
-The Canny algorithm is used to extract the edges of the image.
-Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.
- \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
-\end{enumerate}
-For the first frame, the snake is initialised as a rectangle around the target.
-This rectangle fragments itself into smaller pieces.
-Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].
-If there is no edge within that area, the control point remains were it is.
-This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.
-Once the edge reappears, the control point adapts itself again.
-Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.
-See \alref{alg:regularise}.
-This allows the snake to adapt very quickly to the complexity of the contour it represents.
-
-\begin{figure}
- \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
- \STATE Be $L$ a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
- \FORALL {$L_i$ in $L$}
- \IF {$dist(L_i,L_{i-1}) > max\_distance$}
-  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
- \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
-  \STATE erase $L_i$
- \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
-  \STATE erase $L_i$
- \ENDIF
- \ENDFOR
-\end{algorithmicieee}
-\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
-\end{figure}
-
 \subsection{Evaluation Function}
@@
 % the document is modified later
-%\IEEEtriggeratref{8}
+%\IEEEtriggeratref{15}
 % The "triggered" command can be changed if desired:
@@
 
-%\vfill
+\vfill
 

### 2016-08-11 AnswersToComments.tex
Intro to mass-spring added.

@@
 \usepackage{color}
-\newcommand{\comment}[1]{\textcolor{red}{#1}}
+\newcommand{\comment}[1]{\textcolor{blue}{#1}}
+\newcommand{\marker}{\hspace*{-1.6em}\textcolor{red}{$\Longrightarrow$}}
 
@@
 
+\marker
 1, in line 37 of page 1, "sensorimotor contingencies" needs to be defined or referenced.
 
+\marker
 2, in line 35 of page 1, what does "sequencing the models by mode" mean?
 
+\marker
 3, in line 58 of page 1, the authors claim "In each case the models can be run in two modes, prediction and classification".
@@
 
+\comment{What does he mean by ``filtering mode instead''?}
+
 4, in section IV-A, the author needs to explain how they extend the model proposed in [16].
+
+\comment{The general idea had been explained in the first paragraph of section IV-A, the details came later.  Since it was not clear enough the wording was modified.}
 
@@
 
-\comment{Added an introductory paragraph to section IV.}
+\comment{Added an introduction at the beginning of section IV.}
 
 6, in equation (1), please explain what "particle i" is. Is it vertex i?
+
+\comment{Since for comment 5 the introduction explained that a mass particle is located on vertex i, this has been covered already.}
 

### 2016-08-11 twostagedo.tex
Intro to mass-spring added.

@@
 
-\comment{This work makes use of the ability of a mass-spring model to predict the shape of a deformable object given its shape at time $t$ and the forces acting on it.
+\comment{
+A mass-spring model is a very well established abstraction used to emulate the behaviour of deformable objects.
+As a physics based model, once its parameters are properly assigned, its behaviour is determined by the evolution of a set of differential equations.
+Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the interacting forces are know.
+The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model for the particular object.
+For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.
+In this work we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration to a system that exploits it to solve prediction and classification tasks.
+We emphasize that the physics based model could be easily substituted by another equivalent model.
+For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.
+For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.
+
+%This work makes use of the ability of a mass-spring model to predict the shape of a deformable object given its shape at time $t$ and the forces acting on it.
 The parameters, that attune this generic model to a particular object, will be obtained through machine learning, and that procedure will be explained in \sref{sec:learning}, since it is independent of the model itself.
@@
 
-Mass-spring models are a very well established method to approximately simulate the behaviour of solid deformable objects.
-Here, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).
+In a simple mass-spring model, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).
 At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located.
@@
 
-When one or more particles are displaced from their positions of equilibrium, the forces that tend to restore the body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).
+When one or more particles are displaced from their positions of equilibrium, the forces that tend to restore an elastic body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).
 These forces act upon the particles modifying their positions in time.
@@
 \begin{align}
- F = m \dfrac{\partial^2 p}{{\partial t}^2}
+ F = m \dfrac{\partial^2 p}{{\partial t}^2} \label{eq:newton}
 \end{align}
-where $t$ is time.  However, it is possible to modify this equation according to need if what we are interested in is numerical accuracy, as it will be done here.
-
-Furthermore, the previous equation is a second order differential equation, which is continuous by definition.
-In order to obtain a computer simulation from this equation some numerical method will be selected \cite{Nealen2006review}.
-Time is discretized in intervals of length $h$, positions and forces at discrete past times are used to estimate the new positions at time $t + h$ according to an integration schemed use to approximately solve the differential equation.
-The fact that the value of the forces remain constant for the whole interval $h$ (also know as \textit{time step}), introduces errors which can be severe, depending on their magnitude and the speed at which the force would change during that interval according to the original continuous equation.
-For this reason, computer simulations add procedures that pretend to lessen these effects.
-
-Additionally, the mass spring model of the solid object does not specify how the rest of the world interacts with it.
-Usually, any external force is introduced as a vector of forces acting on a vertex or as a deformation of the mesh, which breaks the state of equilibrium, thus causing a reaction of the object due to internal forces produced by the springs.
-Therefore, to simulate the effects of interacting with the object of interest the model must include the definition of the interactions with surrounding objects (like the effector of the robot).
-Those details are shown below.
+where $t$ is time.
+However, it is possible to modify this equation according to need if what we need is to model a different behaviour.
+%are interested in is numerical accuracy, as it will be done here.
+
+The previous equation is a second order differential equation, which is continuous by definition.
+In order to perform a computer simulation we need to obtain the position $p$ of all vertices as a function of time $p(t)$ using this equation.
+A numerical method is selected to approximate a solution \cite{Nealen2006review}.
+Time is discretized in intervals of length $h$, positions and forces at discrete past times are used to estimate the new positions at time $t + h$, according to an integration scheme which must be selected.
+The fact that the value of the forces remain constant in this approximation, for the whole interval $h$ (also know as \textit{time step}), introduces errors which can be severe.
+For this reason, computer simulations may add procedures that help to alleviate these effects.
+
+To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
+\begin{align}
+ F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
+\end{align}
+Another option could be to induce response forces by displacing vertices from their position of equilibrium.
 }
@@
 Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.
-The potential energy terms for the preservation of length with damping and preservation of area were included in our system.
-To compensate for the lack of a preservation of volume, which tends to avoid distortions of the graph, a new term for the preservation of the angles was included.
+The potential energy terms for the preservation of length with damping and preservation of area were included in our system \comment{exactly as they were defined by Teschner}.
+To compensate for the lack of a preservation of volume, which tends to avoid distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.
 The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.
 
-For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.
+\comment{Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.
+He makes use of different constraints to model the tendency of an object to recover several of its attributes.} For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.
 When the value of the constraint is zero, the potential energy is zero as well.
@@
 \end{align}
-where $k$ is a proportionality constant \eref{eq:EConstraint} and $p_i$ is the position of particle $i$.
+where $k$ is a proportionality constant \sout{\eref{eq:EConstraint}} and $p_i$ is the position of particle $i$.
 

### 2016-08-23 AnswersToComments.tex
Reviewed until point 15.

@@
 
+\marker
 1: in the abstract, the authors claim that five features are simultaneously enabled by the proposed single learning model framework, but the article actually presents a collection of different models and methods to solve different problems, thus it is inappropriate to claim them as a single unified model framework.
@@
 
+\comment{The term was introduced on the second paragraph of section IV.}
+
 8, in line 33 of page 4, please explain why the equations are oscillatory.
+
+\comment{Added explanation that it happens when/because we use $C$s that model springs.}
 
@@
 
+\comment{It is not within the scope of the article to explain that derivation, since it belongs to Teschner's model, however the cite to the article we got it from was repeated here.}
+
 10, in line 54 of page 4, please add reference to "Teschner normalised the value dividing by D0 to make the elasticity constants scale independent."
+
+\comment{Added.}
 
@@
 
+\comment{Changed phrase for the explanation that Teschner gave.}
+
 12, in line 50 of page 4, please explain "the angles between adjacent edges". As adjacent edges form one angle, do the authors mean adjacent angles? Similar ambiguity exists in line 2 of page 5.
+
+\comment{Corrected.}
 
@@
 
+\comment{Reworded sentence to add more detail.}
+
 14, in line 25 of page 4, please explain what "Euler of Verlet" means. Euler or Verlet?
+
+\comment{It was ``or''.  Corrected.}
 
@@
 The mass?
+
+\comment{Explained in more detail and changed notations to make it consistent with previous paragraphs.}
 

### 2016-08-23 twostagedo.tex
Reviewed until point 15.

@@
 \comment{
-A mass-spring model is a very well established abstraction used to emulate the behaviour of deformable objects.
+A mass-spring model is a very well established abstraction used frequently to simulate the behaviour of some deformable objects.
 As a physics based model, once its parameters are properly assigned, its behaviour is determined by the evolution of a set of differential equations.
@@
 In a simple mass-spring model, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).
-At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located.
-The edges between vertices roughly model the interactions between particles that attract each other and at the same time repel each other, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.
+At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located, and is known as a \textit{mass particle}.
+The edges between vertices roughly model the interactions between particles, that attract each other up to a certain critical distance, but repel each other if they become closer, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.
 The simplest model associates a linear spring with each edge, as if those forces would occur only between connected neighbours.
-The resting state of the solid corresponds to a configuration of the masses and particles where the sum of all forces acting over each particle is zero and they are said to be in equilibrium.
+The resting state of the solid corresponds to a configuration of the masses and springs where the sum of all forces acting over each particle is zero, and they are said to be in equilibrium.
 
@@
 
-The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004} was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.
+The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.
 The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.
@@
 \end{align}
-where $k$ is a proportionality constant \sout{\eref{eq:EConstraint}} and $p_i$ is the position of particle $i$.
+where $k$ is a proportionality constant \comment{\sout{\eref{eq:EConstraint}}} and $p_i$ is the position of particle $i$.
 
@@
 \end{align}
-Given the oscilatory nature of these equations, damping can be introduced to stop the oscillations.
-This is modelled with a force acting in the opposite direction to the velocity of the affected particles:
+Given the oscilatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.
+This is modelled with a force acting in the opposite direction to the velocity of the affected particles\comment{, thus slowing down motion.
+Teschner wrote this as a function of the constraints as well \cite{Teschner2004}}:
 \begin{align}
@@
 The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].
-Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.
+Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent \cite{Teschner2004}.
 Morris \cite{Morris2008} omitted that in his final equations for the force.
@@
 This constraint is applied per triangle in the mesh.
-Since Teschner and Morris didn't find it helpful to add damping to the preservation of areas, it is avoided here.
-Therefore the force $\vec{F}^i_A$ is derived as follows:
+Since Teschner and Morris \comment{\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} found that the use of damping for the preservation of areas does not improve significantly the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
 \begin{align}
@@
 [\fref{fig:forces}(c)] gives an idea of how these forces look.
-Here the energy depends on the difference of the angles between adjacent edges.
+Here the energy depends on the difference \comment{\sout{of the angles between adjacent edges.} between the current angle, between adjacent edges, and the angle between them at the equilibrium position.}
 
@@
  \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
- \STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
+ %\STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
+ \comment{\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.} And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
  \IF {$\alpha > max\_\alpha$}
@@
 \label{sec:ischeme}
-Rather than using the traditional Newton equations with a numeric integration scheme like Euler of Verlet, we made the velocities proportional to the forces:
+Rather than using the traditional Newton equations with a numeric integration scheme like Euler \comment{or} Verlet, \comment{used twice to solve the second order differential equiations,} we made the \comment{\sout{velocities proportional to the forces:} force proportional to the velocity:
+\begin{align}
+ F(t) = m \dfrac{\partial \vec{p}}{\partial t}
+\end{align}
+and applied Euler only once to solve for $p$:
 \begin{equation}
- \vec{x}(t+h) = \vec{x}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
-\end{equation}
-with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{x}(t+h)$, the position of the vertex at time $t+h$.
-The simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
-When experimenting with different integration steps, force values were interpolated between measuremente.
+ \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
+\end{equation}}
+with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.
+\comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
+When experimenting with different integration steps, force values were interpolated between measurements.
 

### 2016-08-25 AnswersToComments.tex
Corrected collision detection

@@
 
-\comment{There is an omission here: the article aims at predicting the force and deformation on the visual plane of the three-dimensional object.
-The tracking of the shape is used only during training and for evaluation purposes.}
+\comment{It is important to make a precision here: the article aims at predicting the force and deformation on the visual plane of the three-dimensional object.
+The tracking of the shape is used only during training as part of the evaluation function.}
 
@@
 
+\comment{Improved the description across the whole subsection.}
+
 17, is algorithm 2 designed by the authors?
 
-18, in section V-A, please explain why evolutionary algorithm is used. In algorithm 4, please explain what $\alpha, \beta,\gamma,\zeta$ represent.
+\comment{No, that is why we mentioned in the text that it is a common algorithm.  It can be found across the web in many sites.  Should we mention something about that in the caption?}
+
+18, in section V-A, please explain why evolutionary algorithm is used. In algorithm 4, please explain what $\alpha, \beta, \gamma, \zeta$ represent.
 
@@
 
+\comment{We don't focus on tracking, we want to predict several frames ahead, not follow, that is why closely related works are few.}
+
 Associate Editor's Comments to Author:

### 2016-08-25 twostagedo.tex
Corrected collision detection

@@
 \subsection{Collision Detection}
-In the scenario, rigid objects are treated as geometric obstacles.
-Two types of obstacles are considered: circles and a floor line that can not be crossed.
-For the floor obstacle, the collision is roughly solved by pushing vertices back to the border of the floor.
+In the scenario, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.
+\comment{When the simulation of the behaviour of the mesh, that models the image of the sponge, is carried on, the predicted positions of the vertices may indicate that the mesh penetrates the obstacles.
+This should be interpreted as a collision between the sponge and the obstacle.
+Since the sponge can not penetrate an obstacle, the overlapping between geometries must be solved, and its physical consequences must affect the simulation.
+For example: if vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
+
+In order to model this,} two types of \comment{solid} obstacles are considered: circles \comment{(for the finger)} and a table line that can not be crossed.
+For the table obstacle, the collision \comment{of parts of the sponge with it} is roughly solved by pushing vertices \comment{of the mesh of the sponge} back to the border of the table.
 The springs will eventually propagate the effect to the neighbours.

### 2016-08-25 twostagedo.tex
Started working on learninig section.

@@
 \subsection{Evolutionary Algorithm}
+\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they depend on the geometry of object, and there can be several sets of parameters in different regions of the space of parameters that produce adequate simulations \cite{Morris2008}, we chose to use an evolutionary algorithm to expore different regions and try improvements in the vecinity of successful individuals.}
+
 The evolutionary algorithm listed in \alref{alg:genetic} was used to search the space of parameters of the mass-spring model presented in the previous section.

### 2016-08-25 twostagedo.tex
edited abstract, title and introduction.

--- 2016-08-23 twostagedo.tex
@@
 \markboth{IEEE CDS}%
-{Arriola-Rios \MakeLowercase{\textit{et al.}}: A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
+{Arriola-Rios \MakeLowercase{\textit{et al.}}: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
 % The only time the second header will appear is for the odd numbered pages
@@
 \begin{abstract}
-In this paper we present a new framework for learning generative models of object deformation under robotic action.
+\comment{In this paper we present a modular, multi-modal framework for learning generative models of object deformation under robotic pushing.
 The model is multi-modal in that it is based on integrating force and visual information.
-Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
-While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning model framework, and none has offered classification from a generative predictable model.
+The framework is also modular, being composed of a learned module for each type of material.
+Each such module is learned by parameter estimation for a general model form.
+When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of the modules are compared so as to classify the material.} Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
+While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning framework, and none has offered classification from a generative model.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Deformable objects, prediction, classification, learning.
+deformable objects, prediction, classification, learning.
 \end{IEEEkeywords}
@@
 \IEEEPARstart{A} major challenge in robotic manipulation is to plan actions with objects so as to deform them into new shapes.
-Most objects that are manipulated by animals in natural environments are deformable, and these include objects that are elastic, plastic, breakable, tearable, and spreadable.
-This work is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required.
-These sensorimotor contingencies can then be used to feedback on themselves to predict over long timescales.
-Without such models planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.
-
-There is a wide range of models of deformation used in engineering, that are capable of producing predictions, such as finite element models, but these typically require careful optimisation by hand over many weeks or months.
-In graphics there are many models used for generating qualitatively plausible simulations of objects, but these are not necessarily easy to tune to real data.
-By contrast robots must perform system identification of their model parameters from data, on-line, in reasonable amounts of training time.
-This includes building models from the usually very limited sensing typically available, such as partial visual views of an object, and force-torque monitoring at contact points.
-Thus tackling this problem in robotics adds additional difficulties to an already challenging problem.
+Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable.
+\comment{The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required.
+A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations.
+Thus different materials (elastic, plastic) present have different sensorimotor contingencies.
+The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data.
+These learned models can then be used to make the predictions required for planning.
+The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales.
+Without such predictions planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.
+
+Various models of deformation are used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
+In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data.
+In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time.
+This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points.
+Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.}
 
@@
 
-Previous work on learning sensorimotor contingencies for deformable objects in robotics has been suprisingly limited.
+\comment{Previous work on learning sensorimotor contingencies for deformable objects in robotics has been limited.
 Table~\ref{t:advances} summarises the features of the major efforts to date.
-In each case critical abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g.
+In each case some abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g.
 elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?).
 In this paper we present a framework for solving all these problems simultaneously.
-The key novel contribution we make is to bring together a model of what forces the robot will feel at a contact point with a model of how an object will deform under such applied forces over time.
-This formulation allows the robot to make predictions about the type of object, and thus it's future behaviour using the resistive forces and the object deformations \fref{fig:forceElasticPlastic} and \fref{fig:sstrain}.
-A second critical contribution is that our model is both generative and generalisable with respect to contact point.
-Finally the third major feature of the model is that it is adaptive: a set of model parameters is learned for each material presented from data, thus creating a model library of materials.
-This generative property enables both prediction suitable for use in planning, and also classification of materials.
-The overall model structure is summarised in \fref{fig:system}.
-Essentially the approach is to use visual tracking of shape deformations to create feedback that is used to train the parameters of a model (Model A) that predicts deformation when coupled with contact forces.
-The contact forces are themselves predicted using a second model (Model B).
-In the specific implementation of our system we use a mass spring system for Model A, and a regression model for Model B.
-These models are learned from the same data, and are both indexable by a known material type from the library.
-It is important to note, however, that the model structure is separate from the specific model forms used, and that other model types could be substituted for one or both stages.
+
+There are several novel contributions.
+First, we show how the modular learning framework of \cite{harunowolpert,kopickiwyatt16} can be applied to deformable objects.
+Second, we show how each module can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces.
+Both models are generative, and we show how to learn each from data.
+Third, we show how to use these learned generative models not only to predict deformation, but also to classify an unidentified material.}
+
  %This behaviour is described in \alref{alg:overall}.
@@
 
-In summary the two stage model we present gives us the ability to solve {\em simultaneously} for the first time the following problems.
+In summary the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems.
 Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
@@
 \comment{
-A mass-spring model is a very well established abstraction used frequently to simulate the behaviour of some deformable objects.
+A mass-spring model is a very well established abstraction used to emulate the behaviour of deformable objects.
 As a physics based model, once its parameters are properly assigned, its behaviour is determined by the evolution of a set of differential equations.
@@
 In a simple mass-spring model, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).
-At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located, and is known as a \textit{mass particle}.
-The edges between vertices roughly model the interactions between particles, that attract each other up to a certain critical distance, but repel each other if they become closer, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.
+At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located.
+The edges between vertices roughly model the interactions between particles that attract each other and at the same time repel each other, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.
 The simplest model associates a linear spring with each edge, as if those forces would occur only between connected neighbours.
-The resting state of the solid corresponds to a configuration of the masses and springs where the sum of all forces acting over each particle is zero, and they are said to be in equilibrium.
+The resting state of the solid corresponds to a configuration of the masses and particles where the sum of all forces acting over each particle is zero and they are said to be in equilibrium.
 
@@
 
-The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.
+The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004} was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.
 The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.
@@
 \end{align}
-where $k$ is a proportionality constant \comment{\sout{\eref{eq:EConstraint}}} and $p_i$ is the position of particle $i$.
+where $k$ is a proportionality constant \sout{\eref{eq:EConstraint}} and $p_i$ is the position of particle $i$.
 
@@
 \end{align}
-Given the oscilatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.
-This is modelled with a force acting in the opposite direction to the velocity of the affected particles\comment{, thus slowing down motion.
-Teschner wrote this as a function of the constraints as well \cite{Teschner2004}}:
+Given the oscilatory nature of these equations, damping can be introduced to stop the oscillations.
+This is modelled with a force acting in the opposite direction to the velocity of the affected particles:
 \begin{align}
@@
 The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].
-Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent \cite{Teschner2004}.
+Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.
 Morris \cite{Morris2008} omitted that in his final equations for the force.
@@
 This constraint is applied per triangle in the mesh.
-Since Teschner and Morris \comment{\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} found that the use of damping for the preservation of areas does not improve significantly the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
+Since Teschner and Morris didn't find it helpful to add damping to the preservation of areas, it is avoided here.
+Therefore the force $\vec{F}^i_A$ is derived as follows:
 \begin{align}
@@
 [\fref{fig:forces}(c)] gives an idea of how these forces look.
-Here the energy depends on the difference \comment{\sout{of the angles between adjacent edges.} between the current angle, between adjacent edges, and the angle between them at the equilibrium position.}
+Here the energy depends on the difference of the angles between adjacent edges.
 
@@
  \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
- %\STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
- \comment{\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.} And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
+ \STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
  \IF {$\alpha > max\_\alpha$}
@@
 \label{sec:ischeme}
-Rather than using the traditional Newton equations with a numeric integration scheme like Euler \comment{or} Verlet, \comment{used twice to solve the second order differential equiations,} we made the \comment{\sout{velocities proportional to the forces:} force proportional to the velocity:
-\begin{align}
- F(t) = m \dfrac{\partial \vec{p}}{\partial t}
-\end{align}
-and applied Euler only once to solve for $p$:
+Rather than using the traditional Newton equations with a numeric integration scheme like Euler of Verlet, we made the velocities proportional to the forces:
 \begin{equation}
- \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
-\end{equation}}
-with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.
-\comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
-When experimenting with different integration steps, force values were interpolated between measurements.
+ \vec{x}(t+h) = \vec{x}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
+\end{equation}
+with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{x}(t+h)$, the position of the vertex at time $t+h$.
+The simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
+When experimenting with different integration steps, force values were interpolated between measuremente.
 

### 2016-08-25 twostagedo.tex

@@
 \label{sec:ischeme}
-Rather than using the traditional Newton equations with a numeric integration scheme like Euler of Verlet, we made the velocities proportional to the forces:
+Rather than using the traditional Newton equations with a numeric integration scheme like Euler \comment{or} Verlet, \comment{used twice to solve the second order differential equiations,} we made the \comment{\sout{velocities proportional to the forces:} force proportional to the velocity:
+\begin{align}
+ F(t) = m \dfrac{\partial \vec{p}}{\partial t}
+\end{align}
+and applied Euler only once to solve for $p$:
 \begin{equation}
- \vec{x}(t+h) = \vec{x}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
-\end{equation}
-with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{x}(t+h)$, the position of the vertex at time $t+h$.
-The simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
-When experimenting with different integration steps, force values were interpolated between measuremente.
+ \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
+\end{equation}}
+with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.
+\comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
+When experimenting with different integration steps, force values were interpolated between measurements.
 
 \subsection{Collision Detection}
-In the scenario, rigid objects are treated as geometric obstacles.
-Two types of obstacles are considered: circles and a floor line that can not be crossed.
-For the floor obstacle, the collision is roughly solved by pushing vertices back to the border of the floor.
+In the scenario, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.
+\comment{When the simulation of the behaviour of the mesh, that models the image of the sponge, is carried on, the predicted positions of the vertices may indicate that the mesh penetrates the obstacles.
+This should be interpreted as a collision between the sponge and the obstacle.
+Since the sponge can not penetrate an obstacle, the overlapping between geometries must be solved, and its physical consequences must affect the simulation.
+For example: if vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
+
+In order to model this,} two types of \comment{solid} obstacles are considered: circles \comment{(for the finger)} and a table line that can not be crossed.
+For the table obstacle, the collision \comment{of parts of the sponge with it} is roughly solved by pushing vertices \comment{of the mesh of the sponge} back to the border of the table.
 The springs will eventually propagate the effect to the neighbours.

### 2016-08-26 twostagedo.tex
changes to title etc

@@
 % Do not put math or special symbols in the title.
-\title{A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
+% could we change the title to this? What do you think? I wanted to get the modular idea in there.
+\title{Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects
+%A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
+}
 %
@@
 \author{Veronica E.~Arriola-Rios,
-        Jeremy~L.~Wyatt}
-%\author{Veronica E.~Arriola-Rios,~\IEEEmembership{Member,~IEEE,}
-%        John~Doe,~\IEEEmembership{Fellow,~OSA,}
-%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
-%\thanks{M. Shell is with the Department
-%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
-%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
-%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
+        Jeremy~L.~Wyatt \thanks{V.E. Arriola-Rios is with Facultad de Ciencias, UNAM, Mexico, e-mail: v.arriola@ciencias.unam.mx}
+        \thanks{J.L. Wyatt is with the University of Birmingham, Edgbaston, Birmingham, B15 2TT, UK. e-mail: jeremy.l.wyatt@gmail.com }}%
+% <-this % stops a space
+%\thanks{J.L. Wyatt is with the University of Birmingham, Edgbaston, Birmingham, B15 2TT, UK. e-mail: jeremy.l.wyatt@gmail.com }% <-this % stops a space
 %\thanks{Manuscript received April 19, 2005; revised December 27, 2012.}}
@@
 \markboth{IEEE CDS}%
-{Arriola-Rios \MakeLowercase{\textit{et al.}}: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
+{Arriola-Rios and Wyatt: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
 % The only time the second header will appear is for the odd numbered pages
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{A} major challenge in robotic manipulation is to plan actions with objects so as to deform them into new shapes.
+\IEEEPARstart{A} major challenge in robot manipulation is to plan actions with objects so as to deform them into new shapes.
 Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable.
@@
 A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations.
-Thus different materials (elastic, plastic) present have different sensorimotor contingencies.
+Different materials (elastic, plastic) have different sensorimotor contingencies.
 The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data.
-These learned models can then be used to make the predictions required for planning.
 The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales.
-Without such predictions planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.
-
-Various models of deformation are used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
+These learned models can then be used to make the predictions required for planning manipulation.}
+
+\comment{There are various models of deformation used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
 In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data.
@@
 
-In summary the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems.
+In summary, the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems.
 Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
-In the rest of the paper we present \comment{an overview of the model in more detail (Section~\ref{sec:model_overview});} related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).
+In the rest of the paper we present related work (Section~\ref{sec:background}); \comment{give an overview of the framework (Section~\ref{sec:model_overview});} describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).
 
@@
 
-\section{Model Overview}
-\label{sec:model_overview}
-
-As has been mentioned above there are several choices made in our modeling approach.
-The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour.
-The specific approach taken here separates the predictive model out into components in two ways (see Figure~\ref{fig:system}).
-First we decompose models according to the modality of the information.
-In this case it is important to note that deformation has a particular causal structure related to the modes of sensing that are most easily employed.
-The forces and deformations can be measured separately, using force-torque and visual sensors respectively.
-The sensed forces are the cause of the observed deformation, and this can be exploited by sequencing the models by mode.
-Thus our approach can be seen as a sensorimotor contingency in which two contigencies are sequenced to produce an overall prediction.
-The input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces.
-The second predictor takes this sequence of forces as input and predicts a deformation sequence.
-Both models must be conditioned on the material and the object.
-This is where we decompose the model space in the second way, by specialising models to particular material and object combinations.
-This follows the principle of modular motor learning \cite{KopickiICRA11,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
-Thus we aim to acquire many specialist models of different deformable objects rather than a very few rather general models.
-The specific model forms we choose here could be replaced while retaining the overall scheme.
-In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.
-
-In each case the models can be run in two modes, prediction and classification.
-In prediction mode the models make predictions based on a sequence of robot finger positions.
-To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive.
-This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.
-In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models.
-Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.
-
-The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
-For clarity the algorithm is given in three phases: training, simulation (i.e.
-the prediction phase), and classification (i.e.
-matching prediction to new data).
-In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape.
-Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).
-
 \section{Background}
@@
 
-\begin{figure}[!t]
-\centering
-\setlength\fboxsep{0pt}    % padding
-\setlength\fboxrule{0.5pt} % line width
-\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
-\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
-\label{fig:system}
-\end{figure}
-
 %%%
@@
 
-Frank et al.\ \cite{Frank2010} use a force sensor and a bumblebee stereo camera to measure the Young modulus and the Poisson ratio of unknown deformable objects.
-By assuming a homogeneous material, they propose candidate values for these constitutive parameters and simulate the expected behaviour of the object with a FEM method.
-The volume of the prediction is compared with a 3D reconstruction of the 3D surface of the object.
-A gradient descent search is used to minimize the difference.
-Once the parameters are selected, a collision detection algorithm informs the model of the position of the actuator thus implying the amount of local deformation.
-The FEM simulation provides estimates for the global deformation and response forces.
-However, a quasi-static assumption is made and a detailed study of the dynamic process is not offered\footnote{With the quasi-static assumption a small force is applied and a state of equilibrium is reached before a new force is applied.
-Therefore the deformation process is approximated
-by a succession of states of equilibrium.}.  The resulting model is applied to estimating costs of robotic navigation around deformable objects \cite{Frank2011}.
-
-The advances of this work with respect to these related approaches is summarised in \tref{t:advances}.
-
-\begin{table}[htp]
+\begin{table}[t]
   %\renewcommand{\arraystretch}{1.3}
@@
 
-Other interesting variants are: Conti et al.\ \cite{Conti2003} introduce six-degrees of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects.
+Frank et al.\ \cite{Frank2010} use a force sensor and a bumblebee stereo camera to measure the Young modulus and the Poisson ratio of unknown deformable objects.
+By assuming a homogeneous material, they propose candidate values for these constitutive parameters and simulate the expected behaviour of the object with a FEM method.
+The volume of the prediction is compared with a 3D reconstruction of the 3D surface of the object.
+A gradient descent search is used to minimize the difference.
+Once the parameters are selected, a collision detection algorithm informs the model of the position of the actuator thus implying the amount of local deformation.
+The FEM simulation provides estimates for the global deformation and response forces.
+However, a quasi-static assumption is made and a detailed study of the dynamic process is not offered\footnote{With the quasi-static assumption a small force is applied and a state of equilibrium is reached before a new force is applied.
+Therefore the deformation process is approximated
+by a succession of states of equilibrium.}.  The resulting model is applied to estimating costs of robotic navigation around deformable objects \cite{Frank2011}.
+
+The advances of this work with respect to these related approaches is summarised in \tref{t:advances}.
+
+Other interesting work includes that of Conti et al.\ \cite{Conti2003} which introduced six-degree of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects.
 The open source project Chai3D includes an implementation of this model.
@@
 %Contribution: we learn to predict with a linear regression model.  Also calibrated automatically.
+
+\section{Model Overview}
+\label{sec:model_overview}
+
+\begin{figure}[!t]
+\centering
+\setlength\fboxsep{0pt}    % padding
+\setlength\fboxrule{0.5pt} % line width
+\fbox{\includegraphics[width=3.5in]{modular.png}}
+\caption{The modular architecture, in which each context /material specific module comprises two learned models of sensorimotor contingencies.
+The first predicts the force experienced for some motion and material.
+The second predicts the deformation of the object given the applied force.
+The output of the first predictor gives the input to the second predictor.}
+\label{fig:system}
+\end{figure}
+
+As has been mentioned above there are several choices made in our modeling approach.
+The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour.
+The specific approach taken here separates the predictive model out into components in two ways (see Figure~\ref{fig:system}).
+First we decompose models according to the modality of the information.
+In this case it is important to note that deformation has a particular causal structure related to the modes of sensing that are most easily employed.
+The forces and deformations can be measured separately, using force-torque and visual sensors respectively.
+The sensed forces are the cause of the observed deformation, and this can be exploited by sequencing the models by mode.
+Thus our approach can be seen as a sensorimotor contingency in which two contigencies are sequenced to produce an overall prediction.
+The input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces.
+The second predictor takes this sequence of forces as input and predicts a deformation sequence.
+Both models must be conditioned on the material and the object.
+This is where we decompose the model space in the second way, by specialising models to particular material and object combinations.
+This follows the principle of modular motor learning \cite{KopickiICRA11,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
+Thus we aim to acquire many specialist models of different deformable objects rather than a very few rather general models.
+The specific model forms we choose here could be replaced while retaining the overall scheme.
+In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.
+
+In each case the models can be run in two modes, prediction and classification.
+In prediction mode the models make predictions based on a sequence of robot finger positions.
+To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive.
+This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.
+In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models.
+Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.
+
+\begin{figure}[t]
+\centering
+\setlength\fboxsep{0pt}    % padding
+\setlength\fboxrule{0.5pt} % line width
+\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
+\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
+\label{fig:system}
+\end{figure}
+
+The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
+For clarity the algorithm is given in three phases: training, simulation (i.e.
+the prediction phase), and classification (i.e.
+matching prediction to new data).
+In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape.
+Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).
 

### 2016-08-26 twostagedo.tex
bib and piccies

@@
 There are several novel contributions.
-First, we show how the modular learning framework of \cite{harunowolpert,kopickiwyatt16} can be applied to deformable objects.
+First, we show how the modular learning framework of \cite{kopickiwyatt16, haruno2001mosaic} can be applied to deformable objects.
 Second, we show how each module can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces.
@@
 This is where we decompose the model space in the second way, by specialising models to particular material and object combinations.
-This follows the principle of modular motor learning \cite{KopickiICRA11,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
+This follows the principle of modular motor learning \cite{kopickiwyatt16,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
 Thus we aim to acquire many specialist models of different deformable objects rather than a very few rather general models.

### 2016-08-26 twostagedo.tex
Temp

--- 2016-08-25 twostagedo.tex
@@
 The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales.
-Without such predictions planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.
-
-Various models of deformation are used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
+Without such predictions planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.}
+
+\comment{Various models of deformation are used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
 In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data.
@@
 The Canny algorithm is used to extract the edges of the image.
-Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.
+\comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge }
  \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
@@
  \STATE Be $N$ the number of sets of parameters to be evaluated per generation and $M$ the number of generations to try.
- \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements.
+ \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements\comment{, with: \\
+ $\alpha \leftarrow$ number of elements that survive for the next generation as they are, \\
+ $\beta \leftarrow$ mutated individuals, \\
+ $\gamma \leftarrow$ crossed over, and \\
+ $\zeta \leftarrow$ new individuals. \\
+ which will be chosen as follows:}
  \STATE Be $\sigma_{max} > \sigma_{min}$ standard deviations for a Gaussian

### 2016-08-27 twostagedo.tex

--- 2016-08-26 twostagedo.tex
@@
 
+\section{Model Overview}
+\label{sec:model_overview}
+
+As has been mentioned above there are several choices made in our modeling approach.
+The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour.
+The specific approach taken here separates the predictive model out into components in two ways (see Figure~\ref{fig:system}).
+First we decompose models according to the modality of the information.
+In this case it is important to note that deformation has a particular causal structure related to the modes of sensing that are most easily employed.
+The forces and deformations can be measured separately, using force-torque and visual sensors respectively.
+The sensed forces are the cause of the observed deformation, and this can be exploited by sequencing the models by mode.
+Thus our approach can be seen as a sensorimotor contingency in which two contigencies are sequenced to produce an overall prediction.
+The input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces.
+The second predictor takes this sequence of forces as input and predicts a deformation sequence.
+Both models must be conditioned on the material and the object.
+This is where we decompose the model space in the second way, by specialising models to particular material and object combinations.
+This follows the principle of modular motor learning \cite{KopickiICRA11,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
+Thus we aim to acquire many specialist models of different deformable objects rather than a very few rather general models.
+The specific model forms we choose here could be replaced while retaining the overall scheme.
+In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.
+
+In each case the models can be run in two modes, prediction and classification.
+In prediction mode the models make predictions based on a sequence of robot finger positions.
+To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive.
+This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.
+In filtering mode the predictions of different models are compared to the actual outcomes, and this is used to classify the object as being one of the modular models in the set of models.
+Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.
+
+The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
+For clarity the algorithm is given in three phases: training, simulation (i.e.
+the prediction phase), and classification (i.e.
+matching prediction to new data).
+In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape.
+Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).
+
 \section{Background}
@@
 However, they do not evaluate the quality of the overall shape prediction, and use only the final values of the prediction after the system has stabilised; that is, they do not evaluate the predictions during the dynamic phase.
+
+\begin{figure}[!t]
+\centering
+\setlength\fboxsep{0pt}    % padding
+\setlength\fboxrule{0.5pt} % line width
+\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
+\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
+\label{fig:system}
+\end{figure}
 

### 2016-08-29 twostagedo.tex
Finished learning.

@@
 The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
-Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline}.
-The tracking of the deformable target in the ground truth has two stages:
+\comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}.
+The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
+Even though this implementation is inspired in the theory of active contours, it works with a much simpler and faster, but less accurate algorithm as it is explained below, instead of minimizing a global energy.}  The tracking of the deformable target in the ground truth has two stages:
 \begin{enumerate}
@@
 The Canny algorithm is used to extract the edges of the image.
-\comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge }
+\comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.
+If the average hue of the pixels on an edge fall outside a previously calibrated range, the edge is eliminated.}
  \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
@@
 
-This measurement is used in the computer vision community to evaluate segmentation algorithms.
+This measurement \comment{is know as F-score and} is used in the computer vision community to evaluate segmentation algorithms.
 $F=1$ means that there were no false predictions ($FN+FP = 0$).
 $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).
-$F$ is calculated for every frame and the mark for the video is the mean over all the frames so far, $\mu(F)$.
+$F$ is calculated for every frame and the mark for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
+\comment{\begin{align}
+ \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
+\end{align}}
 If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closest to zero, the more errors there were.
@@
 
-Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the training data set [\fref{fig:sstrain}].
-The curve in this diagram is approximated by segmented regression curves.
+Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{than the mass-spring model.
+This diagram is obtained from the video recorded by the visual system of the robot and the forces recoded by a sensor located at the wrist.
+When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.
+From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.
+Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}
+
+The \comment{stress strain curves} in [\fref{fig:sstrain}] are approximated by segmented regression curves.
 Several candidates are proposed using the least squares technique and the best are chosen as follows.

### 2016-08-30 AnswersToComments.tex
Temp
Finished learning.
Finished reviewing technical ambiguities.

@@
 
+\comment{Section V-B?
+Added a justification at the beginning of the section.
+The meaning of $\alpha, \beta, \gamma, \zeta$ was covered on lines 7,9,10,11 but a shorter description was added when they are firstly introduced.}
+
 19, in line 60 of page 7, please explain how the object is segmented from the environment, which "simple colour segmentation is used", and how to define "edges that not likely to belong to the object of interest".
+
+\comment{Changed paragraph.}
 
@@
 
+\comment{Reference added.  Relationship with active contour explained.}
+
 21, in section V-C, please note the measurement has a name, "F-score", in machine learning community. Also, please define $\mu(f)$ based on "the mean over all the frames so far".
+
+\comment{Added the name ``F-score'' and an equation to define $\mu(f)$}
 
@@
 
+\comment{Added information.}
+
+\marker
 23, in the experiments, please compare with other existing methods.
@@
 
+\comment{Added explanations}
+
 25, in Figure 12, please clarify the legends. For example, why are there two lines with the same marker?
 
+\comment{Sorry, I don't understand this one.
+There are three lines with markers: square, circle and triangle.
+Those represent experimental data (ground data).
+The other two do not have markers because those are regression lines (predictions), but they have different colours.
+Are these ones you mention as ``lines with the same marker''?}
+
 26, in line 56 of page 10, please explain "integration step". Is it the "h" in equation 17?
+
+\comment{Yes, it is $h$, the name of the variable was added.}
 
@@
 
+\comment{Changed ``for novel interactions'' for a more concrete description.}
+
+\comment{Here we are not focusing on classifiying static objects in new environments, but rather in classifying the material an object is made of when the entry is its behaviour in time.
+Visually, the blocks are actually very similar prior to deformation, therefore the most relevant difference lies in their dynamic behaviour.
+That is why the classification task looks different from traditional static datasets.
+Instead of a large collection of labeled images in different backgrounds, we used six videos of two different materials and asked the system to identify which material it looks at through the videos.
+The explanation in the article was modified to be more specific about this.}
+
 28, in Table III, what does "test 1, test 2" mean? Please explain how these numbers are computed. Besides, object classification is typically evaluated with precision-recall curve.
+
+\comment{An explanation about "training", "test 1, test 2" was added.}
+
+\comment{Since there are only two materials the number of false positives for one material is exactly the false negatives for the other material, that is why we considered that recall would be redundant.}
 
@@
 
-\comment{We don't focus on tracking, we want to predict several frames ahead, not follow, that is why closely related works are few.}
+\comment{Added to the background section.}
 

### 2016-08-30 twostagedo.tex
Finished reviewing technical ambiguities.

@@
 
-\section{Model Overview}
-\label{sec:model_overview}
-
-As has been mentioned above there are several choices made in our modeling approach.
+% \section{Model Overview}
+% \label{sec:model_overview}
+% 
+% As has been mentioned above there are several choices made in our modeling approach.
 The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour.
@@
 In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.
-
-In each case the models can be run in two modes, prediction and classification.
+% 
+% In each case the models can be run in two modes, prediction and classification.
 In prediction mode the models make predictions based on a sequence of robot finger positions.
@@
 Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.
-
-The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
-For clarity the algorithm is given in three phases: training, simulation (i.e.
+% 
+% The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
+% For clarity the algorithm is given in three phases: training, simulation (i.e.
 the prediction phase), and classification (i.e.
@@
 Acquired models are used to classify materials through recognition of trained samples, even when the interactions are novel.
-Even though there is abundant literature in the industrial robotics area for 1D and 2D materials (e.g.
-strings, hair, metallic sheets and textiles), there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.
+Even though there \comment{are lots of works on deformable object tracking, some of which make use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are used only for the next frame and can not be used for robotic planning, where predictions must be done for several frames ahead\cite{Cretu2012}.
+Others like \cite{Malassiotis1998, Cremers2006} make use of templates or make assumptions about the shape of the objects that would not work with fully deformable objects like plasticine.
+There is also} abundant literature in the industrial robotics area for 1D and 2D materials (e.g.
+strings, hair, metallic sheets and textiles), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.
 
@@
 \subsection{Evolutionary Algorithm}
-\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they depend on the geometry of object, and there can be several sets of parameters in different regions of the space of parameters that produce adequate simulations \cite{Morris2008}, we chose to use an evolutionary algorithm to expore different regions and try improvements in the vecinity of successful individuals.}
+\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they depend on the geometry of the mesh used to model the object, and there can be several sets of parameters in different regions of the space of parameters that produce adequate simulations \cite{Morris2008}, we chose to use an evolutionary algorithm to expore different regions and try improvements in the vecinity of successful individuals.}
 
@@
 \hline
-\multicolumn{4}{l}{Where the mark is $\mu(F)$ as explained in \sref{sec:evaluation}.}
+\multicolumn{4}{p{88mm}}{Where the mark is $\mu(F)$ as explained in \sref{sec:evaluation}.
+\comment{Mean $\mu(F)$ and standard deviation across 24 training sesions are reported for the best performing models of every training session, worst model and average model.}}
 \end{tabular}
@@
 One hundred photographs and force readings where taken every $0.17\pm0.007 s$ in average.
-The mass spring model runs with an integration step of $0.1s$, where forces that were not measured directly are interpolated.
+The mass spring model runs with an integration step \comment{$h=0.1s$}, where forces that were not measured directly are interpolated.
 The push down movement lasts approximately 50 frames and the retrieval, the other 50.
@@
 
-The calibrated mass-spring models were used to classify objects through recognition of their behaviour, even for novel interactions.
-The models make predictions as if the object was known.
-The model that makes the best predictions indicates which object is being observed.
-There are two criteria for selecting an object:
+The \comment{\sout{calibrated}} mass-spring models \comment{calibrated for the prediction task} were used to classify \comment{\sout{objects} materials} through recognition of their behaviour, even \comment{\sout{for novel interactions} when the objects where pushed and deformed at novel locations.}  The models make predictions as if the \comment{material of the} object was known.
+The model that makes the best predictions indicates which \comment{type of} object is being observed.
+There are two criteria for selecting an \comment{object/material for the video being shown to the system}:
 
 \begin{enumerate}
- \item Per frame: an object is selected for each frame of the interaction. See \tref{tab:classperframe}.
+ \item Per frame: \comment{\sout{an} a candidate} object is selected for each frame of the interaction \comment{according to which model made the best prediction for that frame in particular, as marked by the F-score.
+This test was performed for every frame of the six videos, where the robotic finger pushes the block of material.
+\tref{tab:classperframe} shows the results for the videos that were used as trainning and test sets for the prediction task.}
 
@@
 
- \item Global: The difference in the marks is added for all frames.  Frames where the distinction is more noticeable will receive a greater weight.
+ \item Global: \comment{The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted mark.} The difference in the \comment{F-score} marks is added for all frames.
+Frames where the distinction is more noticeable will receive a greater weight.
 \end{enumerate}

### 2016-09-06 twostagedo.tex
rewrote model overview, fixed biog, added figures, small edits.

@@
 % could we change the title to this? What do you think? I wanted to get the modular idea in there.
-\title{Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects
+\title{A Multi-Modal Model of Object Deformation for Robotics
 %A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
@@
 \markboth{IEEE CDS}%
-{Arriola-Rios and Wyatt: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
+{Arriola-Rios and Wyatt: A Multi-Modal Model of Object Deformation for Robotics}
 % The only time the second header will appear is for the odd numbered pages
@@
 \begin{abstract}
-\comment{In this paper we present a modular, multi-modal framework for learning generative models of object deformation under robotic pushing.
+\comment{In this paper we present a multi-modal framework for learning generative models of object deformation under robotic pushing.
 The model is multi-modal in that it is based on integrating force and visual information.
-The framework is also modular, being composed of a learned module for each type of material.
-Each such module is learned by parameter estimation for a general model form.
-When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of the modules are compared so as to classify the material.} Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
+The framework consists of several sub-models that are independently learned from the same data.
+These component models can be sequenced to provide many-step prediction and classification.
+When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
 While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning framework, and none has offered classification from a generative model.
@@
 A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations.
-Different materials (elastic, plastic) have different sensorimotor contingencies.
+Different materials (here, elastic and plastic materials) have different sensorimotor contingencies.
 The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data.
-The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales.
+The predictions of a learned model can be fed back into itself, so as to predict an object's deformation behaviour over many time steps.
 These learned models can then be used to make the predictions required for planning manipulation.}
@@
 In this paper we present a framework for solving all these problems simultaneously.
+Although our framework requires two separate learned models, they are combined into a single machine for prediction of the behaviour of deformable material.
+In a second machine, multiple copies of this predictor can be compared to an actual outcome deformation, so as to classify the material.
 
 There are several novel contributions.
-First, we show how the modular learning framework of \cite{kopickiwyatt16, haruno2001mosaic} can be applied to deformable objects.
-Second, we show how each module can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces.
+First, we show how the learning problem can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces.
 Both models are generative, and we show how to learn each from data.
@@
 
-\begin{figure}[!t]
-\centering
-\setlength\fboxsep{0pt}    % padding
-\setlength\fboxrule{0.5pt} % line width
-\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
-\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
-\label{fig:system}
-\end{figure}
+%\begin{figure}[!t]
+%\centering
+%\setlength\fboxsep{0pt}    % padding
+%\setlength\fboxrule{0.5pt} % line width
+%\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
+%\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
+%\label{fig:system}
+%\end{figure}
 
@@
 \centering
-\setlength\fboxsep{0pt}    % padding
-\setlength\fboxrule{0.5pt} % line width
-\fbox{\includegraphics[width=3.5in]{modular.png}}
-\caption{The modular architecture, in which each context /material specific module comprises two learned models of sensorimotor contingencies.
-The first predicts the force experienced for some motion and material.
-The second predicts the deformation of the object given the applied force.
-The output of the first predictor gives the input to the second predictor.}
-\label{fig:system}
+%\setlength\fboxsep{0.5pt}    % padding
+%\setlength\fboxrule{0.5pt} % line width
+%\fbox{
+\includegraphics[width=3.5in]{figures/learning.png}%}
+\caption{The learning for a specific material comprises two learned models of sensorimotor contingencies.
+Both are learned from the same data.
+The first (FP) learns to predicts the force experienced for some motion and material.
+The second (SP) predicts the deformation of the object given the applied force.
+The strain $st$ is calculated directly from the interpenetration of the finger into the object.
+This can be calculated by vision (V) or from the planned motion of the finger.}
+\label{fig:learning}
 \end{figure}
 
-As has been mentioned above there are several choices made in our modeling approach.
+\comment{The framework consists of three separate systems: a learning system, a prediction system, and a classification system.
+The learning system is depicted in Figure~\ref{fig:learning}.
+This enables learning of sensorimotor contingencies from real robot data.
+Learning of the force prediction (FP) model relies on comparing the predicted $\hat{f}$ and the experienced $f$ forces at the finger tip, as measured by a force-torque sensor (FM).
+Learning of the shape prediction (SP) model relies on comparing the predicted deformation $\hat{sh}$ with the actual deformation $sh$ measured by visual tracking (V).
+}
+
+\comment{The two learned models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{ig:prediction}).
+This combined model takes a planned motion of the finger, and predicts the resulting forces experienced.
+It then uses both the position and forces to predict the resulting object deformation.
+Typically a sequence of planned finger positions will be known.
+In prediction mode the models make predictions based on this sequence.
+To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive.
+This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.}
+
+\begin{figure}[t]
+\centering
+%\setlength\fboxsep{0pt}    % padding
+%\setlength\fboxrule{0.5pt} % line width
+%\fbox{
+\includegraphics[width=3.5in]{figures/prediction}%}
+\caption{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way.
+The predicted shape is fed back on itself so as to predict over many steps for a planned sequence of finger motion.
+PC is the position controller for the robot which outputs each planned finger position $p$ at each time.
+This is used to calculate the strain $st$ from the predicted interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}
+\label{fig:prediction}
+\end{figure}
+
+\comment{The third system is for material classification (Figure~\ref{fig:classification}).
+This simply comprises multiple copies of the prediction systems, one for each type of material that an unknown object might be made of.
+The predictors each make many-step predictions.
+At each step a classification is made, based on which predictor has made the better prediction of the shape at that step.}
+
+\comment{In summary this way of learning and using models has involved several choices.
+First, the learned models are generative.
+Second, our approach can be seen as a sensorimotor contingency in which two contingencies are sequenced to produce an overall prediction.
+Third, although here we only compare two material models for prediction, the scheme follows the principle of modular motor learning \cite{kopickiwyatt16,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
+Thus the aim of a generalised version of our scheme is to acquire many specialist models of different deformable objects rather than a very few rather general models.
+The specific model forms we choose here could be replaced while retaining the overall scheme.
+In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.}
+
+%As has been mentioned above, there are several choices made in our modeling approach.
 The key choice is that the models are generative, which means that the models can be used to generate predicted sequences of behaviour.
@@
 The sensed forces are the cause of the observed deformation, and this can be exploited by sequencing the models by mode.
-Thus our approach can be seen as a sensorimotor contingency in which two contigencies are sequenced to produce an overall prediction.
+Thus our approach can be seen as a sensorimotor contingency in which two contingencies are sequenced to produce an overall prediction.
 The input to the first predictor is the planned motion of the robot (or the joint torques), and this predictor predicts the resulting sequence of contact forces.
@@
 This follows the principle of modular motor learning \cite{kopickiwyatt16,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
-Thus we aim to acquire many specialist models of different deformable objects rather than a very few rather general models.
+Thus the aim of a generalised version of our scheme is to acquire many specialist models of different deformable objects rather than a very few rather general models.
 The specific model forms we choose here could be replaced while retaining the overall scheme.
@@
 
-In each case the models can be run in two modes, prediction and classification.
+\begin{figure*}[t]
+\centering
+%\setlength\fboxsep{0pt}    % padding
+%\setlength\fboxrule{0.5pt} % line width
+%\fbox{
+\includegraphics[width=0.8\textwidth]{figures/classification}%}
+\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
+The most accurate predictor at any one moment gives the classification for the material type.
+The notation is explained in the text.}
+\label{fig:classification}
+\end{figure*}
+%In each case the resulting learned models can be run in two modes, prediction and classification.
 In prediction mode the models make predictions based on a sequence of robot finger positions.
@@
 Since the model is multi-modal we can compare both force and deformation data to the corresponding model predictions.
-
-\begin{figure}[t]
-\centering
-\setlength\fboxsep{0pt}    % padding
-\setlength\fboxrule{0.5pt} % line width
-\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
-\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
-\label{fig:system}
-\end{figure}
 
@@
 
-\begin{figure}[!t]
+\begin{figure}[t]
 \centering
@@
   Ph.D. in Artificial Intelligence from the University of Edinburgh in
-  1997. He has published more than 80 refereed articles. His interests
+  1997.
+He was a Leverhulme Fellow from 2006-8.
+He has won two best paper prizes, published more than 100 refereed articles, led two international research projects on robot planning and learning (CogX) and robot manipulation (PaCMan), and edited three books.
+His interests
   include machine learning, planning, architectures, vision, mobile robotics and robot manipulation.

### 2016-09-06 twostagedo.tex
small change to model overview, fixed missing cross reference

@@
 
-\comment{The two learned models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{ig:prediction}).
+\comment{The two learned models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{fig:prediction}).
 This combined model takes a planned motion of the finger, and predicts the resulting forces experienced.
@@
 
-The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
-For clarity the algorithm is given in three phases: training, simulation (i.e.
-the prediction phase), and classification (i.e.
-matching prediction to new data).
-In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape.
-Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).
+%The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
+For clarity the algorithm is described in exactly these three phases: training, simulation (i.e.
+the prediction phase), and classification.
+In empirically evaluating the model new data was used, acquired from new finger pushes.
+We now proceed to describe the shape predictor SP (Section~\ref{sec:model}) which is based on a mass-spring system, and then the force predictor (FP), which is a regression model (Section~\ref{sec:forces}).
 

### 2016-09-09 AnswersToComments.tex
corrections to text, and to letter, updated diagram.

@@
 This article aims at tracking three-dimensional deformable objects under robot motion and predicting the force causing the deformation.
-
-\comment{It is important to make a precision here: the article aims at predicting the force and deformation on the visual plane of the three-dimensional object.
-The tracking of the shape is used only during training as part of the evaluation function.}
-
+%
+%\comment{While this is entirely correct we would like to say that this was not the intended emphasis of the article.
+The article primarily aims at predicting both the force and the deformation.
+The tracking of the shape is used only during training as part of the evaluation function, it is not the goal.
+We apologise in case this wasn't clear, probably it was.
+Nevertheless we have revised the paper throughout to try to make this clearer.}
+%
 For these purposes, the authors propose two models, including an extension of the mass-spring model and a regression model, where the parameters of the mass-spring model are learned with an evolutionary searching algorithm on visual data and force data, and the parameters of the regression model are learned with least-square methods on stress-strain graph.
@@
 
+\comment{Thank you for pointing this out.
+We have rewritten the model overview section, parts of the abstract, and the introduction to make this clearer.
+We emphasise that on the one hand we do learn the two models separately, but on the other we also emphasise that they are combined in prediction, and further that exactly the same core prediction model is used to perform classification.
+Thus although we do not claim them as a single model, we do want to claim them as several models linked in a over-arching framework.
+We have made clear the exact structures in three new figures.}
+
 2: section II is not included in the final paragraph of section I, and this section should be included in section I or section IV.
 
-\comment{It was included in the final paragraph of section I.  It was not included in section I to keep the general idea of the model isolated from other aspects in the introduction.}
+\comment{We have now mentioned this.}
 
@@
 
+\comment{Thanks for pointing this out.
+We agree, and have completely rewritten section II.
+In part this addresses this point, and in part it makes clearer the multiple machines that we have in our framework.
+We hope that we have made it much clearer.}
+
 3: section IV-C, "Integration Scheme", seems to be part of algorithm 1 or section IV-B, thus it should be introduced.
 
-\comment{It is neither part of algorithm 1, nor of section IV-B.
-It is part of a numerical algorithm for solving differential equations.
+\comment{We are sorry this was not clear.
+In fact it is neither part of algorithm 1, nor of section IV-B.
+It is simply part of a numerical algorithm for solving differential equations.
 \newline TODO: Add general algorithm which makes use of all the pieces explained in section IV at the beginning of the section.}
@@
 4, section IV-D is not explained: what is the relation between "collision detection" and the proposed model?
+
+\comment{We have now rewritten the start of section IV-D to make this clear.
+Essentially, the simulation must not allow interpenetration between objects.
+Thus, corrections are made to the mass-spring predictions at run time, to respect these constraints.
+Detection of violations requires some collision checking.}
 
@@
 
-\comment{Changed order as suggested.}
+\comment{We have changed the order as suggested.}
 
@@
 
-Associate Editor's Comments to Author:
-
-Associate Editor 
-Comments to the Author: 
-(There are no comments. Please check to see if comments were included as a file attachment with this e-mail or as an attachment in your Author Center.)
-
 I look forward to hearing from you.

### 2016-09-09 twostagedo.tex
corrections to text, and to letter, updated diagram.

@@
 \markboth{IEEE CDS}%
-{Arriola-Rios and Wyatt: A Multi-Modal Model of Object Deformation for Robotics}
+{Arriola-Rios and Wyatt: A Multi-Modal Model of Object Deformation for Robotic Pushing}
 % The only time the second header will appear is for the odd numbered pages
@@
 These component models can be sequenced to provide many-step prediction and classification.
-When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
-While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning framework, and none has offered classification from a generative model.
+When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our approach, which consists of learning and combining multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
+While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model.
+We do so through separately learned models which can be combined in different ways for different purposes.
 \end{abstract}
@@
 elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?).
-In this paper we present a framework for solving all these problems simultaneously.
+In this paper we present an approach for solving all these problems.
 Although our framework requires two separate learned models, they are combined into a single machine for prediction of the behaviour of deformable material.
@@
 
-In summary, the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems.
-Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
+In summary, our approach enables us to learn models of sensorimotor contingencies, and then to use these as components to build different machines.
+This approach enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
 In the rest of the paper we present related work (Section~\ref{sec:background}); \comment{give an overview of the framework (Section~\ref{sec:model_overview});} describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).
@@
 \label{sec:background}
-This work focuses on prediction of response forces and of the deformation of the shape over several frames of elastic and plastic objects when they are pushed and released by a robotic finger.
-Acquired models are used to classify materials through recognition of trained samples, even when the interactions are novel.
-Even though there \comment{are lots of works on deformable object tracking, some of which make use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are used only for the next frame and can not be used for robotic planning, where predictions must be done for several frames ahead\cite{Cretu2012}.
-Others like \cite{Malassiotis1998, Cremers2006} make use of templates or make assumptions about the shape of the objects that would not work with fully deformable objects like plasticine.
-There is also} abundant literature in the industrial robotics area for 1D and 2D materials (e.g.
-strings, hair, metallic sheets and textiles), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.
+This work focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger.
+These predictions include the response forces and the deformations of the object's shape over many future steps.
+The predictive models can be used to classify materials, even when the interactions (e.g.
+the contact point) are novel.
+Although there \comment{is a large body of work on deformable object tracking, some of which makes use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are typically used only for the next frame and can not be used for robotic planning, where predictions must be made many steps into the future \cite{Cretu2012}.
+Others like \cite{Malassiotis1998, Cremers2006} make use of templates or make assumptions about the shape of the objects that would not work with highly deformable objects like plasticine.
+There is also} an abundant literature in industrial robotics for modelling 1D and 2D materials (e.g.
+string, hair, metallic sheet and textile), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.
 
@@
 Their model of the material can predict a net amount of deformation given the applied forces, but does not evaluate details of the deformed shape.
-This model is inspired in the crystalline structure of atoms in solids.
+This model is inspired by the crystalline structure of atoms in solids.
 The space lattice of the crystal is approximated by a particle system where elements with mass are connected by springs and dampers, thus representing the attraction and repulsion between atoms \emph{in equilibrium} with a damped mass-spring system.
@@
 \begin{itemize}
- \item the force applied while pushing against the material with two robotic arms in opposite directions and,
+ \item the force applied while pushing against the material with two robot arms in opposite directions and,
  \item the global displacement and velocity vectors resulting from adding the displacements and velocities of all particles.
 \end{itemize}
-The estimated values of the mass, deformability and damping coefficients are used as the entry of a neural network that outputs the minimum necessary force to lift the object without sliding.
+The estimated values of the mass, deformability and damping coefficients are used as the input to a neural network that outputs the minimum necessary force to lift the object without sliding.
 This function is used to compliantly adjust the force during an interaction.
@@
 
-Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, whose constraints are expressed as potential energy terms.
+Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, in which constraints are expressed as potential energy terms.
 These terms encode the tendency of springs to recover their length (with damping), of triangular faces of a mesh to recover their area and of tetrahedrons to recover their volume.
@@
 However, they only compare the steady states and do not include plastic deformations.
-We extend on this work, by changing the FEM model with force and 2D visual data gathered from real materials.
-Furthermore, we analysed the quality of the predictions while the deformations take place.
+We extend this work, by replacing the FEM model with force and 2D visual data gathered from real materials.
+Furthermore, we analyse the quality of the predictions while the deformations take place.
 
@@
 A gradient descent search is used to minimize the difference.
-Once the parameters are selected, a collision detection algorithm informs the model of the position of the actuator thus implying the amount of local deformation.
+Once the parameters are selected, a collision detection algorithm informs the model of the position of the actuator, thus implying the amount of local deformation.
 The FEM simulation provides estimates for the global deformation and response forces.
-However, a quasi-static assumption is made and a detailed study of the dynamic process is not offered\footnote{With the quasi-static assumption a small force is applied and a state of equilibrium is reached before a new force is applied.
+However, a quasi-static assumption is made and a detailed study of the dynamic process is not offered.\footnote{With the quasi-static assumption a small force is applied and a state of equilibrium is reached before a new force is applied.
 Therefore the deformation process is approximated
-by a succession of states of equilibrium.}.  The resulting model is applied to estimating costs of robotic navigation around deformable objects \cite{Frank2011}.
-
-The advances of this work with respect to these related approaches is summarised in \tref{t:advances}.
+by a series of states of equilibrium.}  The resulting model is applied to estimating costs of robotic navigation around deformable objects \cite{Frank2011}.
+
+The advances of our work with respect to these related approaches is summarised in \tref{t:advances}.
 
@@
 Again, there is no analysis of the dynamics and no real objects are used.
-Furthermore, this model is more adequate for objects that deform preferably around fixed joints.
+Furthermore, this model is more adequate for objects that deform around fixed joints.
 
@@
 
-Additionally, there is a vast amount of literature on modeling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.
-However, these types of models have not been successfully applied to robotic tasks beyond what was presented above.
+Additionally, there is a vast literature on modeling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.
+However, these types of models have not been successfully applied to robotics tasks beyond what is presented above.
 
@@
 \includegraphics[width=3.5in]{figures/learning.png}%}
-\caption{The learning for a specific material comprises two learned models of sensorimotor contingencies.
+\caption{The learning method comprises two learned models of sensorimotor contingencies.
 Both are learned from the same data.
@@
 \comment{The two learned models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{fig:prediction}).
-This combined model takes a planned motion of the finger, and predicts the resulting forces experienced.
-It then uses both the position and forces to predict the resulting object deformation.
-Typically a sequence of planned finger positions will be known.
+This combined model takes a planned motion of the finger, and predicts the resulting forces that will be experienced at the contact points.
+It then uses both the planned positions and predicted forces to predict the resulting object deformations.
+Typically, a sequence of planned finger positions will be known.
 In prediction mode the models make predictions based on this sequence.
@@
 The predicted shape is fed back on itself so as to predict over many steps for a planned sequence of finger motion.
-PC is the position controller for the robot which outputs each planned finger position $p$ at each time.
-This is used to calculate the strain $st$ from the predicted interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}
+PC is the position controller for the robot, which outputs each planned finger position $p$ at each time.
+This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}
 \label{fig:prediction}
@@
 
-\comment{In summary this way of learning and using models has involved several choices.
+\comment{In summary, this way of learning and using models involves several choices.
 First, the learned models are generative.
-Second, our approach can be seen as a sensorimotor contingency in which two contingencies are sequenced to produce an overall prediction.
+Second, our approach can be seen as one in which two contingencies are learned separately, and then sequenced to produce an overall prediction.
 Third, although here we only compare two material models for prediction, the scheme follows the principle of modular motor learning \cite{kopickiwyatt16,haruno2001mosaic}, where prediction and control is specialised into many modules, each of which covers a relatively small portion of the input space, such as an object or material type.
-Thus the aim of a generalised version of our scheme is to acquire many specialist models of different deformable objects rather than a very few rather general models.
+Thus the aim of a generalised version of our scheme will be to acquire many specialist models of different deformable objects rather than a very few rather general models.
 The specific model forms we choose here could be replaced while retaining the overall scheme.
-In this paper we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.}
+In this paper, we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.}
 
@@
 %The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
-For clarity the algorithm is described in exactly these three phases: training, simulation (i.e.
+For clarity, the algorithm is described in exactly these three phases: training, simulation (i.e.
 the prediction phase), and classification.
 In empirically evaluating the model new data was used, acquired from new finger pushes.
-We now proceed to describe the shape predictor SP (Section~\ref{sec:model}) which is based on a mass-spring system, and then the force predictor (FP), which is a regression model (Section~\ref{sec:forces}).
+We now proceed to describe the shape predictor SP (Section~\ref{sec:model}), which is based on a mass-spring system, and then the force predictor (FP), which is a regression model (Section~\ref{sec:forces}).
 
@@
 \comment{
-A mass-spring model is a very well established abstraction used frequently to simulate the behaviour of some deformable objects.
-As a physics based model, once its parameters are properly assigned, its behaviour is determined by the evolution of a set of differential equations.
-Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the interacting forces are know.
-The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model for the particular object.
+A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.
+As a physics based model, once its parameters are acquired, its behaviour is determined by the evolution of a set of differential equations.
+Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the externally applied forces are known.
+The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model as a model of the particular object.
 For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.
-In this work we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration to a system that exploits it to solve prediction and classification tasks.
+In this work, we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration into a system that exploits it to solve prediction and classification tasks.
 We emphasize that the physics based model could be easily substituted by another equivalent model.
@@
 At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located, and is known as a \textit{mass particle}.
-The edges between vertices roughly model the interactions between particles, that attract each other up to a certain critical distance, but repel each other if they become closer, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.
+The edges between vertices roughly model the interactions between particles, which attract each other up to a certain critical distance, but repel each other if they become closer, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material on itself.
 The simplest model associates a linear spring with each edge, as if those forces would occur only between connected neighbours.
@@
 
-When one or more particles are displaced from their positions of equilibrium, the forces that tend to restore an elastic body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).
+When one or more particles are displaced from their equilibirum positions, the forces that tend to restore an elastic body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).
 These forces act upon the particles modifying their positions in time.
@@
 \end{align}
-where $t$ is time.
-However, it is possible to modify this equation according to need if what we need is to model a different behaviour.
-%are interested in is numerical accuracy, as it will be done here.
-
-The previous equation is a second order differential equation, which is continuous by definition.
+where $t$ is time.  However, it is possible to deviate from this equation if we need to model a different behaviour. %are interested in is numerical accuracy, as it will be done here.
+
+The above equation is a second order differential equation, which is continuous by definition.
 In order to perform a computer simulation we need to obtain the position $p$ of all vertices as a function of time $p(t)$ using this equation.
@@
 Time is discretized in intervals of length $h$, positions and forces at discrete past times are used to estimate the new positions at time $t + h$, according to an integration scheme which must be selected.
-The fact that the value of the forces remain constant in this approximation, for the whole interval $h$ (also know as \textit{time step}), introduces errors which can be severe.
+The fact that the values of the forces remain constant in this approximation, for the whole interval $h$ (also know as \textit{time step}), introduces errors which can be severe.
 For this reason, computer simulations may add procedures that help to alleviate these effects.
@@
 The potential energy terms for the preservation of length with damping and preservation of area were included in our system \comment{exactly as they were defined by Teschner}.
-To compensate for the lack of a preservation of volume, which tends to avoid distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.
+To compensate for the lack of a term for preservation of volume, which tends to prevent distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.
 The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.
@@
 \end{align}
-Given the oscilatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.
+Given the oscillatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.
 This is modelled with a force acting in the opposite direction to the velocity of the affected particles\comment{, thus slowing down motion.
@@
 \subsubsection{Preservation of Area}
-This constraint is applied per triangle in the mesh.
-Since Teschner and Morris \comment{\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} found that the use of damping for the preservation of areas does not improve significantly the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
+This constraint is applied per triangle in the mesh.  Since Teschner and Morris \comment{%\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} 
+found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
 \begin{align}
@@
 [\fref{fig:forces}(c)] gives an idea of how these forces look.
-Here the energy depends on the difference \comment{\sout{of the angles between adjacent edges.} between the current angle, between adjacent edges, and the angle between them at the equilibrium position.}
-
-Energy\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
+Here the energy depends on the difference \comment{%\sout{of the angles between adjacent edges.}
+between the current angle, between adjacent edges, and the angle between them at the equilibrium position.} Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
 \begin{align}
@@
 
-Where $\varphi$ is the angle between adjacent edges, $E_\varphi$ is the energy associated to changes in the angle, $k_\varphi$ is the corresponding stiffness constant and the $p_i$s are the Cartesian coordinates of the mass particles.
+where $\varphi$ is the angle between adjacent edges, $E_\varphi$ is the energy associated with changes in the angle, $k_\varphi$ is the corresponding stiffness constant, and the $p_i$s are the Cartesian coordinates of the mass particles.
 
@@
 It tends to restore the angles in the most efficient way, but does not take the original size into account [\fref{fig:forces}(c)].
-Therefore, it helps to recover a similar triangle, but it also may produce tiny or very big triangles if it is not accompanied by some of the other terms that tend to restore the original dimensions, in addition to the angles.
+Therefore, it helps to recover a similar triangle, but it may produce tiny or very big triangles if it is not accompanied by some of the other terms that tend to restore the original dimensions, in addition to the angles.
 The force is derived from:  %An additional line in the code also forces an inverted angle to recover its original orientation.
@@
 \subsection{Collision Detection}
-In the scenario, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.
-\comment{When the simulation of the behaviour of the mesh, that models the image of the sponge, is carried on, the predicted positions of the vertices may indicate that the mesh penetrates the obstacles.
+In the simulation, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.
+\comment{When the simulation of the behaviour of the mesh, which is used to model the sponge, is run, the predicted positions of the vertices may indicate that the mesh penetrates the obstacles.
 This should be interpreted as a collision between the sponge and the obstacle.
-Since the sponge can not penetrate an obstacle, the overlapping between geometries must be solved, and its physical consequences must affect the simulation.
-For example: if vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
-
-In order to model this,} two types of \comment{solid} obstacles are considered: circles \comment{(for the finger)} and a table line that can not be crossed.
-For the table obstacle, the collision \comment{of parts of the sponge with it} is roughly solved by pushing vertices \comment{of the mesh of the sponge} back to the border of the table.
-The springs will eventually propagate the effect to the neighbours.
-For the circle the resolution is in two stages: vertices and edges.
+Since the sponge can not penetrate an obstacle, these overlaps must be eliminated.
+%The physical consequences of this will affect the system predictions.
+For example: if the vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
+}
+
+In order to model this, we model the finger and table using geometric primitives.
+For the table obstacle, any interpenetration by \comment{the sponge} is solved by pushing vertices \comment{of the mesh of the sponge} back to the border of the table.
+The springs will then propagate the effect to their neighbours.
+For the finger obstacle, resolution is in two stages: by vertices and then edges.
 A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.
-For edges, the common algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the circle and the edge.
-Given this distance, the algorithm in \alref{alg:push_edge} indicates how and in which direction the edge must be displaced.
+For edges, the standard algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.
+Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.
 
@@
 
-\subsection{Many Steps Prediction}
+\subsection{Many Step Prediction}
 \label{sec:manysteps}

### 2016-09-09 AnswersToComments.tex
Cover letter for revision is AnswersToComments.tex

@@
 
-Please find attached our submission:
+Please find attached our revised submission:
 \begin{center}
@@
 \end{center}
+
+We thank the reviewer for the comments and we have made our best effort to address them.  In the following text we explain in some detail the changes we did.
+
+Before that, we also wish to apologise because, in spite of our efforts, we still have not been able to implement the comparison with other existing methods, as any fair comparison requires an extension/implementation to closely related methods which do not work under the same conditions that ours.
+We are working on that, but the deadline has come and we will submit our work leaving only this point pending.
+If you could allow us to submit another version in four weeks, we should be able to include this comparison.
+
+Thank you very much for everything,
+
+Veronica E. Arriola
+
+\vspace{2cm}
 
@@
 This article aims at tracking three-dimensional deformable objects under robot motion and predicting the force causing the deformation.
-%
+
+%\marker
 %\comment{While this is entirely correct we would like to say that this was not the intended emphasis of the article.
@@
 %
+
 For these purposes, the authors propose two models, including an extension of the mass-spring model and a regression model, where the parameters of the mass-spring model are learned with an evolutionary searching algorithm on visual data and force data, and the parameters of the regression model are learned with least-square methods on stress-strain graph.
@@
 
-\marker
+%\marker
 1: in the abstract, the authors claim that five features are simultaneously enabled by the proposed single learning model framework, but the article actually presents a collection of different models and methods to solve different problems, thus it is inappropriate to claim them as a single unified model framework.
@@
 It is simply part of a numerical algorithm for solving differential equations.
-\newline TODO: Add general algorithm which makes use of all the pieces explained in section IV at the beginning of the section.}
+We hope the new introduction to section IV will help to make it clearer.}
+% \newline TODO: Add general algorithm which makes use of all the pieces explained in section IV at the beginning of the section.  Canceled: we had one but it was too complex to follow.
 
@@
 
-\comment{TODO: Move section IV-E after IV-A but must emphasize that the geometric constraints in IV-A are enforced through energy preservation, while those in IV-E have immediate effects on the triangulations without directly affecting the equations.}
+\marker
+%\comment{TODO: Move section IV-E after IV-A but must emphasize that the geometric constraints in IV-A are enforced through energy preservation, while those in IV-E have immediate effects on the triangulations without directly affecting the equations.}
+\comment{TODO: Chage title of section E to ``Geometric constraints outside from the mass-spring model''.
+``Even with the addition of the term for preservation of angles to the mass-spring simulation...''}
+
+\comment{Thank you for the observation, it shows the paper was not fully explaining the difference between the constraints presented in section IV-A and those in section IV-E.
+The first belong to the mass-spring model (as everything being explained in subsections A to C).
+The routines in section IV-E are independent and more closely related to mesh-related fixes like the one explained in section IV-D.
+We changed the title of section IV-E to reflect this and further explained this difference in its text.}
 
@@
 
-\comment{There is a missunderstanding here.
-Section IV-F is independent of any training.
-As a physics based model, it depends only on proper entries and parameters, which could be provided by a human expert after manual calibration or by a computer algorithm.
-The last paragraph of this section was modified accordingly and the algorithm introduced for comment 3 should help here.}
+\comment{Sorry the text was confusing.
+Section IV-F does not depend on training.
+As a physics based model, the dynamics of the simulation depend only on proper entries and parameters, which could have been provided by a human expert after manual calibration or by a computer algorithm.
+The introduction to section IV and the last paragraph of this section was modified to emphasise this.}
 
@@
 
-\marker
 1, in line 37 of page 1, "sensorimotor contingencies" needs to be defined or referenced. 
+\comment{Now the first paragraph of the Introduction contains a definition.}
 
-\marker
 2, in line 35 of page 1, what does "sequencing the models by mode" mean? 
+\comment{This has been addressed as part of the rewrite of the Model Overview section.}
 
@@
 
-\comment{The general idea had been explained in the first paragraph of section IV-A, the details came later.  Since it was not clear enough the wording was modified.}
+\comment{Since it was not clear the wording of the first paragraph of section IV-A was modified, where the general idea is explained, the details come in later paragraphs.}
 
@@
 
-\comment{Since for comment 5 the introduction explained that a mass particle is located on vertex i, this has been covered already.}
+\comment{The introduction added to address comment 5 covers this issue as well.}
 
@@
 
-\comment{It is not within the scope of the article to explain that derivation, since it belongs to Teschner's model, however the cite to the article we got it from was repeated here.}
+\comment{It is not within the scope of the article to explain that derivation, since it belongs to Teschner's model, however the reference to the article we got it from was repeated here.}
 
@@
 
-\comment{It was ``or''.  Corrected.}
+\comment{It was ``or'', thanks.  Corrected.}
 
@@
 
-\comment{No, that is why we mentioned in the text that it is a common algorithm.  It can be found across the web in many sites.  Should we mention something about that in the caption?}
+\comment{No, it can be found across the web in many sites.  We changed the word ``common'' in the text, for ``standard'', which should emphasise more this fact.}
 
@@
 
-\comment{Here we are not focusing on classifiying static objects in new environments, but rather in classifying the material an object is made of when the entry is its behaviour in time.
+\comment{Here we are not focusing on classifying static objects in new environments, but rather in classifying the material an object is made of when the entry is its behaviour in time.
 Visually, the blocks are actually very similar prior to deformation, therefore the most relevant difference lies in their dynamic behaviour.
@@
 
-\comment{Added to the background section.}
+\comment{Added to the background section, alongside the statement ``those predictions are used only for the next frame and can not be used for robotic planning, where predictions must be done for several frames ahead''}
 

### 2016-09-09 twostagedo.tex
many small typos corrected.

@@
 % could we change the title to this? What do you think? I wanted to get the modular idea in there.
-\title{A Multi-Modal Model of Object Deformation for Robotics
+\title{A Multi-Modal Model of Object Deformation for Robotic Pushing
 %A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
@@
 When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our approach, which consists of learning and combining multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
-While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model.
-We do so through separately learned models which can be combined in different ways for different purposes.
+\comment{While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model.
+We do so through separately learned models which can be combined in different ways for different purposes.}
 \end{abstract}
@@
 \label{sec:ischeme}
-Rather than using the traditional Newton equations with a numeric integration scheme like Euler \comment{or} Verlet, \comment{used twice to solve the second order differential equiations,} we made the \comment{\sout{velocities proportional to the forces:} force proportional to the velocity:
+\comment{Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations,} we made the \comment{force proportional to the velocity:
 \begin{align}
@@
 with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.
-\comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
+\comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is no inertia in the oscillation.
 When experimenting with different integration steps, force values were interpolated between measurements.
@@
 In the simulation, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.
-\comment{When the simulation of the behaviour of the mesh, which is used to model the sponge, is run, the predicted positions of the vertices may indicate that the mesh penetrates the obstacles.
-This should be interpreted as a collision between the sponge and the obstacle.
+\comment{When the simulation of the mesh deformation runs---used to model the sponge---the predicted positions of the vertices may indicate that the mesh penetrates an obstacle.
+This should be interpreted as a collision between the sponge and an obstacle.
 Since the sponge can not penetrate an obstacle, these overlaps must be eliminated.
@@
 
-In order to model this, we model the finger and table using geometric primitives.
-For the table obstacle, any interpenetration by \comment{the sponge} is solved by pushing vertices \comment{of the mesh of the sponge} back to the border of the table.
+\comment{In order to model this, we model the finger and table using geometric primitives.
+For the table obstacle, any interpenetration by the sponge is solved by pushing vertices of the mesh of the sponge back to the border of the table.
 The springs will then propagate the effect to their neighbours.
-For the finger obstacle, resolution is in two stages: by vertices and then edges.
-A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.
-For edges, the standard algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.
+For the finger obstacle, resolution is in two stages: by vertex and then by edge.}  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.
+For edges, the \comment{standard} algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.
 Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.
@@
   \begin{algorithmicieee}{Closest Point on a Segment to Any Other Point}\label{alg:point_segment}
- \STATE Be $a$, $b$ the ends of the edge $\overline{AB}$, $c$ the external point, and $x$ the closest point to $c$ in the line.
- \STATE Be the vectors $\vec{AB} = b - a$ and $\vec{AC} = c - a$.
+ \STATE Let $a$, $b$ be the ends of the edge $\overline{AB}$, $c$ the external point, and $x$ the closest point to $c$ in the line.
+ \STATE Let the vectors $\vec{AB} = b - a$ and $\vec{AC} = c - a$.
  \STATE The projection of $\vec{AC}$ over $\vec{AB}$ is $\|x\| = \dfrac{\vec{AC} \cdot \vec{AB}}{\|\vec{AB}\|}$
@@
   \begin{algorithmicieee}{Push Edge Out}\label{alg:push_edge}
- \STATE Be $a$, $b$ the ends of the edge $\overline{AB}$, $c$ the centre of the circle.  Use algorithm \ref{alg:point_segment} to find the closest point to the centre in the edge.
+ \STATE Let $a$, $b$ be the ends of the edge $\overline{AB}$, $c$ the centre of the circle.  Use algorithm \ref{alg:point_segment} to find the closest point to the centre in the edge.
  \IF {$\|(closest - c)\| < radius$ (There is an intersection)}
@@
 
-\subsection{Geometric Constraints}
-Even with the addition of the term for preservation of angles, the discretized mass-spring forces can cause triangles of the mesh to overlap each other and be reversed or to be flattened.
-The option of giving additional aid through small geometric subroutines was evaluated as well.
-There are two geometric constraints that the mesh is enforced to maintain:
+\subsection{\comment{Additional} Geometric Constraints}
+\comment{While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function.
+Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.
+It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
+We achieve this through geometric checking subroutines.}  There are two geometric constraints that the mesh is enforced to maintain:
 
 \begin{enumerate}
- \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed beyond half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.
+ \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed more than half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ of this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.
 This is enough for most cases, but triangles in the border of the mesh can still overlap each other.
@@
 
-Given the initial shape of the object and a set of parameters, the mass-spring system allows to make predictions of the deformation of an object for various interactions.
-Given the procedures explained in the previous sections, there are three ways to calculate the subsequent deformations of the shape:
+Given the initial shape of the object and a set of parameters, the mass-spring system allows predictions of the deformation of an object for various interactions.
+Given the procedures explained in previous sections, there are three ways to calculate the subsequent deformations of the shape:
 \begin{enumerate}
@@
  \item \textbf{Forces on vertices.} The forces to be applied on the  vertices around the actuator are approximated by the measured reaction forces.
-Both, the series of positions of the actuator and the reaction forces are necessary.
+Both the series of positions of the actuator and the reaction forces are necessary.
 These forces will be solely responsible for the deformation of the mesh.
-If the deformation is insufficient, the actuator may trespass invalidly the border of the mesh.
+If the deformation is insufficient, the actuator may interpenetrate the border of the mesh.
 The quality of the simulation is much more sensitive to the values of the elasticity-plasticity constants.
 This corresponds to the robot itself pushing the object and feeling the response, while using vision solely to determine the position of the actuator and for evaluation purposes.
- \item \textbf{Both.} The previous two are combined.
+ \item \textbf{Collisions and forces.} The previous two approaches can be combined.
 Visual constraints (collision detection) and forces are used to determine the displacement of nodes around the actuator.
@@
 \end{enumerate}
-The mesh as predicted by one step is fed as current shape for the next step and the cycle is repeated until the simulation is stopped.
-
-The next stage \commentErrase{of} \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.
+The mesh as predicted one step ahead is then fed back as the current shape for the next step and the cycle is repeated until the simulation is stopped.
+
+The next stage \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.
 This problem is addressed in the following section.
 
-\section{Learning to Predict the Deformation}
+\section{\comment{Learning to Predict Deformations}}
 \label{sec:learning}
@@
 \caption{(a) Experimental set up.
-A robotic finger pushes a block of deformable material over a table.
+A robot finger pushes a block of deformable material sitting on a table.
 A camera registers the events of the plane where the main deformations take place.
@@
 Left top: Mesh over real block of material.
-Left bottom: Tipical error of the model vs the ground truth (red) and tipical error of the tracker (blue).
+Left bottom: The typical error of the model vs.
+the ground truth (red) and the typical error of the tracker (blue).
 Right top: tracked ground data.
@@
 The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
-Even though this implementation is inspired in the theory of active contours, it works with a much simpler and faster, but less accurate algorithm as it is explained below, instead of minimizing a global energy.}  The tracking of the deformable target in the ground truth has two stages:
+Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
+This is explained below.}  The tracking of the deformable target in the ground truth has two stages:
 \begin{enumerate}
@@
 The Canny algorithm is used to extract the edges of the image.
-\comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.
-If the average hue of the pixels on an edge fall outside a previously calibrated range, the edge is eliminated.}
+\comment{  Since the object of interest has a distinctive colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.
+If the average hue of the pixels on an edge falls outside a previously calibrated range, the edge is eliminated.}
  \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
@@
 Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].
-If there is no edge within that area, the control point remains were it is.
+If there is no edge within that area, the control point remains where it is.
 This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.
@@
  \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
- \STATE Be $L$ a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
+ \STATE Let $L$ be a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
  \FORALL {$L_i$ in $L$}
@@
 \subsection{Evolutionary Algorithm}
-\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they depend on the geometry of the mesh used to model the object, and there can be several sets of parameters in different regions of the space of parameters that produce adequate simulations \cite{Morris2008}, we chose to use an evolutionary algorithm to expore different regions and try improvements in the vecinity of successful individuals.}
-
-The evolutionary algorithm listed in \alref{alg:genetic} was used to search the space of parameters of the mass-spring model presented in the previous section.
+\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they instead depend on the geometry of the mesh used to model the object.
+There can thus be several sets of parameters in different regions of the parameter space that produce adequate simulations \cite{Morris2008}.
+We therefore chose to use an evolutionary algorithm to both perform global search, and local improvement.}
+
+The evolutionary algorithm described in \alref{alg:genetic} was used to search the parameter space of the mass-spring model presented in the previous section.
 From the parameters that must be chosen for the simulation to work, some were fixed while others had to be searched for.  They are:
@@
   \begin{algorithmicieee}{Evolutionary Search}\label{alg:genetic}
- \STATE Be $N$ the number of sets of parameters to be evaluated per generation and $M$ the number of generations to try.
+ \STATE Let $N$ be the number of sets of parameters to be evaluated per generation and $M$ the number of generations.
  \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements\comment{, with: \\
@@
  which will be chosen as follows:}
- \STATE Be $\sigma_{max} > \sigma_{min}$ standard deviations for a Gaussian
+ \STATE Let $\sigma_{max} > \sigma_{min}$ both be standard deviations of a Gaussian
  \STATE Let $\Delta\sigma \gets (\sigma_{max} - \sigma_{min})/M$.
  \FOR {$i=1 \to M$}
- \STATE Run the $N$ simulations and compare prediction with the ground truth (real object), by comparing the area occupied by the real object with the area occupied by the mesh of the simulation [\fref{fig:scene}(b)].
- \STATE Keep the best $\alpha$ of all simulations.
+ \STATE Run $N$ simulations and compare prediction with ground truth (real object), by comparing the area occupied by the real object with the area occupied by the simulated mesh [\fref{fig:scene}(b)].
+ \STATE Keep the best $\alpha$ from all simulations.
  \STATE $\sigma \gets \sigma - \Delta\sigma$
@@
 
-To evaluate the quality of the predicted shapes, the pixels inside the tracked contour are used as ground truth.
+To evaluate the quality of the predicted deformations, the pixels inside the tracked contour are used as ground truth.
 These are  compared against those inside the mesh of the simulation.
-In machine learning terms, these pixels can be divided among four classes: the \textit{true positives $TP$}, where the material is present, and the simulation predicted it would be present; the \textit{true negatives $TN$}, where the absence of material is in accordance with the prediction; the \textit{false positives $FP$}, where the prediction said there would be material, but there is not; and the \textit{false negatives $FN$}, where there is material, but the prediction said there wouldn't [\fref{fig:scene}(b)].
-Given that maintaining a record of the true negatives (the empty/background space around the material) is mostly irrelevant and highly inefficient, the mark for a whole video was derived from the following quantities:
+In machine learning terms, these pixels can be divided into four classes: the \textit{true positives $TP$}, where the material is present, and the simulation predicted it would be present; the \textit{true negatives $TN$}, where the absence of material is in accordance with the prediction; the \textit{false positives $FP$}, where the prediction said there would be material, but there is not; and the \textit{false negatives $FN$}, where there is material, but the prediction said there wouldn't be [\fref{fig:scene}(b)].
+Given that maintaining a record of the true negatives (the empty/background space around the material) is mostly irrelevant and highly inefficient, the score for a whole video was derived from the following quantities:
 \begin{align}
@@
 
-This measurement \comment{is know as F-score and} is used in the computer vision community to evaluate segmentation algorithms.
+This measurement \comment{is know as the F-score and} is used in the computer vision community to evaluate segmentation algorithms.
 $F=1$ means that there were no false predictions ($FN+FP = 0$).
 $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).
-$F$ is calculated for every frame and the mark for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
+$F$ is calculated for every frame and the score for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
 \comment{\begin{align}
@@
 \end{align}}
-If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closest to zero, the more errors there were.
+If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closer to zero, the more errors there were.
 The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.
@@
 
-\section{Models of Force Learning}
+\section{\comment{Learning to predict forces}}
 %We used two methods to predict the forces.
@@
 
-Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{than the mass-spring model.
-This diagram is obtained from the video recorded by the visual system of the robot and the forces recoded by a sensor located at the wrist.
+Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{as the mass-spring model.
+This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.
 When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.
@@
 \end{align}
-Where $y$ is the dependent variable, $x$ the independent one, $m$ is the sloop of the line, and $y_0$ the $y$ intercept.
+Where $y$ is the dependent variable, $x$ the independent one, $m$ is the slope of the line, and $y_0$ the $y$ intercept.
 To adjust logarithmic curves a change of variable is used.
@@
 
-For the automatic calibration of the pushing phase both types of curves are considered.
-To determine the place of the discontinuities between the segmentes, several candidates are calculated.
-Each candidate is obtained by adjusting a curve through different amounts of data points and calculating the error.
-Beginning with a small number of points (e.g.
-5), as new measurements are taken, new curves are adjusted until the corresponding normalized error increases beyond a threshold value.
-The type of curve that can cover more points with a low fit error is adopted for the model.
+\comment{For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.
+To determine the points of discontinuity between segments, each family is evaluated incrementally.
+The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.
+We begin with a small number of points (e.g.
+five).
+To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.
+This marks the point of discontiuity.
+Having discovered one best parameterized line for each family, the two resulting candidates are compared.
+We select the type of curve that covers more points in a single segment}.
 This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.
@@
 The first set of data was used to train the model for the respective material.
-The set of parameters that allowed the model to behave more like the real object, according to the criterion explained in \sref{sec:evaluation}, was selected.
+The set of parameters that allowed the model to behave most like the real object, according to the criterion explained in \sref{sec:evaluation}, was selected.
 The other two videos were used to test the quality of the predictions made by the model for the new interactions.
 
-To take into account the cylindrical shape of the actuator, the radial component of the registered force is applied on any vertex that happens to be located within the radius of the robotic finger.
+To take into account the cylindrical shape of the actuator, the radial component of the registered force is applied at any vertex that happens to be located within the radius of the robotic finger.
 For collision detection routines the finger is considered a circular obstacle.
@@
 %\includegraphics[width=88mm]{arrio8}
-\caption{The simulation time system with shape and force prediction.
-The position of the actuator is extracted from the image, while the force is predicted by a piece-wise regression model.
-This information is used by the calibrated mass-spring system to predict the shape of the object for several frames.
+\caption{Detail of the simulation system with shape and force prediction.
+The position of the actuator is extracted from the image, while the force is predicted by the piece-wise regression model (FP).
+This information is used by the calibrated mass-spring system (SP) to predict the shape of the object for several frames.
 }\label{fig:diagram}
@@
 Sets of 24 training sessions were run, using different combinations of elasticity and plasticity terms.
-For example: using damped preservation of length only, using preservation of area and angle only or all of them together;  including plastic deformation or removing the collision resolution routines and the aid of geometric constraints.
-For this first rounds of experiments three different integration schemes were used: Euler, Verlet and our Customised version.
+For example, we used damped preservation of length only, preservation of area and angle only, or all of them together.
+Cases were tried including plastic deformation, and removing the collision resolution routines and the aid of geometric constraints.
+For this first round of experiments three different integration schemes were used: Euler, Verlet and our customised version.
 The resulting marks are summarised in \tref{tab:ischemes}.
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Marks of Sets of Parameters Obtained With Different Integration Schemes}
+\caption{Scores of Sets of Parameters Obtained With Different Integration Schemes}
 \label{tab:ischemes}
@@
 The traditional equations of movement, estimated with Euler or Verlet show big oscillations and frequently become unstable.
-Verlet seems to be badly affected by the collision resolution routines and geometric aids, that affect drastically the values of the velocity and acceleration.
-On the contrary, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that resemble more the ground truth, thus obtaining better marks.
-From this group of experiments it was also found that the evolutive learning is as efficient when considering all elastic and plastic terms, as when using only a few of them.
-
-\subsection{Parameters Calibrated Automatically}
-\subsubsection{Performance of the Learning Algorithm for the Mass-spring Model}
+Verlet seems to be badly affected by the collision resolution routines and geometric aids, that drastically affect the values of velocity and acceleration.
+On the other hand, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that more closely resemble the ground truth, thus obtaining better marks.
+From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them.
+
+\subsection{\comment{Results of Automatic Parameter Calibration}}
+\subsubsection{\comment{Performance for the Learned Shape Prediction Model}}
 The performance of the genetic learning algorithm was evaluated using the customised integration scheme and the mass-spring elastic-plastic model.
@@
 \fref{fig:performance} shows how the quality of the simulation throughout the whole video improved with each generation.
-Both graphs show that even though the enforcing of geometric constraints with aid routines may produce simulations with lower marks at the begining of the training, they perform just as well after enough generations.
-These marks reflect the degree at which areas covered by the ground truth and the mesh of the simulation match; however, when looking at the videos produced, it becomes evident that simulations with geometric aids succeed more in keeping the mesh flat.
+Both graphs show that even though the enforcing of geometric constraints with additional routines may produce simulations with lower scores at the begining of the training, they perform just as well after enough generations.
+These scores reflect the degree to which areas covered by the ground truth and the mesh of the simulation match; however, when looking at the videos produced, it becomes evident that simulations with additional geometric constraints succeed more in keeping the mesh flat.
 % Should put images?
@@
 
-\subsubsection{Prediction of the Force}
+\subsubsection{\comment{Performance for the Learned Force Prediction Model}}
 
@@
 See \fref{fig:sstrain}.
-As it was expected from typical stress-strain diagrams of elastic and plastic phases of materials, a linear regression is characteristic of an elastic material, while the logarithmic fits better to a plastic one.
+As was expected from typical stress-strain diagrams of elastic and plastic phases of materials, a linear regression is characteristic of an elastic material, while the logarithmic regression better fits a plastic material.
 For the retrieving movement only a linear model was considered, since the shape deformation of the plastic material makes it harder to measure the strain from the visual information.
 The curves obtained are used to approximately predict the reaction force for unknown interactions.
-These forces are used for the mass-spring model as in the diagram of \fref{fig:diagram}.
+These forces are used for the mass-spring model as in \fref{fig:diagram}.
 
@@
 
-One hundred photographs and force readings where taken every $0.17\pm0.007 s$ in average.
+One hundred photographs and force readings were taken every $0.17\pm0.007 s$ in average.
 The mass spring model runs with an integration step \comment{$h=0.1s$}, where forces that were not measured directly are interpolated.
-The push down movement lasts approximately 50 frames and the retrieval, the other 50.
+The push down movement lasts approximately 50 frames and the retrieval movement the other 50.
 The continuous lines in \fref{fig:sstrain} show the predicted forces, given the amount of compression of the material (strain), while the point-lines show the actual readings.
-\fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the shapes predicted by the mesh.
+\fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the actual shapes predicted \comment{by the combined FP and SP model}.
 
@@
 
-The \comment{\sout{calibrated}} mass-spring models \comment{calibrated for the prediction task} were used to classify \comment{\sout{objects} materials} through recognition of their behaviour, even \comment{\sout{for novel interactions} when the objects where pushed and deformed at novel locations.}  The models make predictions as if the \comment{material of the} object was known.
+The mass-spring models \comment{calibrated for the prediction task} were used to classify \comment{materials} through recognition of their behaviour, even \comment{when the objects where pushed and deformed at novel locations.}  The models make predictions as if the \comment{material of the} object was known.
 The model that makes the best predictions indicates which \comment{type of} object is being observed.
@@
 \begin{enumerate}
- \item Per frame: \comment{\sout{an} a candidate} object is selected for each frame of the interaction \comment{according to which model made the best prediction for that frame in particular, as marked by the F-score.
-This test was performed for every frame of the six videos, where the robotic finger pushes the block of material.
-\tref{tab:classperframe} shows the results for the videos that were used as trainning and test sets for the prediction task.}
+ \item Per frame: \comment{a material is selected for each frame of the interaction according to whichever model made the best prediction for that particular frame, according to the F-score.} This test was performed for every frame of the six videos, where the robot finger pushes the block of material.
+\tref{tab:classperframe} shows the results for the videos that were used as training and test sets for the prediction task.
 
@@
 
- \item Global: \comment{The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted mark.} The difference in the \comment{F-score} marks is added for all frames.
+ \item Global: \comment{The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted score.} The difference in the \comment{F-score} is added for all frames.
 Frames where the distinction is more noticeable will receive a greater weight.
@@
 \subsection{The Learning Method}
-Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to improve the quality of the simulations as much as it could.
-Better simulations have been obtained by hand tunning some of the parameters.
+Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
+Better simulations have been obtained by hand tuning some of the parameters.
 %\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
@@
 
-\subsection{Using Geometric Constraints}
+\subsection{Using \comment{Additional} Geometric Constraints}
 Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.
-Also, when attempting to model the plastic material, the geometric constraints provoqued a much more notorious effect comming from any plasticity term, even if its corresponding constant was small.
+\comment{Also, when attempting to model the plastic material, the geometric constraints provoked a much larger effect caused by the plasticity term}, even if its corresponding constant was small.
 As a consequence the plastic material also recovered its original shape after some extra frames.
-It has been seen that, if the constant for the linear preservation is set to zero, the problem disappears.
+It has been seen that, if the constant for linear preservation is set to zero, the problem disappears.
 Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.
-We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the begining, instant of greatest deformation and final shape.
+We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the beginning, to the instant of greatest deformation and to the final shape.
 Meanwhile, the evaluation function gives the same weight to all frames in the sequence.
-It is possible to use the force learning algorithm and the commands given to the robot (push/retreive) to detect these points, since they are corelated to discontinuities in these other modalities.
+It is possible to use the force learning algorithm and the commands given to the robot (push/retrieve) to detect these points, since they are correlated with discontinuities in these other modalities.
 Future work will evaluate the impact of these modifications on the evaluation function.
@@
 \subsection{Training With Collision Resolution with the Finger vs. Without}
-It was found that the overall mark of the best simulations was similar in both cases, since only geometric similarities are under consideration.
-However, there was a notorious difference on the rate of success for the classification problem.
+It was found that the overall score of the best simulations was similar in both cases, since only geometric similarities are under consideration.
+However, there was a noticeable difference in the rate of success for the classification problem.
 If no collision detection with the finger is used during training, the springs react only to the forces and, if the spring coefficients are not adequate, the shape of the mesh will be very different from the shape of the ground truth.
 
-On the other side, if collision detection is used, the geometric constraints strongly favour a correct configuration of the mesh in the vicinity of the actuator, even though the springs did not react properly.
-For subsequent frames, it is more likely that the neighbours will succeed in propagating the deformation to the rest of the material, than if the positions of the springs had not been tampered by the collision resolution, thus producing better simulations.
-However, this lessens the effectiveness of the mass-spring model in distinguishing the materals.
+On the other hand, if collision detection is used, the geometric constraints strongly favour a correct configuration of the mesh in the vicinity of the actuator, even though the springs did not react properly.
+For subsequent frames, it is more likely, when using collision detection, that the neighbours will succeed in propagating the deformation to the rest of the material, thus producing better simulations than without it.
+However, this lessens the effectiveness of the mass-spring model in distinguishing the materials.
 
@@
 The fact that the spring constants must respond to the applied forces during the training and not only to the visual appearance is reflected in a better differentiation of the behaviour of the models.
-When the calibration is more sensitive to the constants of the springs, it is more robust for novel interactions, therefore the difference in quality of the predictions is a better indicator of which material is beeing used.
+When the calibration is more sensitive to the constants of the springs, it is more robust for novel interactions, therefore the difference in quality of the predictions is a better indicator of which material is being used.
 
 \subsection{Conclusions}
-In summary we have presented a framework, together with specific algorithms for learning the sensorimotor contingencies for the deformable behaviour of objects under robot manipulation.
-We have developed a learning model that learns the parameters of two contingencies that are sequenced, one for kinematic motion to force and the other for force to deformation.
+In summary, we have presented a framework and specific algorithms for learning the sensorimotor contingencies for the deformable behaviour of objects under robot pushing.
+We have developed a learning model that learns the parameters of two contingencies that are sequenced, one for kinematic motion to force (FP) and the other for force to deformation (SP).
 By sequencing these we obtained a compound contingency that predicts deformation given motion.
 The main properties of the model are as follows.
-First, the parameters for the model were adjusted to data from real objects under robotic manipulation.
+First, the parameters for the model components were learned from data from real objects under robotic manipulation.
 Second, we have attempted the automatic calibration of a simulation of plasticity.
-Third we use regression to predict the reaction forces.
-Fourth we include a new energy term for the preservation of angles of the mesh's triangles.
-Fifth we employ a regularization algorithm for a linear snake, that increases or decreases the number of its control points, according to the level of detail required to represent the deformed object.
-Sixth we provide a detailed evaluation of the dynamic simulation, instead of only during its stable states as in previous work.
-Finally we showed the application of the model to the problem of classifying deformable materials.
+Third, we use regression to predict the reaction forces in the force prediction model.
+Fourth, we include a new energy term for the preservation of angles in a mass-spring model of shape deformation.
+Fifth, we employ a regularization algorithm for a linear snake, that increases or decreases the number of its control points, according to the level of detail required to represent the deformed object.
+Sixth, we provide a detailed evaluation of the dynamic simulation, instead of only during its stable states as in previous work.
+Finally we showed the application of the learned prediction models to the problem of classifying deformable materials.
 
@@
 % or if you just want to reserve a space for a photo:
+\newpage
 
@@
 
-%%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{stolk}}]{Rustam Stolkin}
-%%  is currently Senior Research Fellow in robotics, in the School of
-%  Mechanical Engineering, University of Birmingham, UK. Prior to
-%  coming to Birmingham, he was a Research Assistant Professor at
-%  Stevens Institute of Technology, USA. Rustam holds undergraduate and
-%  masters degrees in Engineering Science from University of Oxford,
-%  and a PhD in Robot Vision, undertaken between University College
-%  London and the UK imaging industry. His research interests include
-%  robotics, computer vision, sensor systems, and science and
-%  engineering education
-%\end{IEEEbiography}
-
 % You can push biographies down or up by placing
@@
 \vfill 
-
 % Can be used to pull up biographies so that the bottom of the last one

### 2016-09-09 AnswersToComments.tex
changed letter

@@
 
-\opening{Dear Dr. Ricardo Téllez,}
+\opening{Dear Dr. Ricardo Téllez, Dr Guillem Aleny\`{a},}
 
@@
 
-We thank the reviewer for the comments and we have made our best effort to address them.  In the following text we explain in some detail the changes we did.
+We would like to thank the reviewer for their insightful comments, which have enabled us to significantly improve our paper.
+We have made our best effort to address them.
+In the following pages we explain in some detail the changes we made.
+We have also suggested a slight change to the title make it clear, as the reviewer pointed out, that the paper deals solely with pushing actions and not manipulation in general.
 
-Before that, we also wish to apologise because, in spite of our efforts, we still have not been able to implement the comparison with other existing methods, as any fair comparison requires an extension/implementation to closely related methods which do not work under the same conditions that ours.
-We are working on that, but the deadline has come and we will submit our work leaving only this point pending.
-If you could allow us to submit another version in four weeks, we should be able to include this comparison.
+Before we detail the precise changes made, we also wish to apologise because, in spite of our efforts, we still have not been able to implement the requested comparison with another existing method, as any fair comparison requires an extension to the compared method(s).
+This extension is necessary because none of the methods were designed to work under qualitatively the same conditions as ours.
+Because of this we are still working on an extension of another method, and its implementation.
+We expect to complete it within a month.
+To this end, we wrote to you to request an extension, but we have not yet received a reply.
+Therefore, to conform to the deadline we were given, we are submitting the paper with all the textual revisions completed.
+We would like to submit a further version that includes the results of our comparison.
+We request an extension of four weeks to allow this.
+Please let us know at your convenience whether this is acceptable.
 
-Thank you very much for everything,
+All of the changed parts of the manuscript have been highlighted in red in the re-submitted version.
+We also made some minor changes to the grammar in addition to the requested changes.
+We have not marked these.
+
+We thank you again for your very helpful comments on our paper.
+They have helped us to significantly improve it.
+We hope that we have satisfactorily addressed all your requested changes to the text.
+We will submit a second revision with the experimental comparison as soon as possible.
+We look forward to hearing from you.
+
+\signature{Dra. Verónica Esther Arriola Ríos\\
+Profesora Asociada C de T.C.\newline
+Departamento de Matemáticas, Cub 119.\newline
+Facultad de Ciencias, UNAM \newline
+v.arriola@ciencias.unam.mx \newline
++(52)55 5622 5426}
+
+\closing{Yours sincerely,}
 
@@
 
-\vspace{2cm}
+\newpage
 
-Reviewer(s)' Comments to Author:
-
-Reviewer: 1
-
-Comments to the Author
+Reviewer's Comments \comment{(and responses in blue)}:
 
 This article aims at tracking three-dimensional deformable objects under robot motion and predicting the force causing the deformation.
-
 %\marker
@@
 %
-
 For these purposes, the authors propose two models, including an extension of the mass-spring model and a regression model, where the parameters of the mass-spring model are learned with an evolutionary searching algorithm on visual data and force data, and the parameters of the regression model are learned with least-square methods on stress-strain graph.
@@
 
+\newpage 
 4, section IV-D is not explained: what is the relation between "collision detection" and the proposed model?
@@
 
-\marker
 %\comment{TODO: Move section IV-E after IV-A but must emphasize that the geometric constraints in IV-A are enforced through energy preservation, while those in IV-E have immediate effects on the triangulations without directly affecting the equations.}
-\comment{TODO: Chage title of section E to ``Geometric constraints outside from the mass-spring model''.
-``Even with the addition of the term for preservation of angles to the mass-spring simulation...''}
 
-\comment{Thank you for the observation, it shows the paper was not fully explaining the difference between the constraints presented in section IV-A and those in section IV-E.
-The first belong to the mass-spring model (as everything being explained in subsections A to C).
-The routines in section IV-E are independent and more closely related to mesh-related fixes like the one explained in section IV-D.
-We changed the title of section IV-E to reflect this and further explained this difference in its text.}
+\comment{Thank you for the observation.
+We would very much like to keep the separation of the two sections, as the two constraints are not at all the same, and the additional constraints in section IV-E are both optional and separate from those in the mass-spring model (section IV-A).
+Those belonging to the mass-spring model are soft-constraints implemented through an energy function.
+The additional routines in section IV-E are independent of the mass spring model.
+These are hard constraints, being purely kinematic, and are rather more closely related to the collision detection techniques described in IV-D.
+We changed the title of section IV-E to reflect this and further explained this difference in the text.}
 
@@
 
-\marker
+\newpage
+
 3, in line 58 of page 1, the authors claim "In each case the models can be run in two modes, prediction and classification".
@@
 
-\comment{What does he mean by ``filtering mode instead''?}
+\comment{We hope that we have now addressed this in the model overview section.
+We do not have a filtering mode as such.
+Instead we run models for each material in prediction model, and compare their predictions to the ground truth.
+This is the classification mode.
+This should now be much clearer, especially from the addition of new figures that show the prediction machine (Fig 3.) and the classification machine built on top of it (Fig 4.).}
 
@@
 
-\comment{Since it was not clear the wording of the first paragraph of section IV-A was modified, where the general idea is explained, the details come in later paragraphs.}
+\comment{Since it was not clear the wording of the first paragraph of section IV-A has been modified, where the general idea is explained, the details come in later paragraphs.}
 
@@
 
-\comment{Added an introduction at the beginning of section IV.}
+\comment{We have added an introduction at the beginning of section IV.}
 
@@
 
-\comment{The term was introduced on the second paragraph of section IV.}
+\comment{The term was introduced in the second paragraph of section IV.}
 
@@
 
-\comment{Added explanation that it happens when/because we use $C$s that model springs.}
+\comment{We have added an explanation that it happens when/because we use $C$s that model springs.}
 
@@
 
-\comment{It is not within the scope of the article to explain that derivation, since it belongs to Teschner's model, however the reference to the article we got it from was repeated here.}
+\comment{It is not within the scope of the article to explain that derivation, since it belongs to Teschner's model, and it would take too much space.
+We have, however included a reference to the article for the reader to follow up.}
 
@@
 
-\comment{Changed phrase for the explanation that Teschner gave.}
+\comment{We have changed this phrase to make clearer the explanation that Teschner gave for this.}
 
+\newpage 
 12, in line 50 of page 4, please explain "the angles between adjacent edges". As adjacent edges form one angle, do the authors mean adjacent angles? Similar ambiguity exists in line 2 of page 5.
@@
 
-\comment{Reworded sentence to add more detail.}
+\comment{We have re-worded the sentence to add more detail.}
 
@@
 
-\comment{Explained in more detail and changed notations to make it consistent with previous paragraphs.}
+\comment{We have explained this in more detail and changed the notation to make it consistent with previous paragraphs.}
 
@@
 
-\comment{Improved the description across the whole subsection.}
+\comment{We meant the collisions between both the deformed object and the finger, and between the deformed object and the table.
+We have now improved the description across the whole subsection to make this clear.}
 
@@
 
-\comment{No, it can be found across the web in many sites.  We changed the word ``common'' in the text, for ``standard'', which should emphasise more this fact.}
+\comment{No, it can be found in many places.  We changed the word ``common'' in the text, for ``standard'', which should make this clear.}
 
@@
 
-\comment{Section V-B?
-Added a justification at the beginning of the section.
-The meaning of $\alpha, \beta, \gamma, \zeta$ was covered on lines 7,9,10,11 but a shorter description was added when they are firstly introduced.}
+\comment{We have added this justification at the beginning of the section.
+The meaning of $\alpha, \beta, \gamma, \zeta$ is covered in lines 7,9,10,11 of Algorithm 5.
+We have also added a short description in the main text when they are first introduced.}
 
@@
 
-\comment{Changed paragraph.}
+\comment{We have changed this paragraph to make it clear.}
 
@@
 
-\comment{Reference added.  Relationship with active contour explained.}
+\comment{We have added a reference for this, and explained the relationship with active contours.}
+
+\newpage
 
@@
 
-\comment{Added the name ``F-score'' and an equation to define $\mu(f)$}
+\comment{Thanks. We now use the name ``F-score'' and include an equation to define $\mu(f)$}
 
@@
 
-\comment{Added information.}
+\comment{We have added information on both these points.}
 
-\marker
 23, in the experiments, please compare with other existing methods.
+
+\comment{We are still working on this.
+We requested an extension, as this requires more time than we had available to do the revisions.
+We haven't received a reply to date.
+So we are submitting the paper with the textual changes now.
+If these textual changes are satisfactory we will submit a second revision with the comparison as soon as possible.
+We estimate this will take four weeks from today.
+We are sorry that despite our best efforts this has taken longer that we expected.}
 
@@
 
-\comment{Added explanations}
+\comment{We have added explanations of these values.}
 
@@
 
-\comment{Sorry, I don't understand this one.
+%\comment{Sorry, I don't understand this one.
 There are three lines with markers: square, circle and triangle.
@@
 
-\comment{Yes, it is $h$, the name of the variable was added.}
+\comment{Yes, it is $h$, the name of the variable has been added to make this clear.}
 
@@
 
-\comment{Changed ``for novel interactions'' for a more concrete description.}
+\comment{We have replaced ``for novel interactions'' with a more concrete description.}
 
@@
 Instead of a large collection of labeled images in different backgrounds, we used six videos of two different materials and asked the system to identify which material it looks at through the videos.
-The explanation in the article was modified to be more specific about this.}
+The explanation in the article has been modified to be more specific about this.}
+
+\newpage
 
@@
 
-\comment{An explanation about "training", "test 1, test 2" was added.}
+\comment{An explanation about "training", "test 1, test 2" has been added.}
 
-\comment{Since there are only two materials the number of false positives for one material is exactly the false negatives for the other material, that is why we considered that recall would be redundant.}
+\comment{Since there are only two materials, the number of false positives for one material is exactly the number of false negatives for the other material, that is why we considered that recall would be redundant.}
 
@@
 
-\comment{Added to the background section, alongside the statement ``those predictions are used only for the next frame and can not be used for robotic planning, where predictions must be done for several frames ahead''}
-
-I look forward to hearing from you.
-
-\signature{Dra. Verónica Esther Arriola Ríos\\
-Profesora Asociada C de T.C.\newline
-Departamento de Matemáticas, Cub 119.\newline
-Facultad de Ciencias, UNAM \newline
-v.arriola@ciencias.unam.mx \newline
-+(52)55 5622 5426}
-
-\closing{Yours sincerely,}
+\comment{We have added references on deformable object tracking to the background section.
+We have also made clear why the use o one-step prediction in those algorithms is different from our multi-step prediction problem.
+}
 

### 2016-09-09 twostagedo.tex
highlighting of changes

@@
 \includegraphics[width=3.5in]{figures/learning.png}%}
-\caption{The learning method comprises two learned models of sensorimotor contingencies.
+\caption{\comment{The learning method comprises two learned models of sensorimotor contingencies.
 Both are learned from the same data.
@@
 The strain $st$ is calculated directly from the interpenetration of the finger into the object.
-This can be calculated by vision (V) or from the planned motion of the finger.}
+This can be calculated by vision (V) or from the planned motion of the finger.}}
 \label{fig:learning}
@@
 \includegraphics[width=3.5in]{figures/prediction}%}
-\caption{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way.
+\caption{\comment{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way.
 The predicted shape is fed back on itself so as to predict over many steps for a planned sequence of finger motion.
 PC is the position controller for the robot, which outputs each planned finger position $p$ at each time.
-This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}
+This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}}
 \label{fig:prediction}
@@
 \includegraphics[width=0.8\textwidth]{figures/classification}%}
-\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
+\caption{\comment{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
 The most accurate predictor at any one moment gives the classification for the material type.
-The notation is explained in the text.}
+The notation is explained in the text.}}
 \label{fig:classification}

### 2016-09-09 twostagedo.tex
changed title (again)
bibliography etc

@@
 % could we change the title to this? What do you think? I wanted to get the modular idea in there.
-\title{A Multi-Modal Model of Object Deformation for Robotic Pushing
+\title{A Multi-Modal Model of Object Deformation under Robotic Pushing
 %A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation
@@
 \label{sec:background}
-This work focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger.
-These predictions include the response forces and the deformations of the object's shape over many future steps.
+
+\comment{It is known that the human motor system uses predictive (or forward) models of the effects that motor actions have on sensory state \cite{flanagan03,flanagan06,johansson92}.
+These are believed to be particularly important for dexterous manipulation.
+In robotics, prediction of physical interaction is a well studied but incompletely solved problem.
+There is a large body of work on the effects of pushing on rigid objects.
+This includes physics based models \cite{mason_manipulator_1982,lynch_mechanics_1992,peshkin_motion_1988,mason_mechanics_2001} and learning approaches \cite{montesano08,moldovan12,hermans11,fitzpatrick_learning_2003,ridge2010self,kroemer2014,Stoytchev_affordances_2008,mericli2014, scholz2010combining,kopicki-etal-icra11,kopicki2016learning,belter2014iros}.
+These predictive models have many applications, including visual tracking \cite{morwald2011icra, duff2011icra, pham2015capturing} and push planning \cite{stillman08ijrr,Dogar_2010,lynchmason96,zitoetal-iros12,Cosgun2011}.}
+
+\comment{There is much less work in robotics on predicting the behaviour of deformable objects under manipulation.
+This paper focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger.} These predictions include the response forces and the deformations of the object's shape over many future steps.
 The predictive models can be used to classify materials, even when the interactions (e.g.
@@
 string, hair, metallic sheet and textile), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.
+We now review the most closely related work to that presented here.
 
@@
 
-Other interesting work includes that of Conti et al.\ \cite{Conti2003} which introduced six-degree of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects.
-The open source project Chai3D includes an implementation of this model.
-The spheres are placed along the skeleton of the object and are connected together with elastic links which model elongation, flexion and torsion properties.
-Each vertex of the surface mesh is attached to the nearest sphere or link with a damped spring.
-In \cite{Burion2008}, Burion and Baur, collaborated with Conti's group to automatically calibrate this model using particle filters.
-Again, there is no analysis of the dynamics and no real objects are used.
-Furthermore, this model is more adequate for objects that deform around fixed joints.
-
-Finally, Cretu et al., used growing neural gas networks to learn to predictively track the deformation of objects.
-Even though they managed to make predictions about the deformation of the overall shape of the object, under previously unseen sets of forces, their predictions were evaluated only for the next frame.
-
-Additionally, there is a vast literature on modeling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.
-However, these types of models have not been successfully applied to robotics tasks beyond what is presented above.
-
-%\subsection{Predicting the Shape}
-%Contribution: we analyse the dynamic behaviour of the predictions for several frames.
-%\subsection{Predicting the Force}
-%Physics based models are used to predict reaction forces (surgery simulators).
-%Contribution: we learn to predict with a linear regression model.  Also calibrated automatically.
-
-\section{Model Overview}
-\label{sec:model_overview}
-
 \begin{figure}[!t]
@@
 \end{figure}
+
+Other interesting work includes that of Conti et al.\ \cite{Conti2003} which introduced six-degree of freedom macroscopic elastic spheres described by mass, inertial and volumetric properties that are used to approximate the volume of deformable objects.
+The open source project Chai3D includes an implementation of this model.
+The spheres are placed along the skeleton of the object and are connected together with elastic links which model elongation, flexion and torsion properties.
+Each vertex of the surface mesh is attached to the nearest sphere or link with a damped spring.
+In \cite{Burion2008}, Burion and Baur, collaborated with Conti's group to automatically calibrate this model using particle filters.
+Again, there is no analysis of the dynamics and no real objects are used.
+Furthermore, this model is more adequate for objects that deform around fixed joints.
+
+Finally, Cretu et al., used growing neural gas networks to learn to predictively track the deformation of objects.
+Even though they managed to make predictions about the deformation of the overall shape of the object, under previously unseen sets of forces, their predictions were evaluated only for the next frame.
+
+Additionally, there is a vast literature on modeling of deformable shapes and deformation processes \cite{Gibson1997, McInerney1996, Montagnat2001, Moore2007review, Nealen2006review}, particularly within the area of computer graphics.
+However, these types of models have not been successfully applied to robotics tasks beyond what is presented above.
+
+%\subsection{Predicting the Shape}
+%Contribution: we analyse the dynamic behaviour of the predictions for several frames.
+%\subsection{Predicting the Force}
+%Physics based models are used to predict reaction forces (surgery simulators).
+%Contribution: we learn to predict with a linear regression model.  Also calibrated automatically.
+
+\section{Model Overview}
+\label{sec:model_overview}
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{twostagedo}
+\bibliography{twostagedo,jlw}
 %

### 2016-09-10 twostagedo.tex
fixed references

@@
 \usepackage{tikz}
-
+\usepackage{balance}
 \usepackage{textcomp}
@@
 There is a large body of work on the effects of pushing on rigid objects.
-This includes physics based models \cite{mason_manipulator_1982,lynch_mechanics_1992,peshkin_motion_1988,mason_mechanics_2001} and learning approaches \cite{montesano08,moldovan12,hermans11,fitzpatrick_learning_2003,ridge2010self,kroemer2014,Stoytchev_affordances_2008,mericli2014, scholz2010combining,kopicki-etal-icra11,kopicki2016learning,belter2014iros}.
+This includes physics based models \cite{lynch_mechanics_1992,peshkin_motion_1988,mason_mechanics_2001} and learning approaches \cite{montesano08,moldovan12,fitzpatrick_learning_2003,kroemer2014,Stoytchev_affordances_2008,mericli2014, scholz2010combining,kopicki-etal-icra11,kopickiwyatt16,belter2014iros}.
 These predictive models have many applications, including visual tracking \cite{morwald2011icra, duff2011icra, pham2015capturing} and push planning \cite{stillman08ijrr,Dogar_2010,lynchmason96,zitoetal-iros12,Cosgun2011}.}
@@
 \bibliography{twostagedo,jlw}
+
+\balance
 %
@@
 % or if you just want to reserve a space for a photo:
-\newpage
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{arrio}}]{Veronica E. Arriola-Rios}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.2in,clip,keepaspectratio]{arrio}}]{Veronica E. Arriola-Rios}
   has recently obtained her PhD in Computer Science from the University of Birmingham.
@@
 
-% if you will not have a photo at all:
-%\begin{IEEEbiographynophoto}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{wyatt}}]{Jeremy L. Wyatt} is Professor of Robotics
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{wyatt}}]{Jeremy L. Wyatt}
-%   is Professor of Robotics and Artificial Intelligence at the University of Birmingham. He
-%   obtained a BA in Theology from the University of Bristol, an MSc in
-%   Knowledge Based Systems from the University of Sussex, and his
-%   Ph.D. in Artificial Intelligence from the University of Edinburgh in
-%   1997. He has published more than 80 refereed articles and edited three books. His interests
-%   include machine learning, robot manipulation, robot learning, robot vision, 
-%   robotic architectures, and planning and learning for robots in uncertain 
-%   and incomplete worlds.
-  is Professor of Robotics
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.2in,clip,keepaspectratio]{wyatt}}]{Jeremy L. Wyatt} is Professor of Robotics
   and Artificial Intelligence at the University of Birmingham. He
@@
 
+\vfill
 % insert where needed to balance the two columns on the last page with
 % biographies
-%\newpage
 
@@
 
-\vfill 
 % Can be used to pull up biographies so that the bottom of the last one

### 2016-12-23 twostagedo-rev-Dec-16.tex
my attempts at addressing the issues raised by reviewers (not all), plus response in pages.

--- 2016-09-10 twostagedo.tex
@@
 \begin{abstract}
-\comment{In this paper we present a multi-modal framework for learning generative models of object deformation under robotic pushing.
+In this paper we present a multi-modal framework for \comment{offline} learning of generative models of object deformation under robotic pushing.
 The model is multi-modal in that it is based on integrating force and visual information.
@@
 These component models can be sequenced to provide many-step prediction and classification.
-When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.} Our approach, which consists of learning and combining multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
-\comment{While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model.
-We do so through separately learned models which can be combined in different ways for different purposes.}
+When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.
+Our approach, which consists of \comment{offline learning and combination of} multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
+While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model.
+We do so through separately learned models which can be combined in different ways for different purposes.
 \end{abstract}
@@
 Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable.
-\comment{The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required.
+The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required.
 A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations.
@@
 The predictions of a learned model can be fed back into itself, so as to predict an object's deformation behaviour over many time steps.
-These learned models can then be used to make the predictions required for planning manipulation.}
-
-\comment{There are various models of deformation used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
+These learned models can then be used to make the predictions required for planning manipulation.
+
+There are various models of deformation used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time.
 In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data.
@@
 This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points.
-Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.}
+Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.
 
@@
 
-\comment{Previous work on learning sensorimotor contingencies for deformable objects in robotics has been limited.
+Previous work on learning sensorimotor contingencies for deformable objects in robotics has been limited.
 Table~\ref{t:advances} summarises the features of the major efforts to date.
@@
 
-There are several novel contributions.
+\comment{There are two novel technical contributions.
 First, we show how the learning problem can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces.
 Both models are generative, and we show how to learn each from data.
-Third, we show how to use these learned generative models not only to predict deformation, but also to classify an unidentified material.}
+Second, we show how to use these learned generative models not only to predict deformation, but also to classify an unidentified material.
+To limit the scope of the paper we develop this framework using offline learning methods.
+We do not address the issue of online learning \cite{worgotter09,antonelli14}, leaving this for future work.}
 
@@
 
-In summary, our approach enables us to learn models of sensorimotor contingencies, and then to use these as components to build different machines.
+\comment{These technical contributions} enable us to learn models of sensorimotor contingencies, and then to use these as components to build different machines.
 This approach enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
-In the rest of the paper we present related work (Section~\ref{sec:background}); \comment{give an overview of the framework (Section~\ref{sec:model_overview});} describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).
+In the rest of the paper we present related work (Section~\ref{sec:background}); give an overview of the framework (Section~\ref{sec:model_overview}); describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).
 
@@
 
-\comment{It is known that the human motor system uses predictive (or forward) models of the effects that motor actions have on sensory state \cite{flanagan03,flanagan06,johansson92}.
+It is known that the human motor system uses predictive (or forward) models of the effects that motor actions have on sensory state \cite{flanagan03,flanagan06,johansson92}.
 These are believed to be particularly important for dexterous manipulation.
@@
 This includes physics based models \cite{lynch_mechanics_1992,peshkin_motion_1988,mason_mechanics_2001} and learning approaches \cite{montesano08,moldovan12,fitzpatrick_learning_2003,kroemer2014,Stoytchev_affordances_2008,mericli2014, scholz2010combining,kopicki-etal-icra11,kopickiwyatt16,belter2014iros}.
-These predictive models have many applications, including visual tracking \cite{morwald2011icra, duff2011icra, pham2015capturing} and push planning \cite{stillman08ijrr,Dogar_2010,lynchmason96,zitoetal-iros12,Cosgun2011}.}
-
-\comment{There is much less work in robotics on predicting the behaviour of deformable objects under manipulation.
-This paper focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger.} These predictions include the response forces and the deformations of the object's shape over many future steps.
+These predictive models have many applications, including visual tracking \cite{morwald2011icra, duff2011icra, pham2015capturing} and push planning \cite{stillman08ijrr,Dogar_2010,lynchmason96,zitoetal-iros12,Cosgun2011}.
+
+There is much less work in robotics on predicting the behaviour of deformable objects under manipulation.
+This paper focuses on learning to predict what happens when an elastic or plastic object is pushed by a robot finger.
+These predictions include the response forces and the deformations of the object's shape over many future steps.
 The predictive models can be used to classify materials, even when the interactions (e.g.
 the contact point) are novel.
-Although there \comment{is a large body of work on deformable object tracking, some of which makes use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are typically used only for the next frame and can not be used for robotic planning, where predictions must be made many steps into the future \cite{Cretu2012}.
+Although there is a large body of work on deformable object tracking, some of which makes use of forces, models and predictions to improve tracking accuracy \cite{Chan1994, Greminger2008, Alkkiomaki2009}, those predictions are typically used only for the next frame and can not be used for robotic planning, where predictions must be made many steps into the future \cite{Cretu2012}.
 Others like \cite{Malassiotis1998, Cremers2006} make use of templates or make assumptions about the shape of the objects that would not work with highly deformable objects like plasticine.
-There is also} an abundant literature in industrial robotics for modelling 1D and 2D materials (e.g.
+There is also an abundant literature in industrial robotics for modelling 1D and 2D materials (e.g.
 string, hair, metallic sheet and textile), but there is little research about predicting the deformation of 3D deformable objects \cite{Khalil2010review}.
 We now review the most closely related work to that presented here.
-
-First, Howard and Bekey \cite{Howard2000} address robotic grasping and lifting of viscoelastic objects.
-Their model of the material can predict a net amount of deformation given the applied forces, but does not evaluate details of the deformed shape.
-This model is inspired by the crystalline structure of atoms in solids.
-The space lattice of the crystal is approximated by a particle system where elements with mass are connected by springs and dampers, thus representing the attraction and repulsion between atoms \emph{in equilibrium} with a damped mass-spring system.
-The masses are calculated by measuring the force required to lift the object without sliding.
-The deformation and damping coefficients are estimated as functions of:
-\begin{itemize}
- \item the force applied while pushing against the material with two robot arms in opposite directions and,
- \item the global displacement and velocity vectors resulting from adding the displacements and velocities of all particles.
-\end{itemize}
-The estimated values of the mass, deformability and damping coefficients are used as the input to a neural network that outputs the minimum necessary force to lift the object without sliding.
-This function is used to compliantly adjust the force during an interaction.
-However, they do not evaluate the quality of the overall shape prediction, and use only the final values of the prediction after the system has stabilised; that is, they do not evaluate the predictions during the dynamic phase.
-
-%\begin{figure}[!t]
-%\centering
-%\setlength\fboxsep{0pt}    % padding
-%\setlength\fboxrule{0.5pt} % line width
-%\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
-%\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
-%\label{fig:system}
-%\end{figure}
-
-%%%
-%%% TODO: Text
-%%%
-% \begin{figure}[!t]
-% \centering
-% \setlength\fboxsep{0pt}    % padding
-% \setlength\fboxrule{0.5pt} % line width
-% \fbox{\includegraphics{arrio3}}
-% \caption{The system learns to predict the object's shape and reaction forces from real world data.}
-% \label{fig:system_old}
-% \end{figure}
-
-Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, in which constraints are expressed as potential energy terms.
-These terms encode the tendency of springs to recover their length (with damping), of triangular faces of a mesh to recover their area and of tetrahedrons to recover their volume.
-He sketches the inclusion of plastic deformation.
-Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a FEM model.
-The  search begins with a uniform random sampling of possible values for the elasticity constants, which are latter modified using adaptive simulated annealing to minimize the difference between the deformations in the FEM mesh and those in the mass-spring mesh.
-However, they only compare the steady states and do not include plastic deformations.
-We extend this work, by replacing the FEM model with force and 2D visual data gathered from real materials.
-Furthermore, we analyse the quality of the predictions while the deformations take place.
 
@@
 
+First, Howard and Bekey \cite{Howard2000} address robotic grasping and lifting of viscoelastic objects.
+Their model of the material can predict a net amount of deformation given the applied forces, but does not evaluate details of the deformed shape.
+This model is inspired by the crystalline structure of atoms in solids.
+The space lattice of the crystal is approximated by a particle system where elements with mass are connected by springs and dampers, thus representing the attraction and repulsion between atoms \emph{in equilibrium} with a damped mass-spring system.
+The masses are calculated by measuring the force required to lift the object without sliding.
+The deformation and damping coefficients are estimated as functions of:
+\begin{itemize}
+ \item the force applied while pushing against the material with two robot arms in opposite directions and,
+ \item the global displacement and velocity vectors resulting from adding the displacements and velocities of all particles.
+\end{itemize}
+The estimated values of the mass, deformability and damping coefficients are used as the input to a neural network that outputs the minimum necessary force to lift the object without sliding.
+This function is used to compliantly adjust the force during an interaction.
+However, they do not evaluate the quality of the overall shape prediction, and use only the final values of the prediction after the system has stabilised; that is, they do not evaluate the predictions during the dynamic phase.
+
+%\begin{figure}[!t]
+%\centering
+%\setlength\fboxsep{0pt}    % padding
+%\setlength\fboxrule{0.5pt} % line width
+%\fbox{\includegraphics[width=3.5in]{arrio3.pdf}}
+%\caption{During training time, the system learns to predict the object's shape and reaction forces from real world data.}
+%\label{fig:system}
+%\end{figure}
+
+%%%
+%%% TODO: Text
+%%%
+% \begin{figure}[!t]
+% \centering
+% \setlength\fboxsep{0pt}    % padding
+% \setlength\fboxrule{0.5pt} % line width
+% \fbox{\includegraphics{arrio3}}
+% \caption{The system learns to predict the object's shape and reaction forces from real world data.}
+% \label{fig:system_old}
+% \end{figure}
+
+Teschner et al.\ \cite{Teschner2004} proposed an enhanced 3D mass-spring model, in which constraints are expressed as potential energy terms.
+These terms encode the tendency of springs to recover their length (with damping), of triangular faces of a mesh to recover their area and of tetrahedrons to recover their volume.
+He sketches the inclusion of plastic deformation.
+Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a \comment{finite element model (FEM)}.
+The  search begins with a uniform random sampling of possible values for the elasticity constants, which are latter modified using adaptive simulated annealing to minimize the difference between the deformations in the FEM mesh and those in the mass-spring mesh.
+However, they only compare the steady states and do not include plastic deformations.
+We extend this work, by replacing the FEM model with force and 2D visual data gathered from real materials.
+Furthermore, we analyse the quality of the predictions while the deformations take place.
+
 Frank et al.\ \cite{Frank2010} use a force sensor and a bumblebee stereo camera to measure the Young modulus and the Poisson ratio of unknown deformable objects.
@@
 \includegraphics[width=3.5in]{figures/learning.png}%}
-\caption{\comment{The learning method comprises two learned models of sensorimotor contingencies.
-Both are learned from the same data.
-The first (FP) learns to predicts the force experienced for some motion and material.
+\caption{\comment{The visual tracking and force measurement system subserves the offline learning system.
+The offline learning (or calibration) method comprises two learned models of sensorimotor contingencies.} Both are learned from the same data.
+The first (FP) learns to predict the force experienced for some motion and material.
 The second (SP) predicts the deformation of the object given the applied force.
 The strain $st$ is calculated directly from the interpenetration of the finger into the object.
-This can be calculated by vision (V) or from the planned motion of the finger.}}
+This can be calculated by vision (V) or from the planned motion of the finger.}
 \label{fig:learning}
@@
 
-\comment{The framework consists of three separate systems: a learning system, a prediction system, and a classification system.
+\comment{The framework consists of four separate systems: a visual tracking and force measurement system, an offline learning (or model calibration) system, a prediction system, and a classification system.
+
+Visual tracking and force measurement both serve the learning system (Figure~\ref{fig:learning}).
+The visual analysis is not a contribution of the paper, any method that provides accurate tracking of the outline of the deformed object would suffice.
+We employ a Canny edge detector, filtering out irrelevant edges using the colour of the manipulated object.
+Then we fit the outline and track using a linear spline.
+This is also used to initialise the mesh for the spring mass system.}
+
 The learning system is depicted in Figure~\ref{fig:learning}.
@@
 Learning of the shape prediction (SP) model relies on comparing the predicted deformation $\hat{sh}$ with the actual deformation $sh$ measured by visual tracking (V).
-}
-
-\comment{The two learned models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{fig:prediction}).
+
+\comment{In machine learning there is division between discriminative and generative models.
+Discriminative models are powerful for classification, but cannot be used to regenerate the data, or to generate new imagined data.
+In our application this would mean we could not predict deformations, only being able to classify materials directly based on their deformations.
+Instead, following others, we use a generative model.
+But we show how this can be used not just for prediction, but also for material classification.
+Thus a generative model is appealing because it allows us to solve both the prediction and classification problems.}
+
+The two learned generative models, FP and SP, can be sequenced to create a single, multi-modal predictor (Figure~\ref{fig:prediction}).
 This combined model takes a planned motion of the finger, and predicts the resulting forces that will be experienced at the contact points.
@@
 To create the prediction for time $t+1$ the previous predicted state at time $t$ is required: thus the predictions are recursive.
-This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.}
+This enables predictions over arbitrarily long time periods, but makes learning a good predictor challenging.
 
@@
 \includegraphics[width=3.5in]{figures/prediction}%}
-\caption{\comment{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way.
+\caption{The two learned models can be sequenced, so as to predict an object's shape and reaction forces when the robot finger will push in a particular way.
 The predicted shape is fed back on itself so as to predict over many steps for a planned sequence of finger motion.
 PC is the position controller for the robot, which outputs each planned finger position $p$ at each time.
-This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}}
+This is used to calculate the strain $st$ from the interpenetration using a fixed function, as well as providing an input to the shape predictor (SP).}
 \label{fig:prediction}
@@
 
-\comment{The third system is for material classification (Figure~\ref{fig:classification}).
+The third system is for material classification (Figure~\ref{fig:classification}).
 This simply comprises multiple copies of the prediction systems, one for each type of material that an unknown object might be made of.
 The predictors each make many-step predictions.
-At each step a classification is made, based on which predictor has made the better prediction of the shape at that step.}
-
-\comment{In summary, this way of learning and using models involves several choices.
+At each step a classification is made, based on which predictor has made the better prediction of the shape at that step.
+
+In summary, this way of learning and using models involves several choices.
 First, the learned models are generative.
@@
 The specific model forms we choose here could be replaced while retaining the overall scheme.
-In this paper, we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.}
+In this paper, we implement the scheme using a regression approach to force prediction, and we follow several other authors in employing a mass-spring system with learnable parameters for the deformation prediction.
 
@@
 \includegraphics[width=0.8\textwidth]{figures/classification}%}
-\caption{\comment{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
+\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
 The most accurate predictor at any one moment gives the classification for the material type.
-The notation is explained in the text.}}
+The notation is explained in the text.}
 \label{fig:classification}
@@
 
-\comment{
 A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.
@@
 Another option could be to induce response forces by displacing vertices from their position of equilibrium.
-}
 

### 2016-12-23 response-to-reviewers-second-resubmission.txt
further small improvements

Response to reviewers’ comments

Reviewer 1

In this revision, the authors solved the initial problems on article structure, citations, figures, and contents.

We thank the reviewer for their confirmation that we have satisfied all their comments, and would like to thank them again for their remarks, which significantly improved our paper.

Reviewer 2

Page 1. Last Paragraph. There is an enumeration of novel contributions, you mentioned the First, and the Third, to avoid any confusion I would recommend to remark which is the second.

We apologise, this was a typo. We have listed two main technical contributions.

Page 1-2. It is possible to infer the meaning of FEM, but in order to avoid any possible confusion you might indicate its meaning explicitly when the concept is first used.

We have fixed this.

I find Fig 2, 3, and 4 difficult to read, however I don’t consider they should be necessarily changed.

We agree this was difficult to interpret. To aid the reader we have added some text descriptions for the most important elements to make the diagram more readable.

At the end of subsection VII. B, when the authors refer to “all elastic and plastic terms” might be illustrative to add references to the corresponding equations.

Vero?

Within the Model Overview, it would be interesting if the authors would briefly highlight the advantage of having generative models for those who are not vastly familiarized with the topic.

Thank you for this suggestion. We have added a short explanation within the section Model Overview. This explains that generative models are appealing because they allow us to solve both prediction and classification problems, whereas discriminative models are typically only suitable for classification.

When interpreting results what are the consequences when triangles in the border of the mesh overlap each other? Is there any consequence that might be significant to the result interpretation?

Vero?

In section V, during the training the segmentation assumes that the color of the object is known. Therefore, when you say that the algorithm could learn and classify previously unknown materials, at least the machine must partially know the material. Also the authors mention that the simplified active contours approach is not the most accurate, What are the consequences of this approximation? Regarding to the chosen integration
time step, does the results vary considerably when 0.01 is considered?

Vero?

In section VII.E,  Table III,  Is there any special reason to have the
same result for the training and test1 sets  in the case of the sponge? It
would be better if the authors offer  deeper insights regarding the
interpretation of the figures shown in Table III.
Vero?

Reviewer 3

Reading the abstract and the introduction, I was expecting a framework in which the robot creates a representation of the environment (in this case the two objects) while it is interacting with it (looking at and pushing the object).

We are sorry that this set false expectations. We have altered the wording in several places to make the reader aware that we are not tackling this problem on-line, but off-line. We have also read and noted the references you mention in the introduction.

While the authors recognize the issue of the on-line learning in the introduction (page 1, col 2, line 35), they do not address that in the paper.

From your comments, we believe that mentioning on-line learning, while important, was misleading in terms of our paper, so we have withdrawn mention of it.

Also they refer the problem of getting access to partial views of the objects but it seems that in this paper the object is viewed from a convenient point of view.

We have added text to make clear the restrictions in our study.

Finally, the word “learning from data” suggests to me that the model acquires a representation of the environment based on some kind on non-parametric model, such as neural network. However, the models seem carefully designed to represent the object deformation and the stress-gain diagram. Even if the difference between learning and calibration is subtle, in my opinion, the use of the word “learning” creates an expectation that is not fulfilled in the rest of the paper.

We have changed the wording in the abstract to note that we calibrate the models from data. We hope this makes it clearer where on the spectrum of parameter tuning to full autonomous learning we lie. We now emphasise that we perform off-line learning, we sometimes refer to it as calibration, and we mention that we don’t address the issues of on-line learning. We hope this changes the tone of the introduction enough so as not to mislead.

What happen if we add a new material? Do we need to add a new feature space?

If the object is plastic or elastic then we don’t need to add a new feature space, but we would be wise to calibrate a new model. We have added a comment to this effect in the conclusion.

I would add a figure like Fig. 9a at this point. In this way the reader can have an idea of the setup, how vision acquire information and how the mass-spring model depends on the visual information itself. For example, it might be interesting to combine Fig 9a and Fig 2 in single sketch of the model.

Thank you for this suggestion. We have prepared a new version of Figure 2, as you suggest This incorporates elements from the old Figure 2 and Figure 9.

A consideration about the organisation of the following sections. Given that the framework consists of two models, Force prediction and Shape
prediction, I would expect section III.1 to be Force prediction and III.2
to the shape prediction (or vice-versa). However, the organization of
Sections IV, V, VI, in my opinion is misleading.

Finally, I would start describing the algorithm from the visual system. It seems that the visual system has a very important role in this paper but it is not highlighted. Indeed, the mass-spring model described is section IV should be created within the perimeter of the segmented object. Thus, it requires a visual system capable of localizing and segmenting the object.
In my understanding, this information is necessary to initialize the position of the particles (described in section IV) inside the object. Regarding this point, I would like to know how the particles are placed inside the object. What happen if we change the size of the object? Do we
need to change the number of particles? What happen if the object is not rectangular?

IV. A MASS-SPRING MODEL OF DEFORMATION

My main concern about this section is Eq. 19. I really do not understand
how the force could be proportional to the velocity. This point should be
clarified. My understanding is that the authors assumed the acceleration
(a) to be constant during the interval h, so that h*F = h*a*m => F =
v*m/h.
Given that h is a constant, the force (F), the velocity (v) or the mass
(m)
are recovered with a scale factor. Is that correct? If so, what is the
difference between this approach and the quasi-static approach proposed by
Frank. Please, explain.

Other considerations about this section. This section is quite long and I
think it could be simplified notably. The authors could simply state the
energy E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2], where N is the number of
constraints. There are several constraints to preserve the length, the
area, etc. and the force is given as negative gradient of the potential
energy with respect to the position of the particle. Details about all the
equation could be reported in an appendix to make the paper more readable.
About the equation, personally I don’t like the symbol composed of
multiple letters such as Area, forcemag, forcedir, etc.

VII. EXPERIMENTAL RESULTS

Do the authors use just one video for the training and two video for the
testing? It seems that the dataset is quite small to have a systematic
test of the algorithm. Even if there are only two objects, the authors should
test different motions of the finger (i.e. pushing velocity, penetration,
…)

I don’t see the results of the classification. I would expect a confusion
matrix to show how many times the sponge was classified as sponge and as
plasticine and vice-versa. Table III seems to show the results about the
classification of the pixels.

Moreover, I would like to see a comparison against a baseline approach.
They could some of the models proposed in the literature or some simple
models. For example, the authors could assume that the material
“disappears” when the finger penetrates the object and report the F-Score
of this method. Alternatively, they could use a simple model with only one
spring between the finger and the base of the object.

VIII.DISCUSSION

Page 12. Col 2. Line 3. Please, can you clarify the hand tuning? It seems
that the genetic algorithm is not the best way to find the parameters?
Have you tried to use other method like gradient descent? Is it possible to use
the model to mass-spring model to predict the force exerted on the finger?
In that case, the authors could use this prediction error as fitness
function and use the gradient descent to find the parameters.

Other comments:

Fig. 1 is not referenced in the paper.
Page 2, Col 1, Line 17. Section VI is not mentioned.
Page 2. Col 2, Line 5: FEM is not defined. Finite-element models?
Fig 2. Instead of use symbols, I would write Vision, strain, force, etc.
Page 3, Col 1, line 57. Cretu et al. -> cite [10]?
Page 8. Col 2, Line 27. Linear snake = linear spline?
Algorithm 5. I would specify that the F-Score is the fitness function of
the algorithm. Moreover, it would be nice to know the values used for the
parameters of the algorithm (perhaps using a table).
Page 10. Col 1. Line 45. Please specify that there are two objects, a
sponge and plasticine, with a rectangular shape and so on.

### 2016-12-23 twostagedo-rev-Dec-16.tex
further small improvements

@@
 The model is multi-modal in that it is based on integrating force and visual information.
-The framework consists of several sub-models that are independently learned from the same data.
+The framework consists of several sub-models that are independently \comment{calibrated} from the same data.
 These component models can be sequenced to provide many-step prediction and classification.
@@
 In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data.
-In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time.
-This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points.
-Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.
+\comment{In contrast, robots must learn model parameters from data, in reasonable amounts of training time.
+This includes learning from the limited sensing typically available, such as training data gathered from a specific view(s) of an object, and force-torque monitoring only at contact points.} Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.
+\comment{In this paper we address some of these---we learn from real visual and force-torque data---while simplifying others---we set a convenient fixed viewing location and perform off-line rather than on-line learning.}
 
@@
 We note the following points about the results and the method in general.
-\subsection{The Learning Method}
+\subsection{The Learning/Calibration Method}
 Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
@@
 \subsection{Conclusions}
-In summary, we have presented a framework and specific algorithms for learning the sensorimotor contingencies for the deformable behaviour of objects under robot pushing.
-We have developed a learning model that learns the parameters of two contingencies that are sequenced, one for kinematic motion to force (FP) and the other for force to deformation (SP).
+In summary, we have presented a framework and specific algorithms for off-line learning of the sensorimotor contingencies for the deformable behaviour of objects under robot pushing.
+We have developed an approach that learns/calibrates the parameters of two contingencies that are sequenced, one for kinematic motion to force (FP) and the other for force to deformation (SP).
 By sequencing these we obtained a compound contingency that predicts deformation given motion.
@@
 Second, we have attempted the automatic calibration of a simulation of plasticity.
-Third, we use regression to predict the reaction forces in the force prediction model.
+\comment{If a new material were added, provided it were elastic or plastic, it would be possible to calibrate a model for that material, independent of the existing calibrated models.} Third, we use regression to predict the reaction forces in the force prediction model.
 Fourth, we include a new energy term for the preservation of angles in a mass-spring model of shape deformation.
@@
 While the framework presented here is promising, we regard the actual deformation model representation as an open research problem.
+\comment{Finally, all the work here is for off-line learning, from a carefully chosen viewpoint.
+The issue of online learning of these representations, including how to explore so as to learn, is an open issue.}
 

### 2016-12-24 twostagedo-rev-Dec-16.tex
lots of small changes to satisfy reviewer 3.

@@
 
-\begin{figure}[!t]
-\centering\includegraphics[width=3.5in]{arrio17.png}
-\caption{Force information acquired while working with a block of sponge (elastic) and a piece of plasticine (plastic).
+%\begin{figure}[!t]
+%\centering\includegraphics[width=3.5in]{arrio17.png}
+%\caption{Force information acquired while working with a block of sponge (elastic) and a piece of plasticine (plastic).
 Notice the difference in magnitude between the force used with the sponge and the plasticine.}
-\label{fig:forceElasticPlastic}
-\end{figure}
+%\label{fig:forceElasticPlastic}
+%\end{figure}
 
@@
 This approach enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
-In the rest of the paper we present related work (Section~\ref{sec:background}); give an overview of the framework (Section~\ref{sec:model_overview}); describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).
+In the rest of the paper we present related work (Section~\ref{sec:background}); give an overview of the framework (Section~\ref{sec:model_overview}); describe the two component models in detail (Section~\ref{sec:model}); describe how the model parameters are learned from real data for the shape prediction model (Section~\ref{sec:learning}) and for the force prediction model (Section~\ref{sec:forces}); report experiments on prediction and classification with elastic and plastic objects (Section~\ref{sec:experiments}); and finish with a short discussion (Section~\ref{sec:discussion}).
 
@@
 
-Finally, Cretu et al., used growing neural gas networks to learn to predictively track the deformation of objects.
+Finally, Cretu et al.
+\cite{Cretu2012}, used growing neural gas networks to learn to predictively track the deformation of objects.
 Even though they managed to make predictions about the deformation of the overall shape of the object, under previously unseen sets of forces, their predictions were evaluated only for the next frame.

### 2016-12-24 response-to-reviewers-second-resubmission.txt
copies of responses

@@
 
-We have added text to make clear the restrictions in our study.
+We have added text to make clear the restriction in our study of using a single view chosen by the user.
 
@@
 
-Thank you for this suggestion. We have prepared a new version of Figure 2, as you suggest This incorporates elements from the old Figure 2 and Figure 9.
+Thank you for this suggestion. We have prepared a new version of Figure 2, as you suggest. This incorporates elements from the old Figure 2 and Figure 9.
 
-A consideration about the organisation of the following sections. Given that the framework consists of two models, Force prediction and Shape
-prediction, I would expect section III.1 to be Force prediction and III.2
-to the shape prediction (or vice-versa). However, the organization of
-Sections IV, V, VI, in my opinion is misleading.
+A consideration about the organisation of the following sections.
+Given that the framework consists of two models, force prediction and shape prediction, I would expect section III.1 to be Force prediction and III.2 to the shape prediction (or vice-versa).
+However, the organization of Sections IV, V, VI, in my opinion is misleading.
+
+We have reorganised the paper as you request.
 
@@
 Regarding this point, I would like to know how the particles are placed inside the object.
-What happen if we change the size of the object?
-Do we
-need to change the number of particles? What happen if the object is not rectangular?
+What happens if we change the size of the object?
+Do we need to change the number of particles?
+What happen if the object is not rectangular?
 
-IV. A MASS-SPRING MODEL OF DEFORMATION
+Vero?
 
-My main concern about this section is Eq. 19. I really do not understand
-how the force could be proportional to the velocity. This point should be
-clarified. My understanding is that the authors assumed the acceleration
-(a) to be constant during the interval h, so that h*F = h*a*m => F =
-v*m/h.
-Given that h is a constant, the force (F), the velocity (v) or the mass
-(m)
-are recovered with a scale factor. Is that correct? If so, what is the
-difference between this approach and the quasi-static approach proposed by
-Frank. Please, explain.
+Eq.
+19.
+I really do not understand how the force could be proportional to the velocity.
+This point should be clarified.
+My understanding is that the authors assumed the acceleration (a) to be constant during the interval h, so that h*F = h*a*m => F = v*m/h.
+Given that h is a constant, the force (F), the velocity (v) or the mass (m) are recovered with a scale factor.
+Is that correct?
+If so, what is the difference between this approach and the quasi-static approach proposed by Frank.
+Please, explain.
 
-Other considerations about this section. This section is quite long and I
-think it could be simplified notably. The authors could simply state the
-energy E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2], where N is the number of
-constraints. There are several constraints to preserve the length, the
-area, etc. and the force is given as negative gradient of the potential
-energy with respect to the position of the particle. Details about all the
-equation could be reported in an appendix to make the paper more readable.
-About the equation, personally I don’t like the symbol composed of
-multiple letters such as Area, forcemag, forcedir, etc.
+Vero?
 
-VII. EXPERIMENTAL RESULTS
+This section is quite long and I think it could be simplified notably. The authors could simply state the
+energy E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2], where N is the number of constraints.
+There are several constraints to preserve the length, the area, etc.
+and the force is given as negative gradient of the potential
+energy with respect to the position of the particle.
+Details about all the equation could be reported in an appendix to make the paper more readable.
+About the equation, personally I don’t like the symbol composed of multiple letters such as Area, forcemag, forcedir, etc.
 
-Do the authors use just one video for the training and two video for the
-testing? It seems that the dataset is quite small to have a systematic
-test of the algorithm. Even if there are only two objects, the authors should
-test different motions of the finger (i.e. pushing velocity, penetration,
-…)
+We have shortened the section as requested.
 
-I don’t see the results of the classification. I would expect a confusion
-matrix to show how many times the sponge was classified as sponge and as
-plasticine and vice-versa. Table III seems to show the results about the
-classification of the pixels.
+Do the authors use just one video for the training and two video for the testing?
+It seems that the dataset is quite small to have a systematic test of the algorithm.
+Even if there are only two objects, the authors should test different motions of the finger (i.e.
+pushing velocity, penetration, …)
+
+This is rather difficult to modify as we gathered the data set three years ago, and it would not be possible to extend it quickly enough using the same set up.
+However, to try to satisfy your request we are currently working on an extension of the data set using simulated data from a physics engine.
+While this will not be physically accurate it will allow further testing of the model.
+We are afraid we have not yet completed this work for this revision, as we focussed on completing the study of the comparable method, but we will submit our additional data and the results with a new revision when they are available.
+
+I don’t see the results of the classification.
+I would expect a confusion matrix to show how many times the sponge was classified as sponge and as plasticine and vice-versa.
+Table III seems to show the results about the classification of the pixels.
+
+Vero?
 
 Moreover, I would like to see a comparison against a baseline approach.
-They could some of the models proposed in the literature or some simple
-models. For example, the authors could assume that the material
-“disappears” when the finger penetrates the object and report the F-Score
-of this method. Alternatively, they could use a simple model with only one
-spring between the finger and the base of the object.
+They could some of the models proposed in the literature or some simple models.
+For example, the authors could assume that the material “disappears” when the finger penetrates the object and report the F-Score of this method.
+Alternatively, they could use a simple model with only one spring between the finger and the base of the object.
 
-VIII.DISCUSSION
+We have implemented a growing neural gas network previously published by Cretu et al. We have added additional results in a comparison of this method with our own.
 
-Page 12. Col 2. Line 3. Please, can you clarify the hand tuning? It seems
-that the genetic algorithm is not the best way to find the parameters?
-Have you tried to use other method like gradient descent? Is it possible to use
-the model to mass-spring model to predict the force exerted on the finger?
-In that case, the authors could use this prediction error as fitness
-function and use the gradient descent to find the parameters.
+Vero?
+
+Page 12.
+Col 2.
+Line 3.
+Please, can you clarify the hand tuning?
+It seems that the genetic algorithm is not the best way to find the parameters?
+Have you tried to use other method like gradient descent?
+Is it possible to use the model to mass-spring model to predict the force exerted on the finger?
+In that case, the authors could use this prediction error as fitness function and use the gradient descent to find the parameters.
+
+Vero?
 
@@
 Fig. 1 is not referenced in the paper.
+
+We have decided that this figure added little, hence it was not referenced naturally anywhere in the text.
+We have thus decided to remove it to save space, since other requested changes have added considerable length.
+
 Page 2, Col 1, Line 17. Section VI is not mentioned.
+
+We have fixed this.
+
 Page 2. Col 2, Line 5: FEM is not defined. Finite-element models?
+
+We have defined this.
+
 Fig 2. Instead of use symbols, I would write Vision, strain, force, etc.
+
+Thank you, this chimes with the comments of Reviewer 2, we have fixed this in Figures 1, 2 and 3.
+
 Page 3, Col 1, line 57. Cretu et al. -> cite [10]?
+
+We have added this reference.
+
 Page 8. Col 2, Line 27. Linear snake = linear spline?
+
+Vero?
+
 Algorithm 5. I would specify that the F-Score is the fitness function of

### 2016-12-25 twostagedo.tex
Answered to most technical details.  Two are missing.

@@
 \begin{align}
- F = m \dfrac{\partial^2 p}{{\partial t}^2} \label{eq:newton}
+ F = m \dfrac{\partial \vec{p}}{\partial t} \label{eq:newton}
 \end{align}
@@
 \begin{align}
- F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
+ F + F_{ext} = m \dfrac{\partial \vec{p}}{\partial t}
 \end{align}
 Another option could be to induce response forces by displacing vertices from their position of equilibrium.
-}
 
@@
 Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.
-The potential energy terms for the preservation of length with damping and preservation of area were included in our system \comment{exactly as they were defined by Teschner}.
-To compensate for the lack of a term for preservation of volume, which tends to prevent distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.
+\comment{Since the location of the masses and direction of the springs will bias the deformation response of the object, it is important to distribute the masses uniformly and place the springs with a symetric arrangement for an isotropic material.
+This placement can be done with an algorithm that takes into account the symmetries of the object at the begining of the experiment.
+If the material gets permanently deformed a properly callibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}
+
+The potential energy terms for the preservation of length with damping and preservation of area were included in our system exactly as they were defined by Teschner.
+To compensate for the lack of a term for preservation of volume, which tends to prevent distortions of the graph in the original model, we derived a new term for the preservation of the angles in every triangle.
 The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.
 
-\comment{Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.
-He makes use of different constraints to model the tendency of an object to recover several of its attributes.} For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.
+Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.
+He makes use of different constraints to model the tendency of an object to recover several of its attributes.
+For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.
 When the value of the constraint is zero, the potential energy is zero as well.
@@
 \end{align}
-where $k$ is a proportionality constant \comment{\sout{\eref{eq:EConstraint}}} and $p_i$ is the position of particle $i$.
+where $k$ is a proportionality constant and $p_i$ is the position of particle $i$.
 
@@
 \end{align}
-Given the oscillatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.
-This is modelled with a force acting in the opposite direction to the velocity of the affected particles\comment{, thus slowing down motion.
-Teschner wrote this as a function of the constraints as well \cite{Teschner2004}}:
+Given the oscillatory nature of these equations, when using constraints that model the behaviour of springs, damping can be introduced to stop the oscillations.
+This is modelled with a force acting in the opposite direction to the velocity of the affected particles, thus slowing down motion.
+Teschner wrote this as a function of the constraints as well \cite{Teschner2004}:
 \begin{align}
@@
 \subsubsection{Preservation of Area}
-This constraint is applied per triangle in the mesh.  Since Teschner and Morris \comment{%\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} 
-found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
+This constraint is applied per triangle in the mesh.  Since Teschner and Morris
+found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.
+Therefore the force $\vec{F}^i_A$ is derived as follows:
 \begin{align}
@@
 [\fref{fig:forces}(c)] gives an idea of how these forces look.
-Here the energy depends on the difference \comment{%\sout{of the angles between adjacent edges.}
-between the current angle, between adjacent edges, and the angle between them at the equilibrium position.} Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
+Here the energy depends on the difference between the current angle, between adjacent edges, and the angle between them at the equilibrium position.
+Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
 \begin{align}
@@
  \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
- %\STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
- \comment{\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.} And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
+ \STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.
+And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
  \IF {$\alpha > max\_\alpha$}
@@
 \label{sec:ischeme}
-\comment{Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations,} we made the \comment{force proportional to the velocity:
+Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations, we made the force proportional to the velocity:
+\comment{
 \begin{align}
- F(t) = m \dfrac{\partial \vec{p}}{\partial t}
+ F(t) = \vec{p}
 \end{align}
-and applied Euler only once to solve for $p$:
+which implies the following assumptions:
+\begin{align}
+ \dfrac{\partial \vec{p}}{\partial t} = \vec{p}
+\end{align}
+this equation can only be satisfied with an exponential, thus avoiding the oscillatory behaviour of springs in favour of forcing critically damped springs to model the elastic materials.
+}
+
+Then we applied Euler only once to solve for $p$:
 \begin{equation}
  \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
-\end{equation}}
+\end{equation}
 with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.
-\comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is no inertia in the oscillation.
+We acknowledge that this is not the proper Newtonian model, however the simulations obtained were closer to the observed behaviour, because there is no inertia in the oscillation.
 When experimenting with different integration steps, force values were interpolated between measurements.
@@
 \subsection{Collision Detection}
-In the simulation, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.
-\comment{When the simulation of the mesh deformation runs---used to model the sponge---the predicted positions of the vertices may indicate that the mesh penetrates an obstacle.
+In the simulation, rigid objects (like the robotic finger and the table) are treated as geometric obstacles.
+When the simulation of the mesh deformation runs---used to model the sponge---the predicted positions of the vertices may indicate that the mesh penetrates an obstacle.
 This should be interpreted as a collision between the sponge and an obstacle.
@@
 For example: if the vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
-}
-
-\comment{In order to model this, we model the finger and table using geometric primitives.
+
+In order to model this, we model the finger and table using geometric primitives.
 For the table obstacle, any interpenetration by the sponge is solved by pushing vertices of the mesh of the sponge back to the border of the table.
 The springs will then propagate the effect to their neighbours.
-For the finger obstacle, resolution is in two stages: by vertex and then by edge.}  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.
-For edges, the \comment{standard} algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.
+For the finger obstacle, resolution is in two stages: by vertex and then by edge.
+A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.
+For edges, the standard algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.
 Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.
@@
 
-\subsection{\comment{Additional} Geometric Constraints}
-\comment{While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function.
+\subsection{Additional Geometric Constraints}
+While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function.
 Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.
-It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
-We achieve this through geometric checking subroutines.}  There are two geometric constraints that the mesh is enforced to maintain:
+\comment{All of these situations are innadecuate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
+We achieve this through geometric checking subroutines.
+There are two geometric constraints that the mesh is enforced to maintain:
 
@@
  \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed more than half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ of this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.
-This is enough for most cases, but triangles in the border of the mesh can still overlap each other.
+This is enough for most cases, but triangles in the border of the mesh can still overlap each other \comment{which would render a bad numerical approximation to the behaviour of a surface}.
 
@@
 
-The next stage \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.
-This problem is addressed in the following section.
-
-\section{\comment{Learning to Predict Deformations}}
+The next stage is the learning process which consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.
+
+\section{Learning to Predict Deformations}
 \label{sec:learning}
@@
 The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
-\comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}.
+Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.
+Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}.
 The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
 Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
-This is explained below.}  The tracking of the deformable target in the ground truth has two stages:
+\comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.
+The tracking of the deformable target in the ground truth has two stages:
 \begin{enumerate}
@@
 The Canny algorithm is used to extract the edges of the image.
-\comment{  Since the object of interest has a distinctive colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.
-If the average hue of the pixels on an edge falls outside a previously calibrated range, the edge is eliminated.}
+Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.
+Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.
+If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
  \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
@@
 \subsection{Evolutionary Algorithm}
-\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they instead depend on the geometry of the mesh used to model the object.
+\label{sub:EvolAlg}
+
+Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they instead depend on the geometry of the mesh used to model the object.
 There can thus be several sets of parameters in different regions of the parameter space that produce adequate simulations \cite{Morris2008}.
-We therefore chose to use an evolutionary algorithm to both perform global search, and local improvement.}
+We therefore chose to use an evolutionary algorithm to both perform global search, and local improvement.
 
@@
  \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
- \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well.)}
+ \item The integration time step $\Delta t$.
+\textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well; \comment{however, since they didn't improve stability or accuracy in a visible way, but made the process much slower, they were disregarded as impractical}.)}
  \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
@@
  \STATE Let $N$ be the number of sets of parameters to be evaluated per generation and $M$ the number of generations.
- \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements\comment{, with: \\
+ \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements, with: \\
  $\alpha \leftarrow$ number of elements that survive for the next generation as they are, \\
@@
  $\zeta \leftarrow$ new individuals. \\
- which will be chosen as follows:}
+ which will be chosen as follows:
  \STATE Let $\sigma_{max} > \sigma_{min}$ both be standard deviations of a Gaussian
@@
 
-This measurement \comment{is know as the F-score and} is used in the computer vision community to evaluate segmentation algorithms.
+This measurement is know as the F-score and is used in the computer vision community to evaluate segmentation algorithms.
 $F=1$ means that there were no false predictions ($FN+FP = 0$).
 $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).
-$F$ is calculated for every frame and the score for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
-\comment{\begin{align}
+$F$ is calculated for every frame and the score for the video at frame $i$ is the mean over all the frames so far, $\mu(F)$:
+\begin{align}
  \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
-\end{align}}
+\end{align}
 If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closer to zero, the more errors there were.
@@
 
-\section{\comment{Learning to predict forces}}
+\section{Learning to predict forces}
 %We used two methods to predict the forces.
@@
 
-Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{as the mass-spring model.
+Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the same training data set as the mass-spring model.
 This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.
@@
 From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.
-Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}
-
-The \comment{stress strain curves} in [\fref{fig:sstrain}] are approximated by segmented regression curves.
+Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.
+
+The stress strain curves in [\fref{fig:sstrain}] are approximated by segmented regression curves.
 Several candidates are proposed using the least squares technique and the best are chosen as follows.
@@
 
-\comment{For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.
+For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.
 To determine the points of discontinuity between segments, each family is evaluated incrementally.
@@
 Having discovered one best parameterized line for each family, the two resulting candidates are compared.
-We select the type of curve that covers more points in a single segment}.
+We select the type of curve that covers more points in a single segment.
 This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.
@@
 A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].
+\comment{This set up allowed us to study the effect of using an automatically calibrated physics based model as model for the material, while using 2D meshes, even though the same principles can be used if we use a 3D point cloud and a 3D mesh.
+To generate the initial mesh for the mass-spring model a routine placed the nodes for the masses evenly spaced on a rectangular grid, given the number of squares we wanted per side.
+We made experiments with grids of $3x2, 10x5$ and $20x10$ squares.
+To aliviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}
+
 The force sensor was a DAQ-FT-Gamma, that can measure forces up to 200 Newtons in the $z$ direction (up to 65N in the others) and torques up to 5 Newton-metre.
@@
 \multicolumn{4}{p{88mm}}{Where the mark is $\mu(F)$ as explained in \sref{sec:evaluation}.
-\comment{Mean $\mu(F)$ and standard deviation across 24 training sesions are reported for the best performing models of every training session, worst model and average model.}}
+Mean $\mu(F)$ and standard deviation across 24 training sesions are reported for the best performing models of every training session, worst model and average model.}
 \end{tabular}
@@
 On the other hand, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that more closely resemble the ground truth, thus obtaining better marks.
-From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them.
-
-\subsection{\comment{Results of Automatic Parameter Calibration}}
-\subsubsection{\comment{Performance for the Learned Shape Prediction Model}}
+From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them\footnote{\comment{The full list of terms is summarised at the end of \sref{sub:EvolAlg}.}}.
+
+\subsection{Results of Automatic Parameter Calibration}
+\subsubsection{Performance for the Learned Shape Prediction Model}
 The performance of the genetic learning algorithm was evaluated using the customised integration scheme and the mass-spring elastic-plastic model.
@@
 
-\subsubsection{\comment{Performance for the Learned Force Prediction Model}}
+\subsubsection{Performance for the Learned Force Prediction Model}
 
@@
 One hundred photographs and force readings were taken every $0.17\pm0.007 s$ in average.
-The mass spring model runs with an integration step \comment{$h=0.1s$}, where forces that were not measured directly are interpolated.
+The mass spring model runs with an integration step $h=0.1s$, where forces that were not measured directly are interpolated.
 The push down movement lasts approximately 50 frames and the retrieval movement the other 50.
 The continuous lines in \fref{fig:sstrain} show the predicted forces, given the amount of compression of the material (strain), while the point-lines show the actual readings.
-\fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the actual shapes predicted \comment{by the combined FP and SP model}.
+\fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the actual shapes predicted by the combined FP and SP model.
 
@@
 
+\comment{We also compared the generalisation capabilities of our model against the system proposed in \cite{Cretu2012}.
+This system uses a Growing Neural Gas (GNG) Network to calibrate a colour segmentation algorithm.
+We couldn't use the HSV space as suggested because our videos have lots of shadows which produce holes inside the material regions which cause problems in further stages.
+We got better results with Luv color space.
+We clarify that, even though we used the HSV space in our tracking algorithm, we didn't have the same problems because the nodes of our linear snake search for nearby edges and are not affected by shadows in the middle of the object or temporary failures in detecting an edge.
+Another GNG network is used to select the control points adequate for approximating the contour in the first frame.
+Later, Cretu et al.
+use a Neural Gas (NG) algorithm to adjust that fixed number of control points to the contour of the material at every frame.
+Using the coordinates of the finger, the sensed force and the coordinates of those control points through time, they train a Feedforward Neural Network (FFNN) to predict the contour of the material given the position of the finger and sensed force.
+We reproduced those steps and their training method on our training video.
+However, when we traid to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it got tied to the specific location and type of interaction it was trained with, while our physics based model had no problem at making predictions for new pushing actions.
+The results with this other method can be seen on \fref{fig:cretu_prediction}.
+
+Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.
+However, even if we could change that algorithm, we would expect the system to have the same problems and predict the deformation at the middle of the material even if is being pushed somewhere else.
+}
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3.5in]{sponge__result.jpg}
+\caption{Neural network approach fits quite well the training data set, but can not generalise to new pushing actions.
+The green line shows the border used as ground truth, while blue lines show prediction from the FFNN.}\label{fig:cretu_prediction}
+\end{figure}
+
 \subsection{Classification}
 
-The mass-spring models \comment{calibrated for the prediction task} were used to classify \comment{materials} through recognition of their behaviour, even \comment{when the objects where pushed and deformed at novel locations.}  The models make predictions as if the \comment{material of the} object was known.
-The model that makes the best predictions indicates which \comment{type of} object is being observed.
-There are two criteria for selecting an \comment{object/material for the video being shown to the system}:
+The mass-spring models calibrated for the prediction task were used to classify materials through recognition of their behaviour, even when the objects where pushed and deformed at novel locations.
+The models make predictions as if the material of the object was known.
+The model that makes the best predictions indicates which type of object is being observed.
+There are two criteria for selecting an object/material for the video being shown to the system:
 
 \begin{enumerate}
- \item Per frame: \comment{a material is selected for each frame of the interaction according to whichever model made the best prediction for that particular frame, according to the F-score.} This test was performed for every frame of the six videos, where the robot finger pushes the block of material.
-\tref{tab:classperframe} shows the results for the videos that were used as training and test sets for the prediction task.
+ \item Per frame: a material is selected for each frame of the interaction according to whichever model made the best prediction for that particular frame, according to the F-score.
+This test was performed for every frame of the six videos, where the robot finger pushes the block of material.
+
+ \item Global: The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted score.
+The difference in the F-score is added for all frames.
+Frames where the distinction is more noticeable will receive a greater weight.
+\end{enumerate}
+
+\section{Discussion}
+\label{sec:discussion}
+We note the following points about the results and the method in general.
+\subsection{The Learning Method}
+Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
+Better simulations have been obtained by hand tuning some of the parameters.
+%\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
+The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}
+
+\subsection{Using Additional Geometric Constraints}
+Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.
+Also, when attempting to model the plastic material, the geometric constraints provoked a much larger effect caused by the plasticity term, even if its corresponding constant was small.
+As a consequence the plastic material also recovered its original shape after some extra frames.
+It has been seen that, if the constant for linear preservation is set to zero, the problem disappears.
+Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.
+We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the beginning, to the instant of greatest deformation and to the final shape.
+Meanwhile, the evaluation function gives the same weight to all frames in the sequence.
+It is possible to use the force learning algorithm and the commands given to the robot (push/retrieve) to detect these points, since they are correlated with discontinuities in these other modalities.
+Future work will evaluate the impact of these modifications on the evaluation function.
+
+\subsection{Training With Collision Resolution with the Finger vs. Without}
+It was found that the overall score of the best simulations was similar in both cases, since only geometric similarities are under consideration.
+However, there was a noticeable difference in the rate of success for the classification problem.
+If no collision detection with the finger is used during training, the springs react only to the forces and, if the spring coefficients are not adequate, the shape of the mesh will be very different from the shape of the ground truth.
+
+On the other hand, if collision detection is used, the geometric constraints strongly favour a correct configuration of the mesh in the vicinity of the actuator, even though the springs did not react properly.
+For subsequent frames, it is more likely, when using collision detection, that the neighbours will succeed in propagating the deformation to the rest of the material, thus producing better simulations than without it.
+However, this lessens the effectiveness of the mass-spring model in distinguishing the materials.
+
+\subsection{Classification}
 
@@
 \hline
-%\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.
+\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN measured in number of frames.
+TP + FN = total number of frames.
+Final class assigned according to the performance of the models' predictions across the whole video.} \end{footnotesize}}
+\end{tabular}
+\end{table}
+% \begin{table}[!t]
+% \renewcommand{\arraystretch}{1.3}
+% \caption{Classification per Frame}
+% \label{tab:classperframe}
+% \centering
+% \begin{tabular}{ccccccccc}
+% \hline
+%  & & \multicolumn{4}{c}{\bfseries Sponge} \\
+% Data Set & & TP & FN & TN & FP & Precision \\
+% \hline\hline
+%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
+% \hline
+%    training & & 167 & 132 & & & 0.56  \\
+%    test1 & & 167 & 132 & & & 0.56  \\
+%    test2 & & 242 & 57 & & & 0.81 \\ 
+% \hline
+% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.
 Final class assigned according to the performance of the models' predictions across the whole video.
 \end{footnotesize}}
-\end{tabular}
-\end{table}
-
- \item Global: \comment{The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted score.} The difference in the \comment{F-score} is added for all frames.
-Frames where the distinction is more noticeable will receive a greater weight.
-\end{enumerate}
-
-\section{Discussion}
-\label{sec:discussion}
-We note the following points about the results and the method in general.
-\subsection{The Learning Method}
-Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
-Better simulations have been obtained by hand tuning some of the parameters.
-%\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
-The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}
-
-\subsection{Using \comment{Additional} Geometric Constraints}
-Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.
-\comment{Also, when attempting to model the plastic material, the geometric constraints provoked a much larger effect caused by the plasticity term}, even if its corresponding constant was small.
-As a consequence the plastic material also recovered its original shape after some extra frames.
-It has been seen that, if the constant for linear preservation is set to zero, the problem disappears.
-Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.
-We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the beginning, to the instant of greatest deformation and to the final shape.
-Meanwhile, the evaluation function gives the same weight to all frames in the sequence.
-It is possible to use the force learning algorithm and the commands given to the robot (push/retrieve) to detect these points, since they are correlated with discontinuities in these other modalities.
-Future work will evaluate the impact of these modifications on the evaluation function.
-
-\subsection{Training With Collision Resolution with the Finger vs. Without}
-It was found that the overall score of the best simulations was similar in both cases, since only geometric similarities are under consideration.
-However, there was a noticeable difference in the rate of success for the classification problem.
-If no collision detection with the finger is used during training, the springs react only to the forces and, if the spring coefficients are not adequate, the shape of the mesh will be very different from the shape of the ground truth.
-
-On the other hand, if collision detection is used, the geometric constraints strongly favour a correct configuration of the mesh in the vicinity of the actuator, even though the springs did not react properly.
-For subsequent frames, it is more likely, when using collision detection, that the neighbours will succeed in propagating the deformation to the rest of the material, thus producing better simulations than without it.
-However, this lessens the effectiveness of the mass-spring model in distinguishing the materials.
-
-\subsection{Classification}
+% \end{tabular}
+% 
+% \vspace*{1em}
+% \begin{tabular}{ccccccccc}
+% \hline
+%  & & \multicolumn{4}{c}{\bfseries Plasticine} \\
+% Data Set & & TP & FN & TN & FP & Precision\\
+% \hline\hline
+%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
+% \hline
+%    training & & 186 & 113 & & & 0.62 \\
+%    test1 & & 271 & 28 & & & 0.91 \\
+%    test2 & & 289 & 10 & & & 0.97 \\ 
+% \hline
+% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.
+Final class assigned according to the performance of the models' predictions across the whole video.
+\end{footnotesize}}
+% \end{tabular}
+% \end{table}
+
+\comment{
+\tref{tab:classperframe} shows the results for the classification task, frame per frame, for the videos that were used as training and test sets for the prediction task.
+When the sponge model performed better for a frame of the sponge video it is marked as a true positive, when plasticine does better it is marked as a false negative (since it didn't recognize it as a sponge).
+The same process is used for the other material.
+Since the concept was proved with only two materials, true positives and false negatives provide all the information.
+
+It calls for attention that in some of the experiments, like the one illustrated above, the performance of classification for the sponge on the training and test1 sets are the same.
+We interpret that the mass-spring physics based model generalizes naturally when the force/deformation conditions are similar, even if the place for pushing is different.
+Between pushing in the middle and pushing closer to the border the qualitative behaviour of the elastic material was no so different for these models; unlike the data from the third set, where the distribution of mass and maximum force/deformation is more different from the ones seen on the training set of the model.
+It is interesting to confirm that the difference in predictions between the models is bigger here in favour of a better classification.
+
+It is also remarkable that, even though, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.
+That is, the parameters for the sponge are not good at modeling the plasticine on the test sets, thus yielding better results at the classification task.
+This could happen because the plastic behaviour (that is, the permanent deformation of a spring) gets activated with stronger forces after the $yield$ limit is reached, which only appear when pushing the plasticine.
+With the small forces feedback by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour do not manifest.
+}
+
 The best results for classification with the mass-spring model are obtained when the training of the models uses only forces to displace the nodes in the mass-spring model\footnote{Case 2 in \sref{sec:manysteps}}, but the combination of force readings and collision resolution is used during classification\footnote{Case 3 in \sref{sec:manysteps}}.

### 2016-12-25 twostagedo-rev-Dec-16.tex
merged changes from Vero into Jeremy’s changes

@@
 \begin{align}
- F = m \dfrac{\partial^2 p}{{\partial t}^2} \label{eq:newton}
+F = m \dfrac{\partial \vec{p}}{\partial t} \label{eq:newton}
 \end{align}
@@
 \begin{align}
- F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
+F + F_{ext} = m \dfrac{\partial \vec{p}}{\partial t}
 \end{align}
@@
  \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
- %\STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
- \comment{\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.} And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
+\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.
+And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
  \IF {$\alpha > max\_\alpha$}
@@
  The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
-\comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}.
+Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.
+Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}.
 The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
 Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
-This is explained below.}  The tracking of the deformable target in the ground truth has two stages:
+\comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.
+The tracking of the deformable target in the ground truth has two stages:
 \begin{enumerate}
@@
  \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
- \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well.)}
+  \item The integration time step $\Delta t$.
+\textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well; \comment{however, since they didn't improve stability or accuracy in a visible way, but made the process much slower, they were disregarded as impractical}.)}
  \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
@@
 
-This measurement \comment{is know as the F-score and} is used in the computer vision community to evaluate segmentation algorithms.
+This measurement is known as the F-score and is used in the computer vision community to evaluate segmentation algorithms.
 $F=1$ means that there were no false predictions ($FN+FP = 0$).
 $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).
-$F$ is calculated for every frame and the score for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
-\comment{\begin{align}
+$F$ is calculated for every frame and the score for the video at frame $i$ is the mean over all the frames so far, $\mu(F)$:
+\begin{align}
  \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
-\end{align}}
+\end{align}
 If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closer to zero, the more errors there were.
@@
 From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.
-Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}
-
-The \comment{stress strain curves} in [\fref{fig:sstrain}] are approximated by segmented regression curves.
+Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.
+
+The stress strain curves in [\fref{fig:sstrain}] are approximated by segmented regression curves.
 Several candidates are proposed using the least squares technique and the best are chosen as follows.
 Since there is only one independent variable the equations for adjusting a line are straightforward:
+
 \begin{align}
@@
 \hline
-%\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.
+\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN measured in number of frames.
+TP + FN = total number of frames.
+Final class assigned according to the performance of the models' predictions across the whole video.} \end{footnotesize}}
+\end{tabular}
+\end{table}
+% \begin{table}[!t]
+% \renewcommand{\arraystretch}{1.3}
+% \caption{Classification per Frame}
+% \label{tab:classperframe}
+% \centering
+% \begin{tabular}{ccccccccc}
+% \hline
+%  & & \multicolumn{4}{c}{\bfseries Sponge} \\
+% Data Set & & TP & FN & TN & FP & Precision \\
+% \hline\hline
+%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
+% \hline
+%    training & & 167 & 132 & & & 0.56  \\
+%    test1 & & 167 & 132 & & & 0.56  \\
+%    test2 & & 242 & 57 & & & 0.81 \\ 
+% \hline
+% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.
 Final class assigned according to the performance of the models' predictions across the whole video.
 \end{footnotesize}}
-\end{tabular}
-\end{table}
-
- \item Global: \comment{The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted score.} The difference in the \comment{F-score} is added for all frames.
-Frames where the distinction is more noticeable will receive a greater weight.
-\end{enumerate}
-
-\section{Discussion}
-\label{sec:discussion}
-We note the following points about the results and the method in general.
-\subsection{The Learning/Calibration Method}
-Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
-Better simulations have been obtained by hand tuning some of the parameters.
-%\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
-The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}
-
-\subsection{Using \comment{Additional} Geometric Constraints}
-Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.
-\comment{Also, when attempting to model the plastic material, the geometric constraints provoked a much larger effect caused by the plasticity term}, even if its corresponding constant was small.
-As a consequence the plastic material also recovered its original shape after some extra frames.
-It has been seen that, if the constant for linear preservation is set to zero, the problem disappears.
-Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.
-We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the beginning, to the instant of greatest deformation and to the final shape.
-Meanwhile, the evaluation function gives the same weight to all frames in the sequence.
-It is possible to use the force learning algorithm and the commands given to the robot (push/retrieve) to detect these points, since they are correlated with discontinuities in these other modalities.
-Future work will evaluate the impact of these modifications on the evaluation function.
-
-\subsection{Training With Collision Resolution with the Finger vs. Without}
-It was found that the overall score of the best simulations was similar in both cases, since only geometric similarities are under consideration.
-However, there was a noticeable difference in the rate of success for the classification problem.
-If no collision detection with the finger is used during training, the springs react only to the forces and, if the spring coefficients are not adequate, the shape of the mesh will be very different from the shape of the ground truth.
-
-On the other hand, if collision detection is used, the geometric constraints strongly favour a correct configuration of the mesh in the vicinity of the actuator, even though the springs did not react properly.
-For subsequent frames, it is more likely, when using collision detection, that the neighbours will succeed in propagating the deformation to the rest of the material, thus producing better simulations than without it.
-However, this lessens the effectiveness of the mass-spring model in distinguishing the materials.
-
-\subsection{Classification}
+% \end{tabular}
+% 
+% \vspace*{1em}
+% \begin{tabular}{ccccccccc}
+% \hline
+%  & & \multicolumn{4}{c}{\bfseries Plasticine} \\
+% Data Set & & TP & FN & TN & FP & Precision\\
+% \hline\hline
+%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
+% \hline
+%    training & & 186 & 113 & & & 0.62 \\
+%    test1 & & 271 & 28 & & & 0.91 \\
+%    test2 & & 289 & 10 & & & 0.97 \\ 
+% \hline
+% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.
+Final class assigned according to the performance of the models' predictions across the whole video.
+\end{footnotesize}}
+% \end{tabular}
+% \end{table}
+
+\comment{
+\tref{tab:classperframe} shows the results for the classification task, frame per frame, for the videos that were used as training and test sets for the prediction task.
+When the sponge model performed better for a frame of the sponge video it is marked as a true positive, when plasticine does better it is marked as a false negative (since it didn't recognize it as a sponge).
+The same process is used for the other material.
+Since the concept was proved with only two materials, true positives and false negatives provide all the information.
+
+It calls for attention that in some of the experiments, like the one illustrated above, the performance of classification for the sponge on the training and test1 sets are the same.
+We interpret that the mass-spring physics based model generalizes naturally when the force/deformation conditions are similar, even if the place for pushing is different.
+Between pushing in the middle and pushing closer to the border the qualitative behaviour of the elastic material was no so different for these models; unlike the data from the third set, where the distribution of mass and maximum force/deformation is more different from the ones seen on the training set of the model.
+It is interesting to confirm that the difference in predictions between the models is bigger here in favour of a better classification.
+
+It is also remarkable that, even though, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.
+That is, the parameters for the sponge are not good at modeling the plasticine on the test sets, thus yielding better results at the classification task.
+This could happen because the plastic behaviour (that is, the permanent deformation of a spring) gets activated with stronger forces after the $yield$ limit is reached, which only appear when pushing the plasticine.
+With the small forces feedback by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour do not manifest.
+}
+
 The best results for classification with the mass-spring model are obtained when the training of the models uses only forces to displace the nodes in the mass-spring model\footnote{Case 2 in \sref{sec:manysteps}}, but the combination of force readings and collision resolution is used during classification\footnote{Case 3 in \sref{sec:manysteps}}.

### 2016-12-25 twostagedo-rev-Dec-16.tex
made changes to response, and to paper

@@
 \subsection{Scenario}
-For the experimental scenario a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth, is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.
+For the experimental scenario we used two materials: sponge and plasticine.
+Each formed a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth.
+One of the objects is selected, and is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.
 A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].

### 2016-12-25 twostagedo-rev-Dec-16.tex
Final revisions from Jeremy!

@@
 
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/classification}
+\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
+The most accurate predictor at any one moment gives the classification for the material type.
+The notation is explained in the text.}
+\label{fig:classification}
+\end{figure*}
+
 The third system is for material classification (Figure~\ref{fig:classification}).
@@
 
-\begin{figure*}[t]
-\centering
-%\setlength\fboxsep{0pt}    % padding
-%\setlength\fboxrule{0.5pt} % line width
-%\fbox{
-\includegraphics[width=0.8\textwidth]{figures/classification}%}
-\caption{Prediction sequences for several predictors--each of which has been learned for a different material--can be compared against real data.
-The most accurate predictor at any one moment gives the classification for the material type.
-The notation is explained in the text.}
-\label{fig:classification}
-\end{figure*}
 %In each case the resulting learned models can be run in two modes, prediction and classification.
@@
 %The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
-For clarity, the algorithm is described in exactly these three phases: training, simulation (i.e.
+For clarity, the algorithm is described in exactly these phases: training, simulation (i.e.
 the prediction phase), and classification.
 In empirically evaluating the model new data was used, acquired from new finger pushes.
-We now proceed to describe the shape predictor SP (Section~\ref{sec:model}), which is based on a mass-spring system, and then the force predictor (FP), which is a regression model (Section~\ref{sec:forces}).
+We now proceed to describe first the visual analysis and force measurement, the force predictor (FP), which is a regression model (Section~\ref{sec:forces}), and then the shape predictor SP (Section~\ref{sec:model}), which is based on a mass-spring system.
+
+\section{Visual Analysis}
+\label{sec:vision}
+
+Although the visual analysis is not an original contribution of the paper, it is necessary to track the shape of the object as it deforms.
+This tracking is performed on the two training movies to provide training data for the shape predictor (SP).
+For the test movies the system only requires visual analysis of the first frame to initialise the predictors.
+The test movies are analysed for each frame, but only for experimental evaluation purposes.
+We now give details of the visual analysis method.
+
+\subsection{Tracking to Provide Ground Truth}
+\begin{figure}[!t]
+ \begin{algorithmicieee}{Regularised Linear Spline}\label{alg:regularise}
+ \STATE Let $L$ be a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
+ \FORALL {$L_i$ in $L$}
+ \IF {$dist(L_i,L_{i-1}) > max\_distance$}
+  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
+ \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
+  \STATE erase $L_i$
+ \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
+  \STATE erase $L_i$
+ \ENDIF
+ \ENDFOR
+\end{algorithmicieee}
+\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
+\end{figure}
+ The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
+Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.
+Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}.
+The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
+Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
+\comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.
+The tracking of the deformable target in the ground truth has two stages:
+ \begin{figure}[t]
+\centering
+\includegraphics[width=88mm]{arrio2}
+\caption{(a) Experimental set up.
+A robot finger pushes a block of deformable material sitting on a table.
+A camera registers the events of the plane where the main deformations take place.
+(b) Evaluation function.
+Left top: Mesh over real block of material.
+Left bottom: The typical error of the model vs.
+the ground truth (red) and the typical error of the tracker (blue).
+Right top: tracked ground data.
+Middle: TP in yellow, FP in blue, FN in red and TN in gray.
+Bottom: simulated mesh.
+}
+\label{fig:scene}
+\end{figure}
+
+\begin{enumerate}
+ \item The target object is segmented from the environment.
+The Canny algorithm is used to extract the edges of the image.
+Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.
+Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.
+If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
+ \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
+\end{enumerate}
+For the first frame, the snake is initialised as a rectangle around the target.
+This rectangle fragments itself into smaller pieces.
+Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].
+If there is no edge within that area, the control point remains where it is.
+This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.
+Once the edge reappears, the control point adapts itself again.
+Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.
+See \alref{alg:regularise}.
+This allows the snake to adapt very quickly to the complexity of the contour it represents.
+
+\section{Learning to predict forces}
+\label{sec:forces}
+
+It is possible to learn the force predictor (FP), by using information from the stress-strain diagram obtained from the data set of measured positions and forces.
+This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by the sensor located at the wrist.
+When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.
+From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.
+Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.
+
+The stress strain curves in [\fref{fig:sstrain}] are approximated by segmented regression curves.
+Several candidates are proposed using the least squares technique and the best are chosen as follows.
+Since there is only one independent variable the equations for adjusting a line are straightforward:
+
+\begin{align}
+ y & = mx + y_0 + Error \\
+ m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
+ y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
+ Error & = \sum(y_i - (mx_i+y0))^2
+\end{align}
+Where $y$ is the dependent variable, $x$ the independent one, $m$ is the slope of the line, and $y_0$ the $y$ intercept.
+To adjust logarithmic curves a change of variable is used.
+Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.
+In this case, the final equation will be of the form: $y = m \ln x + y_0$.
+It would be possible to add other types of curves, but these were enough for the materials covered.
+
+For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.
+To determine the points of discontinuity between segments, each family is evaluated incrementally.
+The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.
+We begin with a small number of points (e.g.
+five).
+To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.
+This marks the point of discontiuity.
+Having discovered one best parameterized line for each family, the two resulting candidates are compared.
+We select the type of curve that covers more points in a single segment.
+This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.
 
 \section{A Mass-Spring Model of Deformation}
-
+The shape predictor (SP) is based on a mass-spring model of deformation.
 A mass-spring model is a very well established abstraction, commonly used to simulate the behaviour of certain types of deformable objects.
@@
 This placement can be done with an algorithm that takes into account the symmetries of the object at the begining of the experiment.
-If the material gets permanently deformed a properly callibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}
+If the material gets permanently deformed a properly calibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}
 
@@
 Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.
-\comment{All of these situations are innadecuate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
+\comment{All of these situations are inadequate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
 We achieve this through geometric checking subroutines.
@@
 \label{sec:learning}
-
-\subsection{Tracking to Provide Ground Truth}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=88mm]{arrio2}
-\caption{(a) Experimental set up.
-A robot finger pushes a block of deformable material sitting on a table.
-A camera registers the events of the plane where the main deformations take place.
-(b) Evaluation function.
-Left top: Mesh over real block of material.
-Left bottom: The typical error of the model vs.
-the ground truth (red) and the typical error of the tracker (blue).
-Right top: tracked ground data.
-Middle: TP in yellow, FP in blue, FN in red and TN in gray.
-Bottom: simulated mesh.
-}
-\label{fig:scene}
-\end{figure}
-
- The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour.
-Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.
-Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}.
-The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
-Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
-\comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.
-The tracking of the deformable target in the ground truth has two stages:
-\begin{enumerate}
- \item The target object is segmented from the environment.
-The Canny algorithm is used to extract the edges of the image.
-Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.
-Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.
-If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
- \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
-\end{enumerate}
-For the first frame, the snake is initialised as a rectangle around the target.
-This rectangle fragments itself into smaller pieces.
-Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].
-If there is no edge within that area, the control point remains where it is.
-This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.
-Once the edge reappears, the control point adapts itself again.
-Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.
-See \alref{alg:regularise}.
-This allows the snake to adapt very quickly to the complexity of the contour it represents.
-
-\begin{figure}
- \begin{algorithmicieee}{Regularise Linear Spline}\label{alg:regularise}
- \STATE Let $L$ be a linear spline representing the contour of a $2D$ shape, and $L_i$ the $i^{th}$ control point in a cyclic order.
- \FORALL {$L_i$ in $L$}
- \IF {$dist(L_i,L_{i-1}) > max\_distance$}
-  \STATE Insert $midpoint(L_i,L_{i-1})$ after $L_i$
- \ELSIF {$dist(L_i,L_{i-1}) < min\_distance$}
-  \STATE erase $L_i$
- \ELSIF {$angle(L_{i-1},L_i,L_{i+1}) < min\_angle$}
-  \STATE erase $L_i$
- \ENDIF
- \ENDFOR
-\end{algorithmicieee}
-\caption{Algorithm used to regularise the number of elements of a linear spline, in accordance with the level of detail required.}
-\end{figure}
 
@@
 
-\section{Learning to predict forces}
-%We used two methods to predict the forces.
-The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.
-This force is feed as input for the following simulation step.
-\label{sec:forces}
-
-Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the same training data set as the mass-spring model.
-This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.
-When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.
-From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.
-Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.
-
-The stress strain curves in [\fref{fig:sstrain}] are approximated by segmented regression curves.
-Several candidates are proposed using the least squares technique and the best are chosen as follows.
-Since there is only one independent variable the equations for adjusting a line are straightforward:
-
-\begin{align}
- y & = mx + y_0 + Error \\
- m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
- y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
- Error & = \sum(y_i - (mx_i+y0))^2
-\end{align}
-Where $y$ is the dependent variable, $x$ the independent one, $m$ is the slope of the line, and $y_0$ the $y$ intercept.
-To adjust logarithmic curves a change of variable is used.
-Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.
-In this case, the final equation will be of the form: $y = m \ln x + y_0$.
-It would be possible to add other types of curves, but these were enough for the materials covered.
-
-For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.
-To determine the points of discontinuity between segments, each family is evaluated incrementally.
-The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.
-We begin with a small number of points (e.g.
-five).
-To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.
-This marks the point of discontiuity.
-Having discovered one best parameterized line for each family, the two resulting candidates are compared.
-We select the type of curve that covers more points in a single segment.
-This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.
-
 \section{Experimental Results}
@@
 To generate the initial mesh for the mass-spring model a routine placed the nodes for the masses evenly spaced on a rectangular grid, given the number of squares we wanted per side.
-We made experiments with grids of $3x2, 10x5$ and $20x10$ squares.
-To aliviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}
+We made experiments with grids of $3 \times 2, 10 \times 5$ and $20 \times 10$ squares.
+To alleviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}
 
@@
 This system uses a Growing Neural Gas (GNG) Network to calibrate a colour segmentation algorithm.
-We couldn't use the HSV space as suggested because our videos have lots of shadows which produce holes inside the material regions which cause problems in further stages.
-We got better results with Luv color space.
-We clarify that, even though we used the HSV space in our tracking algorithm, we didn't have the same problems because the nodes of our linear snake search for nearby edges and are not affected by shadows in the middle of the object or temporary failures in detecting an edge.
+We couldn't use the HSV space, as suggested by Cretu, because our videos have lots of shadows which produce holes inside the material regions which cause problems in further stages.
+We obtained better results for GNGN with the Luv color space.
+Note that, even though we used the HSV space in our tracking algorithm, this didn't have the same problems because the nodes of our linear snake search for nearby edges and are not affected by shadows in the middle of the object or temporary failures in detecting an edge.
 Another GNG network is used to select the control points adequate for approximating the contour in the first frame.
 Later, Cretu et al.
-use a Neural Gas (NG) algorithm to adjust that fixed number of control points to the contour of the material at every frame.
+used a Neural Gas (NG) algorithm to adjust that fixed number of control points to the contour of the material at every frame.
 Using the coordinates of the finger, the sensed force and the coordinates of those control points through time, they train a Feedforward Neural Network (FFNN) to predict the contour of the material given the position of the finger and sensed force.
 We reproduced those steps and their training method on our training video.
-However, when we traid to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it got tied to the specific location and type of interaction it was trained with, while our physics based model had no problem at making predictions for new pushing actions.
-The results with this other method can be seen on \fref{fig:cretu_prediction}.
-
-Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.
+However, when we tried to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it became tied to the specific location and type of interaction it was trained with, while our physics based model had no problem making predictions for new pushing actions.
+The results with the GNGN method can be seen in \fref{fig:cretu_prediction}.
+
+%Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.
 However, even if we could change that algorithm, we would expect the system to have the same problems and predict the deformation at the middle of the material even if is being pushed somewhere else.
@@
 \hline
-\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN measured in number of frames.
-TP + FN = total number of frames.
-Final class assigned according to the performance of the models' predictions across the whole video.} \end{footnotesize}}
+\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN are measured in terms of numbers of frames.
+TP + FN = 299.
+The final class assigned according to the performance of the models' predictions across the whole video.
+See text for details.} \end{footnotesize}}
 \end{tabular}
 \end{table}
+
 % \begin{table}[!t]
@@
 \comment{
-\tref{tab:classperframe} shows the results for the classification task, frame per frame, for the videos that were used as training and test sets for the prediction task.
-When the sponge model performed better for a frame of the sponge video it is marked as a true positive, when plasticine does better it is marked as a false negative (since it didn't recognize it as a sponge).
-The same process is used for the other material.
-Since the concept was proved with only two materials, true positives and false negatives provide all the information.
-
-It calls for attention that in some of the experiments, like the one illustrated above, the performance of classification for the sponge on the training and test1 sets are the same.
+\tref{tab:classperframe} shows the results for the classification task, frame by frame for each video.
+When the sponge model performed better for a frame of the sponge video it is marked as a true positive (TP), when plasticine does better it is marked as a false negative (FN) ---since it didn't recognize it as a sponge.
+The analogous process is used for the plasticine.
+The total TP+FN=299 which is the total frames per movie.
+The overall classification for a movie can be determined by whether TP > FN or vice versa.
+Using this rule gives a perfect classification of the six movies.
+
+It should be noted that, in some of the experiments, the performance of classification for the sponge on the training and test1 sets are the same.
 We interpret that the mass-spring physics based model generalizes naturally when the force/deformation conditions are similar, even if the place for pushing is different.
 Between pushing in the middle and pushing closer to the border the qualitative behaviour of the elastic material was no so different for these models; unlike the data from the third set, where the distribution of mass and maximum force/deformation is more different from the ones seen on the training set of the model.
-It is interesting to confirm that the difference in predictions between the models is bigger here in favour of a better classification.
-
-It is also remarkable that, even though, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.
+It is interesting to confirm that the difference in predictions between the models is bigger here, resulting in better classification.
+
+It is also notable that, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.
 That is, the parameters for the sponge are not good at modeling the plasticine on the test sets, thus yielding better results at the classification task.
 This could happen because the plastic behaviour (that is, the permanent deformation of a spring) gets activated with stronger forces after the $yield$ limit is reached, which only appear when pushing the plasticine.
-With the small forces feedback by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour do not manifest.
+With the small forces fed back by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour are not manifest.
 }
@@
 The current search procedure doesn't always find good solutions for the plastic deformation model.
+A third extension is to extend the fairly trivial visual system to initialise mass-spring models for objects of abitrary shape.
+We propose using a Voronoi triangulation rather than a grid.
 But a more basic problem is that mass-spring models restrict the range of deformations that can be modelled.

### 2016-12-26 AnswersToComments-secondreview.tex
Added results of GNG for plasticine.  Gave format to answer to comments.

\documentclass[letterpaper,12pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}

\usepackage[top=1in, left=1.2in,right=1.2in]{geometry}
%\longindentation=0pt

\renewcommand*{\opening}[2]{\ifx\@empty\fromaddress
 \thispagestyle{firstpage}%
 {\raggedleft#2\par}%
 \else % home address
 \thispagestyle{empty}%
 {\raggedleft\begin{tabular}{l@{}}\ignorespaces
 \fromaddress \\*[2\parskip]%
 #2 \end{tabular}\par}%
 \fi
 \vspace{10\parskip}%
 {\raggedright \toname \\ \toaddress \par}%
 \vspace{2\parskip}%
 #1\par\nobreak}

%% --- Font

\renewcommand*{\familydefault}{\sfdefault}
\renewcommand*{\sfdefault}{pag} % Avant Garde

\usepackage{amssymb}
\def\labelitemii{$\blacktriangle$} % Windows no encuentra el símbolo solicitado
\renewcommand{\theenumii}{\alph{enumii}}

%% -- Logo
\usepackage{graphicx}

\author{Verónica Esther Arriola Ríos}

\begin{document}
\begin{letter}{\textbf{}}
\includegraphics[height=4cm, keepaspectratio=true]{escudo_UNAM.jpg}\vspace*{-4cm}

\address{Universidad Nacional Autónoma de México\\ Facultad de Ciencias,\\ Departamento de Matemáticas,\\ México, D.F.}
\telephone{5622 5426}

\opening{Dear Editors,}{\today}
\vspace{1cm}

Please find attached the pdf of our revised submission to the special issue on “Sensorimotor Contingencies for Cognitive Robotics”.

We would like to thank you and the three reviewers for timely and informative reviews, and for the comments, which have helped us to substantially improve our paper. In this revised version we have now included information on an implementation of a comparison model: a growing neural gas network. We show that it produces significantly poorer results on the prediction problem, and that it is thus unsuitable to tackle the classification problem we defined. In addition, we have carefully addressed all the reviewers’ concerns about the text. We detail below how we have addressed each reviewer point one by one.

We hope that you find the resulting paper much improved.

\vspace{1cm}

\signature{Verónica E. Arriola-Rios\\
Jeremy L Wyatt}

\closing{Yours sincerely,}

%%enclosure listing
%%\encl{}

\end{letter}

\textbf{Response to reviewers’ comments}

We would like to thank all the reviewers most sincerely for their effort in reading and commenting on our paper. The remarks helped us understand where we had gone wrong, and we have tried our best to incorporate all your requests and suggestions. Thank you.

\textbf{Reviewer 1}

\emph{In this revision, the authors solved the initial problems on article structure, citations, figures, and contents.}

We thank the reviewer for their confirmation that we have satisfied all their comments, and would like to thank them again for their remarks, which significantly improved our paper.

\textbf{Reviewer 2}

\emph{Page 1. Last Paragraph. There is an enumeration of novel contributions, you mentioned the First, and the Third, to avoid any confusion I would recommend to remark which is the second.}

We apologise, this was a typo. We have listed two main technical contributions.

\emph{Page 1-2. It is possible to infer the meaning of FEM, but in order to avoid any possible confusion you might indicate its meaning explicitly when the concept is first used.}

We have fixed this.

\emph{I find Fig 2, 3, and 4 difficult to read, however I don’t consider they should be necessarily changed.}

We agree this was difficult to interpret. To aid the reader we have added some text descriptions for the most important elements to make the diagram more readable.

\emph{At the end of subsection VII. B, when the authors refer to “all elastic and plastic terms” might be illustrative to add references to the corresponding equations.}

We have added a footnote with a reference to the section where all terms are summarised.

\emph{Within the Model Overview, it would be interesting if the authors would briefly highlight the advantage of having generative models for those who are not vastly familiarized with the topic.}

Thank you for this suggestion. We have added a short explanation within the section Model Overview. This explains that generative models are appealing because they allow us to solve both prediction and classification problems, whereas discriminative models are typically only suitable for classification.

\emph{When interpreting results what are the consequences when triangles in the border of the mesh overlap each other? Is there any consequence that might be significant to the result interpretation?}

We have added some comments to clarify these points in the subsection “Additional Geometric Constraints”.

\emph{In section V, during the training the segmentation assumes that the color of the object is known. Therefore, when you say that the algorithm could learn and classify previously unknown materials, at least the machine must partially know the material.}

In fact, no assumption about the colour of the object is made. The object is simply marked with a bounding box in the first frame. This is used to determine the dominant hue. This hue is then used to distinguish edge pixels bounding the object from those gathered elsewhere in the scene. We have added this to the explanation of the first step of the algorithm in Section V.A.

Also the authors mention that the simplified active contours approach is not the most accurate, What are the consequences of this approximation?

The errors of the approximation are small, and are relatively small in comparison with the errors introduced by prediction with the mass-spring model. We added a note to this effect in section V.A  on page 8.

Regarding to the chosen integration time step, does the results vary considerably when 0.01 is considered?

We used a time step of 0.1. We have previously experimented with various time step parameters, and found that a time step of 0.01 didn’t improve the quality of prediction, but increased computation time.  We have added a comment on this point to section V.B.

\emph{In section VII.E, Table III, is there any special reason to have the same result for the training and test1 sets  in the case of the sponge? It would be better if the authors offer deeper insights regarding the interpretation of the figures shown in Table III.}

Thanks for raising this point, we needed to explain this more clearly.  To enable this Table III has been shifted to section VIII, where the discussion has been expanded to cover these issues.

\textbf{Reviewer 3}

\emph{Reading the abstract and the introduction, I was expecting a framework in which the robot creates a representation of the environment (in this case the two objects) while it is interacting with it (looking at and pushing the object).}

We are sorry that this set false expectations. We have altered the wording in several places to make the reader aware that we are not tackling this problem on-line, but off-line. We have also read and noted the references you mention in the introduction.

\emph{While the authors recognize the issue of the on-line learning in the introduction (page 1, col 2, line 35), they do not address that in the paper.}

From your comments, we believe that mentioning on-line learning, while important, was misleading in terms of our paper, so we have withdrawn mention of it.

\emph{Also they refer the problem of getting access to partial views of the objects but it seems that in this paper the object is viewed from a convenient point of view.}

We have added text to make clear the restriction in our study of using a single view chosen by the user.

Finally, the word “learning from data” suggests to me that the model acquires a representation of the environment based on some kind on non-parametric model, such as neural network. However, the models seem carefully designed to represent the object deformation and the stress-gain diagram. Even if the difference between learning and calibration is subtle, in my opinion, the use of the word “learning” creates an expectation that is not fulfilled in the rest of the paper.

We have changed the wording in the abstract to note that we calibrate the models from data. We hope this makes it clearer where on the spectrum of parameter tuning to full autonomous learning we lie. We now emphasise that we perform off-line learning, we sometimes refer to it as calibration, and we mention that we don’t address the issues of on-line learning. We hope this changes the tone of the introduction enough so as not to mislead.

\emph{What happen if we add a new material? Do we need to add a new feature space?}

If the object is plastic or elastic then we don’t need to add a new feature space, but we would be wise to calibrate a new model. We have added a comment to this effect in the conclusion.

\emph{I would add a figure like Fig. 9a at this point. In this way the reader can have an idea of the setup, how vision acquire information and how the mass-spring model depends on the visual information itself. For example, it might be interesting to combine Fig 9a and Fig 2 in single sketch of the model.}

Thank you for this suggestion. We have prepared a new version of Figure 2, as you suggest. This incorporates elements from the old Figure 2 and Figure 9.

\emph{A consideration about the organisation of the following sections. Given that the framework consists of two models, force prediction and shape prediction, I would expect section III.1 to be Force prediction and III.2 to the shape prediction (or vice-versa). However, the organization of Sections IV, V, VI, in my opinion is misleading.}

We have reorganised the paper as you request.

\emph{Finally, I would start describing the algorithm from the visual system. It seems that the visual system has a very important role in this paper but it is not highlighted.}

We have slightly expanded and moved the section on the visual system as part of the paper reorganization. We also spend a little more space on vision in the model overview section.

\emph{I would like to know how the particles are placed inside the object. What happens if we change the size of the object? Do we need to change the number of particles? What happen if the object is not rectangular?}

We describe the way that the particles are placed inside the object in detail in Section VII.A on p.10. We specify a total number of cells in the mass-spring model, and then the particles (which are the nodes of the mass-spring model) are then spaced evenly in a grid. If we change the size of the object the method will allocate particles more widely spaced for a given grid size. No version of the algorithm has been specified for non-rectangular, we suggest that a Voronoi triangulation would be the way to proceed. We mention this in the conclusion.

\emph{Eq. 19. I really do not understand how the force could be proportional to the velocity. This point should be clarified. My understanding is that the authors assumed the acceleration (a) to be constant during the interval h, so that h*F = h*a*m => F = v*m/h. Given that h is a constant, the force (F), the velocity (v) or the mass (m) are recovered with a scale factor. Is that correct? If so, what is the difference between this approach and the quasi-static approach proposed by Frank. Please, explain.}

The essential difference is that Frank models the system as being quasi-static, i.e. that acceleration is not only constant, but is zero. In contrast we model the acceleration as non-zero and non-constant. One point of possible confusion is that we are using a numerical method for approximate solution of the differential equations that model each mass particle. This numerical method simply assumes that between two successive points in time that acceleration is constant. Over multiple time steps, however, the acceleration does change.

Regarding the physics, the corresponding math would be F = c m v  =>  a = v  =>  (a = dv/dt) = v which can only be solved with an exponential. Thus the constraint a=v forces us to solve with an exponential, and has the benefit of preventing oscillations of the springs in the model.  This explanation has been added to the “Integration scheme” section.

\emph{This section is quite long and I think it could be simplified notably. The authors could simply state the energy $E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2]$, where N is the number of constraints. There are several constraints to preserve the length, the area, etc. and the force is given as negative gradient of the potential energy with respect to the position of the particle. Details about all the equation could be reported in an appendix to make the paper more readable. About the equation, personally I don’t like the symbol composed of multiple letters such as Area, forcemag, forcedir, etc.}

We have shortened the section as requested.

\emph{Do the authors use just one video for the training and two video for the testing? It seems that the dataset is quite small to have a systematic test of the algorithm. Even if there are only two objects, the authors should test different motions of the finger (i.e. pushing velocity, penetration, …)}

We accept that the test data set is rather small. We believe however that it is sufficient to show through a comparison with the GNGN method that our approach is considerably better than the state of the art when it comes to generalising deformation predictions with respect to push position. Also, we should note that the data needed for successful training is also very small: just one movie of each object. We will gather additional data, but our current plan is to use this for future work.

\emph{I don’t see the results of the classification. I would expect a confusion matrix to show how many times the sponge was classified as sponge and as plasticine and vice-versa. Table III seems to show the results about the classification of the pixels.}

Table III was moved to the discussion section were this point was clarified in the caption. The results show the classification not of the pixel, but of the frame. The classification of a whole movie is described in the text.

\emph{Moreover, I would like to see a comparison against a baseline approach. They could some of the models proposed in the literature or some simple models. For example, the authors could assume that the material “disappears” when the finger penetrates the object and report the F-Score of this method. Alternatively, they could use a simple model with only one spring between the finger and the base of the object.}

We have implemented a Growing Neural Gas Network (GNGN) previously published by Cretu et al. We have added additional results in a comparison of this method with our own. This is now described in section VII.D. The very poor generalization performance of the GNGN on the prediction problem is shown in a new Fig 17.

\emph{Page 12. Col 2. Line 3. Please, can you clarify the hand tuning? It seems that the genetic algorithm is not the best way to find the parameters? Have you tried to use other method like gradient descent? Is it possible to use the model to mass-spring model to predict the force exerted on the finger? In that case, the authors could use this prediction error as fitness function and use the gradient descent to find the parameters.}

The hand tuned solutions we tried were obtained by slightly modifying some of the parameters of the solutions proposed by the algorithm.  However, the f-score of the solution chosen by the search algorithm is better than the f-score of the solution that was hand tuned, that is why the algorithm does not select it. This explanation was added to section IX.A.

Gradient descent was not used because the space of parameters for a mass-spring model can offer good solutions in very separated regions.  Therefore, gradient descent would not make a proper exploration of the space and would get stuck on local minima quite quickly.

It could be possible to estimate the reaction force by adding the total forces on the particles in touch with the finger.  At some point we did check if they were similar to the measured forces and found out that for some selected sets they are similar, but we didn’t try to use this criterion as evaluation function.  The suggestion was added in the discussion section (IX.A).

Other comments:

\emph{Fig. 1 is not referenced in the paper.}

We have decided that this figure added little, hence it was not referenced naturally anywhere in the text. We have thus decided to remove it to save space, since other requested changes have added considerable length.

\emph{Page 2, Col 1, Line 17. Section VI is not mentioned.}

We have fixed this.

\emph{Page 2. Col 2, Line 5: FEM is not defined. Finite-element models?}

We have defined this.

\emph{Fig 2. Instead of use symbols, I would write Vision, strain, force, etc.}

Thank you, this chimes with the comments of Reviewer 2, we have fixed this in Figures 1, 2 and 3.

\emph{Page 3, Col 1, line 57. Cretu et al. -> cite [10]?}

We have added this reference.

\emph{Page 8. Col 2, Line 27. Linear snake = linear spline?}

The term “linear spline” is associated with the geometric curve, while “linear snake” to the tracking algorithm that makes use of the linear spline.  A clarification was added in section IV.A and the use of both terms was reviewed to make sure it was consistent with their definition.  Thank you for noticing.

\emph{Algorithm 5. I would specify that the F-Score is the fitness function of the algorithm. Moreover, it would be nice to know the values used for the parameters of the algorithm (perhaps using a table).}

This information has been added to section VIII. C., the values for the parameters are on Table III.

\emph{Page 10. Col 1. Line 45. Please specify that there are two objects, a sponge and plasticine, with a rectangular shape and so on.}

We have added this information.
\end{document}

### 2016-12-26 twostagedo-rev-Dec-16.tex
Added results of GNG for plasticine.  Gave format to answer to comments.

@@
 We employ a Canny edge detector, filtering out irrelevant edges using the colour of the manipulated object.
-Then we fit the outline and track using a linear spline.
+Then we fit the outline and track using a linear snake.
 This is also used to initialise the mesh for the spring mass system.}
@@
 The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
-Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
+\comment{We call our algorithm for tracking a \emph{linear snake}, because}  this implementation is inspired by the theory of active contours, however it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
 \comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.
@@
 Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.
-Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping}, after some manipulation we get:
-\begin{align}
- \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
- & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
-\end{align}
-where $k_L$ is the linear spring constant and $k_d$ the damping constant.
+Details about the final form of this and following equations are in \aref{ap:formulas}.
 
@@
 found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.
-Therefore the force $\vec{F}^i_A$ is derived as follows:
+Therefore the force $\vec{F}^i_A$ is derived from the constraint and \eref{eq:EForce}:
 \begin{align}
- C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
- Area &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
- \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( Area-A_0 \right) \nonumber \\
- & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
+ C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0}
 \end{align}
@@
 
-It is possible to rewrite \eref{eq:area_force} in a more geometrically intuitive manner by rewriting the direction of the gradient according to Morris' geometric derivation \cite{Morris2008}, but keeping the area normalisation constant\footnote{This substitution is valid because both are expressions for the gradients of the area of the triangle formed by the three verticies.}.
-It becomes evident that the forces pull along the heights of the triangles [\fref{fig:forces}(b)].
-Then the magnitude of the force is given by:
-\begin{equation}
- forcemag_A(\vec{p}_i)= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0^2}
-\end{equation}
-
-While the rest becomes:
-\begin{align}
- \vec{F}_A(\vec{p}_i) & = k_A \cdot forcemag_A(\vec{p}_i) \cdot forcedir_A(\vec{p}_i) \nonumber \\
- forcedir_A(\vec{p}_i) & = \dfrac{\vec{F}_A(\vec{p}_i)}{|\vec{F}_A(\vec{p}_i)|} = \dfrac{areagradient(\vec{p}_i)}{|areagradient(\vec{p}_i)|} \nonumber \\
- areagradient(\vec{p}_i) & = (\vec{p}_i-\vec{p}_j)- \nonumber \\
- & \left( (\vec{p}_k-\vec{p}_j) \cdot \dfrac{(\vec{p}_k-\vec{p}_j) \cdot (\vec{p}_i-\vec{p}_j)}{(\vec{p}_k-\vec{p}_j) \cdot (\vec{p}_k-\vec{p}_j)} \right)
-\end{align}
-
 \subsubsection{Preservation of Angles}
 The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.
-For this reason we included a new term that enforces the preservation of the original angles of the triangles.
+For this reason we included a new term that enforces the preservation of the original angles of the triangles, \comment{which is not a valid state since the mesh models the surface of a 3D object}.
 [\fref{fig:forces}(c)] gives an idea of how these forces look.
@@
 Therefore, it helps to recover a similar triangle, but it may produce tiny or very big triangles if it is not accompanied by some of the other terms that tend to restore the original dimensions, in addition to the angles.
-The force is derived from:  %An additional line in the code also forces an inverted angle to recover its original orientation.
-\begin{align}
- F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
- \dfrac{\partial \varphi}{\partial p_i} & = \dfrac{\partial}{\partial p_i} arccos \left(\frac{(\vec{p}_j-\vec{p}_i)\cdot(\vec{p}_k-\vec{p}_i)}{\left\|\vec{p}_j-\vec{p}_i\right\|\left\|\vec{p}_k-\vec{p}_i\right\|}\right) \displaybreak[0] \\
-\end{align}
-
-The force is:
-\begin{align}
- F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
- \dfrac{\partial \varphi}{\partial p_i}(p_i) & =
-    \dfrac{1}{d_{ji}d_{ki}\sqrt{1-\left[\dfrac{pp}{(d_{ji}) (d_{ki})}\right]^2}} \nonumber \\
- & \left\{\left[1-\dfrac{pp}{d_{ki}^2}\right](p_k-p_i)+\left[1-\dfrac{pp}{d_{ji}^2}\right](p_j-p_i) \right\} \nonumber \\
- pp(p_i) & =(p_j-p_i)\cdot(p_k-p_i) \nonumber \\
- d_{ji}(p_i) & = \left\| p_j-p_i \right\| \nonumber \\
- d_{ki}(p_i) & = \left\| p_k-p_i \right\|
-\end{align}
-
+
+\vspace*{1em}
 The terms for the preservation of length and angle model elastic deformations.
@@
 \end{align}
-which implies the following assumptions:
+which implies the following assumption:
 \begin{align}
@@
 \end{align}
-this equation can only be satisfied with an exponential, thus avoiding the oscillatory behaviour of springs in favour of forcing critically damped springs to model the elastic materials.
+this equation can only be satisfied with an exponential function, thus avoiding the oscillatory behaviour of sinusoid solutions for springs, in favour of forcing critically damped springs to model the elastic materials.
 }
@@
 At this point no prediction of the force was included, rather, the reaction forces measured with the force sensor were used.
+\comment{The F-score was used as fitness function.
+The values used for the parameters of \alref{alg:genetic} are summarised in \tref{tab:gen_params}, also the minimum and maximum limits for the values of parameters for the mass-spring model.}
+
+\begin{table}[!t]
+ \renewcommand{\arraystretch}{1.3}
+ \caption{Parameters for the Genetic Algorithm}
+ \label{tab:gen_params}
+ \centering
+\comment{
+\begin{tabular}{cc}
+\hline
+ \bfseries Parameter & \bfseries Value/Range \\  \hline
+ \multicolumn{2}{c}{Genetic Algorithm} \\ \hline
+ M & 30 \\
+ $\alpha = \beta = \gamma = \zeta$  & 5 \\
+ $\sigma_{max}$                     & 1000.0 \\
+ $\sigma_{min}$                     & 10.0 \\
+ $P(s)$                             & 0.05 \\ \hline
+ \multicolumn{2}{c}{Elastic Parameters } \\ \hline
+ $k_d$                              & $[0.00001, 2]$ \\
+ $K_L$                              & $[0.00001, 10000]$ \\
+ $K_A$                              & $[0.00001, 10000]$ \\
+ $K_{\varphi}$                      & $[0.00001, 100]$ \\
+ $max(F)$                           & $[5, 100000]$ \\ \hline
+ \multicolumn{2}{c}{Plastic Parameters } \\ \hline
+ $\alpha$                           & $[0.00001, 0.9]$ \\
+ $yield$                            & $[0.00001, 1.0]$ \\
+ $creep$                            & $[0.00001, 1.0]$ \\ \hline
+\end{tabular}
+
+\vspace*{1em}
+* Where $P(s)$ is the probability of using a special value, like $0.0$, for the spring constants.
+}
+\end{table}
+
 \fref{fig:performance} shows how the quality of the simulation throughout the whole video improved with each generation.
@@
 \caption{Neural network approach fits quite well the training data set, but can not generalise to new pushing actions.
-The green line shows the border used as ground truth, while blue lines show prediction from the FFNN.}\label{fig:cretu_prediction}
+The red line shows the border used as ground truth in the training set, while yellow lines show prediction from the FFNN.}\label{fig:cretu_prediction}
+\end{figure}
+\begin{figure}[!t]
+\centering
+\includegraphics[width=3.5in]{plasticine__result.jpg}
+\caption{GNG segmentation didn't work with a shadowed material (video of plasticine).
+This image shows the segmentation results previous to the application the median filter, which was applied to perform tracking.
+Results also do not generalise to unseen pushing actions.
+The red line shows the border used as ground truth in the training set, while yellow lines show prediction from the FFNN.}\label{fig:cretu_prediction}
 \end{figure}
@@
 Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
-Better simulations have been obtained by hand tuning some of the parameters.
-%\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.
-The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}
+Better simulations have been obtained by hand tuning some of the \comment{best sets of parameters obtained by the genetic algorithn.
+This is not due to an error of the search algorithm, but due to a difference between what humans visually qualify as a good solution and what the evaluation function does.
+For example: humans would prefer the solution shown in \fref{fig:simulation_plasticine_hand}, because this model of the plasticine does not tend to recover its shape, as the one selected by the computer does.
+However, the f-score of the solution chosen by the computer is better.
+The similarity between the sensed force and the forces that the mass-particles produce on the finger could be added as a second criteria to evaluate the models.
+Even though we found that the automatic calibration procedure sometimes produces models were this difference is small, we have left for future work the addition of this criterion to the evaluation function.}
 
@@
 %
+\appendix[Forces Derived from Preservation Constraints]
+\label{ap:formulas}
+
+\subsubsection{Preservation of Length}
+
+Substituting \eref{eq:clength} in \eref{eq:Fdamping}, after some manipulation we get:
+\begin{align}
+ \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
+ & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
+\end{align}
+where $k_L$ is the linear spring constant and $k_d$ the damping constant.
+
+\subsubsection{Preservation of Area}
+
+The force $\vec{F}^i_A$ is derived as follows: 
+\begin{align}
+ C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
+ A &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
+ \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( A-A_0 \right) \nonumber \\
+ & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
+\end{align}
+where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.
+
+It is possible to rewrite \eref{eq:area_force} in a more geometrically intuitive manner by rewriting the direction of the gradient according to Morris' geometric derivation \cite{Morris2008}, but keeping the area normalisation constant\footnote{This substitution is valid because both are expressions for the gradients of the area of the triangle formed by the three verticies.}.
+It becomes evident that the forces pull along the heights of the triangles [\fref{fig:forces}(b)].
+Then the magnitude of the force is given by:
+\begin{equation}
+ \lVert \vec{F} \rVert_A(\vec{p}_i)= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0^2}
+\end{equation}
+
+While the rest becomes:
+\begin{align}
+ \vec{F}_A(\vec{p}_i) & = k_A \cdot \lVert \vec{F} \rVert_A(\vec{p}_i) \cdot forcedir_A(\vec{p}_i) \nonumber \\
+ forcedir_A(\vec{p}_i) & = \dfrac{\vec{F}_A(\vec{p}_i)}{|\vec{F}_A(\vec{p}_i)|} = \dfrac{\nabla A(\vec{p}_i)}{|\nabla A(\vec{p}_i)|} \nonumber \\
+ \nabla A(\vec{p}_i) & = (\vec{p}_i-\vec{p}_j)- \nonumber \\
+ & \left( (\vec{p}_k-\vec{p}_j) \cdot \dfrac{(\vec{p}_k-\vec{p}_j) \cdot (\vec{p}_i-\vec{p}_j)}{(\vec{p}_k-\vec{p}_j) \cdot (\vec{p}_k-\vec{p}_j)} \right)
+\end{align}
+
+\subsubsection{Preservation of Angles}
+
+The force is derived from:  %An additional line in the code also forces an inverted angle to recover its original orientation.
+\begin{align}
+ F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
+ \dfrac{\partial \varphi}{\partial p_i} & = \dfrac{\partial}{\partial p_i} arccos \left(\frac{(\vec{p}_j-\vec{p}_i)\cdot(\vec{p}_k-\vec{p}_i)}{\left\|\vec{p}_j-\vec{p}_i\right\|\left\|\vec{p}_k-\vec{p}_i\right\|}\right) \displaybreak[0] \\
+\end{align}
+
+The force is:
+\begin{align}
+ F_\varphi(p_i) & = k_\varphi(\varphi - \varphi_0)\dfrac{\partial \varphi}{\partial p_i} \\
+ \dfrac{\partial \varphi}{\partial p_i}(p_i) & =
+    \dfrac{1}{d_{ji}d_{ki}\sqrt{1-\left[\dfrac{pp}{(d_{ji}) (d_{ki})}\right]^2}} \nonumber \\
+ & \left\{\left[1-\dfrac{pp}{d_{ki}^2}\right](p_k-p_i)+\left[1-\dfrac{pp}{d_{ji}^2}\right](p_j-p_i) \right\} \nonumber \\
+ pp(p_i) & =(p_j-p_i)\cdot(p_k-p_i) \nonumber \\
+ d_{ji}(p_i) & = \left\| p_j-p_i \right\| \nonumber \\
+ d_{ki}(p_i) & = \left\| p_k-p_i \right\|
+\end{align}
 
@@
 \fi
+
+\balance
 
@@
 
-\balance
 %

### 2016-12-26 twostagedo-rev-Dec-16.tex
Fixed missing reference to figure.

@@
 However, when we tried to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it became tied to the specific location and type of interaction it was trained with, while our physics based model had no problem making predictions for new pushing actions.
-The results with the GNGN method can be seen in \fref{fig:cretu_prediction}.
+The results with the GNGN method can be seen in \fref{fig:cretu_prediction_sponge} and \fref{fig:cretu_prediction_plasticine}.
 
@@
 \caption{Neural network approach fits quite well the training data set, but can not generalise to new pushing actions.
-The red line shows the border used as ground truth in the training set, while yellow lines show prediction from the FFNN.}\label{fig:cretu_prediction}
+The red line shows the border used as ground truth in the training set, while yellow lines show prediction from the FFNN.}\label{fig:cretu_prediction_sponge}
 \end{figure}
@@
 Results also do not generalise to unseen pushing actions.
-The red line shows the border used as ground truth in the training set, while yellow lines show prediction from the FFNN.}\label{fig:cretu_prediction}
+The red line shows the border used as ground truth in the training set, while yellow lines show prediction from the FFNN.}\label{fig:cretu_prediction_plasticine}
 \end{figure}

### 2017-01-24 twostagedo-rev-final.tex
Added source of final version.

--- 2016-12-26 twostagedo-rev-Dec-16.tex
@@
 \begin{abstract}
-In this paper we present a multi-modal framework for \comment{offline} learning of generative models of object deformation under robotic pushing.
+In this paper we present a multi-modal framework for offline learning of generative models of object deformation under robotic pushing.
 The model is multi-modal in that it is based on integrating force and visual information.
-The framework consists of several sub-models that are independently \comment{calibrated} from the same data.
+The framework consists of several sub-models that are independently calibrated from the same data.
 These component models can be sequenced to provide many-step prediction and classification.
 When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of modules for different materials are compared so as to classify the unknown material.
-Our approach, which consists of \comment{offline learning and combination of} multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
+Our approach, which consists of offline learning and combination of multiple models, goes beyond previous techniques by enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.
 While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered a way to achieve them all, and none has offered classification from a generative model.
@@
 In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data.
-\comment{In contrast, robots must learn model parameters from data, in reasonable amounts of training time.
-This includes learning from the limited sensing typically available, such as training data gathered from a specific view(s) of an object, and force-torque monitoring only at contact points.} Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.
-\comment{In this paper we address some of these---we learn from real visual and force-torque data---while simplifying others---we set a convenient fixed viewing location and perform off-line rather than on-line learning.}
+In contrast, robots must learn model parameters from data, in reasonable amounts of training time.
+This includes learning from the limited sensing typically available, such as training data gathered from a specific view(s) of an object, and force-torque monitoring only at contact points.
+Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.
+In this paper we address some of these---we learn from real visual and force-torque data---while simplifying others---we set a convenient fixed viewing location and perform off-line rather than on-line learning.
 
@@
 
-\comment{There are two novel technical contributions.
+There are two novel technical contributions.
 First, we show how the learning problem can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces.
@@
 To limit the scope of the paper we develop this framework using offline learning methods.
-We do not address the issue of online learning \cite{worgotter09,antonelli14}, leaving this for future work.}
+We do not address the issue of online learning \cite{worgotter09,antonelli14}, leaving this for future work.
 
@@
 
-\comment{These technical contributions} enable us to learn models of sensorimotor contingencies, and then to use these as components to build different machines.
+These technical contributions enable us to learn models of sensorimotor contingencies, and then to use these as components to build different machines.
 This approach enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed.
@@
 He sketches the inclusion of plastic deformation.
-Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a \comment{finite element model (FEM)}.
+Morris and Salisbury \cite{Morris2008} developed a method to calibrate Teschner's model automatically with respect to a finite element model (FEM).
 The  search begins with a uniform random sampling of possible values for the elasticity constants, which are latter modified using adaptive simulated annealing to minimize the difference between the deformations in the FEM mesh and those in the mass-spring mesh.
@@
 \includegraphics[width=3.5in]{figures/learning.png}%}
-\caption{\comment{The visual tracking and force measurement system subserves the offline learning system.
-The offline learning (or calibration) method comprises two learned models of sensorimotor contingencies.} Both are learned from the same data.
+\caption{The visual tracking and force measurement system subserves the offline learning system.
+The offline learning (or calibration) method comprises two learned models of sensorimotor contingencies.
+Both are learned from the same data.
 The first (FP) learns to predict the force experienced for some motion and material.
@@
 
-\comment{The framework consists of four separate systems: a visual tracking and force measurement system, an offline learning (or model calibration) system, a prediction system, and a classification system.
+The framework consists of four separate systems: a visual tracking and force measurement system, an offline learning (or model calibration) system, a prediction system, and a classification system.
 
@@
 Then we fit the outline and track using a linear snake.
-This is also used to initialise the mesh for the spring mass system.}
+This is also used to initialise the mesh for the spring mass system.
 
@@
 
-\comment{In machine learning there is division between discriminative and generative models.
+In machine learning there is division between discriminative and generative models.
 Discriminative models are powerful for classification, but cannot be used to regenerate the data, or to generate new imagined data.
@@
 But we show how this can be used not just for prediction, but also for material classification.
-Thus a generative model is appealing because it allows us to solve both the prediction and classification problems.}
+Thus a generative model is appealing because it allows us to solve both the prediction and classification problems.
 
@@
 The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.
-\comment{We call our algorithm for tracking a \emph{linear snake}, because}  this implementation is inspired by the theory of active contours, however it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
-\comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.
+We call our algorithm for tracking a \emph{linear snake}, because this implementation is inspired by the theory of active contours, however it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.
+This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.
+This is explained below.
 The tracking of the deformable target in the ground truth has two stages:
@@
 The Canny algorithm is used to extract the edges of the image.
-Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.
-Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.
+Since the object of interest has a distinctive colour, its characteristic hue range is extracted from the area enclosed by the initial contour.
+Then the hue of the pixels on every edge are used to keep only edges that could belong to it.
 If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
@@
 \begin{align}
-F = m \dfrac{\partial \vec{p}}{\partial t} \label{eq:newton}
+F = m \dfrac{\partial^2 \vec{p}}{\partial t^2} \label{eq:newton}
 \end{align}
-where $t$ is time.  However, it is possible to deviate from this equation if we need to model a different behaviour. %are interested in is numerical accuracy, as it will be done here.
+where $t$ is time and $\vec{p}$, the position.
+However, it is possible to deviate from this equation if we need to model a different behaviour.
+%are interested in is numerical accuracy, as it will be done here.
 
@@
 \begin{align}
-F + F_{ext} = m \dfrac{\partial \vec{p}}{\partial t}
+F + F_{ext} = m \dfrac{\partial^2 \vec{p}}{\partial t^2}
 \end{align}
@@
 Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.
-\comment{Since the location of the masses and direction of the springs will bias the deformation response of the object, it is important to distribute the masses uniformly and place the springs with a symetric arrangement for an isotropic material.
+Since the location of the masses and direction of the springs will bias the deformation response of the object, it is important to distribute the masses uniformly and place the springs with a symetric arrangement for an isotropic material.
 This placement can be done with an algorithm that takes into account the symmetries of the object at the begining of the experiment.
-If the material gets permanently deformed a properly calibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}
+If the material gets permanently deformed a properly calibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.
 
@@
 \subsubsection{Preservation of Angles}
-The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.
-For this reason we included a new term that enforces the preservation of the original angles of the triangles, \comment{which is not a valid state since the mesh models the surface of a 3D object}.
+The previous terms can not do anything to restore the mesh if triangles get flipped during deformation, which is not a valid state since the mesh models the surface of a 3D object.
+For this reason we included a new term that enforces the preservation of the original angles of the triangles.
 [\fref{fig:forces}(c)] gives an idea of how these forces look.
-Here the energy depends on the difference \comment{%\sout{of the angles between adjacent edges.}
-between the current angle, between adjacent edges, and the angle between them at the equilibrium position.} Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
+Here the energy depends on the difference between the current angle, between adjacent edges, and the angle between them at the equilibrium position.
+Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
 \begin{align}
@@
 \label{sec:ischeme}
-Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations, we made the force proportional to the velocity:
-\comment{
+Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equations, we made the force proportional to the velocity:
 \begin{align}
- F(t) = \vec{p}
+ F(t) = m\dfrac{\partial \vec{p}(t)}{\partial t} = m \vec{v}(t)
 \end{align}
@@
 \begin{align}
- \dfrac{\partial \vec{p}}{\partial t} = \vec{p}
+ \vec{a}(t) = \dfrac{\partial \vec{v}(t)}{\partial t} = \vec{v}(t)
 \end{align}
-this equation can only be satisfied with an exponential function, thus avoiding the oscillatory behaviour of sinusoid solutions for springs, in favour of forcing critically damped springs to model the elastic materials.
-}
+this equation of the form $f=f'$ can only be satisfied with an exponential function $f(x)=e^x$, thus avoiding the oscillatory behaviour of sinusoid solutions for springs, in favour of forcing critically damped springs to model the elastic materials.
 
@@
 Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.
-\comment{All of these situations are inadequate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
+All of these situations are inadequate since this mesh is used to model one face of a 3D object.
+It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.
 We achieve this through geometric checking subroutines.
@@
  \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed more than half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ of this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.
-This is enough for most cases, but triangles in the border of the mesh can still overlap each other \comment{which would render a bad numerical approximation to the behaviour of a surface}.
+This is enough for most cases, but triangles in the border of the mesh can still overlap each other which would render a bad numerical approximation to the behaviour of a surface.
 
@@
   \item The integration time step $\Delta t$.
-\textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well; \comment{however, since they didn't improve stability or accuracy in a visible way, but made the process much slower, they were disregarded as impractical}.)}
+\textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well; however, since they didn't improve stability or accuracy in a visible way, but made the process much slower, they were disregarded as impractical.)}
  \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
@@
 A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].
-\comment{This set up allowed us to study the effect of using an automatically calibrated physics based model as model for the material, while using 2D meshes, even though the same principles can be used if we use a 3D point cloud and a 3D mesh.
+This set up allowed us to study the effect of using an automatically calibrated physics based model as model for the material, while using 2D meshes, even though the same principles can be used if we use a 3D point cloud and a 3D mesh.
 To generate the initial mesh for the mass-spring model a routine placed the nodes for the masses evenly spaced on a rectangular grid, given the number of squares we wanted per side.
 We made experiments with grids of $3 \times 2, 10 \times 5$ and $20 \times 10$ squares.
-To alleviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}
+To alleviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.
 
@@
 On the other hand, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that more closely resemble the ground truth, thus obtaining better marks.
-From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them\footnote{\comment{The full list of terms is summarised at the end of \sref{sub:EvolAlg}.}}.
+From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them\footnote{The full list of terms is summarised at the end of \sref{sub:EvolAlg}.}.
 
@@
 At this point no prediction of the force was included, rather, the reaction forces measured with the force sensor were used.
-\comment{The F-score was used as fitness function.
-The values used for the parameters of \alref{alg:genetic} are summarised in \tref{tab:gen_params}, also the minimum and maximum limits for the values of parameters for the mass-spring model.}
+The F-score was used as fitness function.
+The values used for the parameters of \alref{alg:genetic} are summarised in \tref{tab:gen_params}, also the minimum and maximum limits for the values of parameters for the mass-spring model.
 
@@
  \centering
-\comment{
 \begin{tabular}{cc}
@@
 * Where $P(s)$ is the probability of using a special value, like $0.0$, for the spring constants.
-}
 \end{table}
@@
 
-\comment{We also compared the generalisation capabilities of our model against the system proposed in \cite{Cretu2012}.
+We also compared the generalisation capabilities of our model against the system proposed in \cite{Cretu2012}.
 This system uses a Growing Neural Gas (GNG) Network to calibrate a colour segmentation algorithm.
@@
 The results with the GNGN method can be seen in \fref{fig:cretu_prediction_sponge} and \fref{fig:cretu_prediction_plasticine}.
-
-%Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.
-However, even if we could change that algorithm, we would expect the system to have the same problems and predict the deformation at the middle of the material even if is being pushed somewhere else.
-}
 
@@
 Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.
-Better simulations have been obtained by hand tuning some of the \comment{best sets of parameters obtained by the genetic algorithn.
+Better simulations have been obtained by hand tuning some of the best sets of parameters obtained by the genetic algorithn.
 This is not due to an error of the search algorithm, but due to a difference between what humans visually qualify as a good solution and what the evaluation function does.
@@
 The similarity between the sensed force and the forces that the mass-particles produce on the finger could be added as a second criteria to evaluate the models.
-Even though we found that the automatic calibration procedure sometimes produces models were this difference is small, we have left for future work the addition of this criterion to the evaluation function.}
+Even though we found that the automatic calibration procedure sometimes produces models were this difference is small, we have left for future work the addition of this criterion to the evaluation function.
 
@@
 \hline
-\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN are measured in terms of numbers of frames.
+\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}* TP and FN are measured in terms of numbers of frames.
 TP + FN = 299.
 The final class assigned according to the performance of the models' predictions across the whole video.
-See text for details.} \end{footnotesize}}
+See text for details.
+\end{footnotesize}}
 \end{tabular}
@@
 
-\comment{
 \tref{tab:classperframe} shows the results for the classification task, frame by frame for each video.
@@
 With the small forces fed back by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour are not manifest.
-}
 
@@
 Second, we have attempted the automatic calibration of a simulation of plasticity.
-\comment{If a new material were added, provided it were elastic or plastic, it would be possible to calibrate a model for that material, independent of the existing calibrated models.} Third, we use regression to predict the reaction forces in the force prediction model.
+If a new material were added, provided it were elastic or plastic, it would be possible to calibrate a model for that material, independent of the existing calibrated models.
+Third, we use regression to predict the reaction forces in the force prediction model.
 Fourth, we include a new energy term for the preservation of angles in a mass-spring model of shape deformation.
@@
 While the framework presented here is promising, we regard the actual deformation model representation as an open research problem.
-\comment{Finally, all the work here is for off-line learning, from a carefully chosen viewpoint.
-The issue of online learning of these representations, including how to explore so as to learn, is an open issue.}
+Finally, all the work here is for off-line learning, from a carefully chosen viewpoint.
+The issue of online learning of these representations, including how to explore so as to learn, is an open issue.
 

### 2017-01-25 twostagedo-rev-final.tex
Constraints...

@@
 \subsubsection{Preservation of Area}
-This constraint is applied per triangle in the mesh.  Since Teschner and Morris
-found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.
-Therefore the force $\vec{F}^i_A$ is derived from the constraint and \eref{eq:EForce}:
+The constraint for the potential energy is given by:
 \begin{align}
@@
 \end{align}
-where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.
+where $A_0$ is the initial area of the triangle.
+This constraint is applied per triangle in the mesh.
+Since Teschner and Morris found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.
+Therefore the force $\vec{F}^i_A$ is derived from the constraint and \eref{eq:EForce} as shown in the appendix.
 
@@
 Here the energy depends on the difference between the current angle, between adjacent edges, and the angle between them at the equilibrium position.
-Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
+The constraint for this potential energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
 \begin{align}
- E_\varphi(\varphi) & = \frac{1}{2}k_\varphi(\varphi - \varphi_0)^2 \\
+ %E_\varphi(\varphi) & = \frac{1}{2}k_\varphi(\varphi - \varphi_0)^2 \\
+ C &= \varphi - \varphi_0 \\
  \varphi(p_i,p_j,p_k) & = \arccos \left( \dfrac{(p_j - p_i) \cdot (p_k - p_i)}{ \left\| p_j - p_i \right\| \left\|p_k - p_i \right\|} \right)^2 \nonumber
 \end{align}
-
-where $\varphi$ is the angle between adjacent edges, $E_\varphi$ is the energy associated with changes in the angle, $k_\varphi$ is the corresponding stiffness constant, and the $p_i$s are the Cartesian coordinates of the mass particles.
+where $\varphi$ is the angle between adjacent edges,% $E_\varphi$ is the energy associated with changes in the angle, $k_\varphi$ is the corresponding stiffness constant, 
+and the $p_i$s are the Cartesian coordinates of the mass particles.
 
@@
 \end{align}
+where $k_\varphi$ is the corresponding stiffness constant.
 

### 2017-01-25 twostagedo-rev-final.tex
Constraints... 2

@@
 \end{align}
-where $\varphi$ is the angle between adjacent edges,% $E_\varphi$ is the energy associated with changes in the angle, $k_\varphi$ is the corresponding stiffness constant, 
-and the $p_i$s are the Cartesian coordinates of the mass particles.
+where $\varphi$ is the angle between adjacent edges, and $\varphi_0$ is the angle at rest.
+% $E_\varphi$ is the energy associated with changes in the angle, $k_\varphi$ is the corresponding stiffness constant,
+%and the $p_i$s are the Cartesian coordinates of the mass particles.
 

### 2017-01-25 AnswersToComments-finalreview.tex
Added cover letter for final review.

--- 2016-12-26 AnswersToComments-secondreview.tex
@@
 \usepackage[spanish, mexico]{babel}
+\usepackage{calligra}
 
@@
 
-We would like to thank you and the three reviewers for timely and informative reviews, and for the comments, which have helped us to substantially improve our paper.
-In this revised version we have now included information on an implementation of a comparison model: a growing neural gas network.
-We show that it produces significantly poorer results on the prediction problem, and that it is thus unsuitable to tackle the classification problem we defined.
-In addition, we have carefully addressed all the reviewers’ concerns about the text.
+We would like to thank you and all reviewers for timely and informative reviews, and for the comments, which have helped us to substantially improve our paper.
+In this final revised version we address the comments from the two reviewers who still made observations.
 We detail below how we have addressed each reviewer point one by one.
 
-We hope that you find the resulting paper much improved.
+We look forward to hearing from you.
 
@@
 
-We would like to thank all the reviewers most sincerely for their effort in reading and commenting on our paper.
-The remarks helped us understand where we had gone wrong, and we have tried our best to incorporate all your requests and suggestions.
-Thank you.
+We would like to thank all the reviewers most sincerely for their effort in reading and commenting on our paper.  We address now their final remarks.
 
-\textbf{Reviewer 1}
+\textbf{Reviewer: 1}
 
-\emph{In this revision, the authors solved the initial problems on article structure, citations, figures, and contents.}
+{\calligra\large The authors have successfully modified the work according to the reviewers comments.
+In my opinion the work is ready to be published under one condition: equation 11 should be replaced with equation 23 from the appendix.}
 
-We thank the reviewer for their confirmation that we have satisfied all their comments, and would like to thank them again for their remarks, which significantly improved our paper.
+Thanks to this observation we noticed an inconsistency between the text and the equations, which suggested we should include equation 23 were we had 11.
+However, since all lengthy equations were moved to the appendix, instead of bringing equation 23 back, we arranged the text in section VI.A.
+so that only equations for the constraints would be shown here, and all derived forces could be found in the appendix.
+We hope this will make the presentation clearer.
 
-\textbf{Reviewer 2}
+\textbf{Reviewer: 2}
 
-\emph{Page 1. Last Paragraph. There is an enumeration of novel contributions, you mentioned the First, and the Third, to avoid any confusion I would recommend to remark which is the second.}
+{\calligra\large In this revision, the authors solved the major problems in previous manuscript, but there are several small problems:}
 
-We apologise, this was a typo. We have listed two main technical contributions.
+{\calligra\large 1, Equation (5) shows $F=m\frac{\partial p}{\partial t}$, but it seems that $F=m\frac{\partial^2 p}{\partial t^2}$ is more reasonable, since force equals mass multiplying acceleration.
+This problem remains in Equation (6) and the paragraph between Equation (5) and Equation (6).}
 
-\emph{Page 1-2. It is possible to infer the meaning of FEM, but in order to avoid any possible confusion you might indicate its meaning explicitly when the concept is first used.}
+Sorry you are right, we mixed notation which was not used in the article where the momentum of the particle is defined as $p=mv$.
+All the affected equations were modified to avoid this and keep $p$ as the position of the particle.
 
-We have fixed this.
+{\calligra\large 2, Equation (13) and Equation (14) are confusing. Is it a symbol conflict? Why does the derivative of $\vec{p}$ equal $\vec{p}$?}
 
-\emph{I find Fig 2, 3, and 4 difficult to read, however I don’t consider they should be necessarily changed.}
+It is a mathematical trick to force over damped systems of springs, which do not oscillate.
+The text in this paragraph was modified to make this more explicit.
+Also, we added a quick explanation about the differential equation of the form $f(x) = f'(x)$.
 
-We agree this was difficult to interpret. To aid the reader we have added some text descriptions for the most important elements to make the diagram more readable.
-
-\emph{At the end of subsection VII. B, when the authors refer to “all elastic and plastic terms” might be illustrative to add references to the corresponding equations.}
-
-We have added a footnote with a reference to the section where all terms are summarised.
-
-\emph{Within the Model Overview, it would be interesting if the authors would briefly highlight the advantage of having generative models for those who are not vastly familiarized with the topic.}
-
-Thank you for this suggestion.
-We have added a short explanation within the section Model Overview.
-This explains that generative models are appealing because they allow us to solve both prediction and classification problems, whereas discriminative models are typically only suitable for classification.
-
-\emph{When interpreting results what are the consequences when triangles in the border of the mesh overlap each other? Is there any consequence that might be significant to the result interpretation?}
-
-We have added some comments to clarify these points in the subsection “Additional Geometric Constraints”.
-
-\emph{In section V, during the training the segmentation assumes that the color of the object is known.
-Therefore, when you say that the algorithm could learn and classify previously unknown materials, at least the machine must partially know the material.}
-
-In fact, no assumption about the colour of the object is made.
-The object is simply marked with a bounding box in the first frame.
-This is used to determine the dominant hue.
-This hue is then used to distinguish edge pixels bounding the object from those gathered elsewhere in the scene.
-We have added this to the explanation of the first step of the algorithm in Section V.A.
-
-Also the authors mention that the simplified active contours approach is not the most accurate, What are the consequences of this approximation?
-
-The errors of the approximation are small, and are relatively small in comparison with the errors introduced by prediction with the mass-spring model.
-We added a note to this effect in section V.A  on page 8.
-
-Regarding to the chosen integration time step, does the results vary considerably when 0.01 is considered?
-
-We used a time step of 0.1.
-We have previously experimented with various time step parameters, and found that a time step of 0.01 didn’t improve the quality of prediction, but increased computation time.
-We have added a comment on this point to section V.B.
-
-\emph{In section VII.E, Table III, is there any special reason to have the same result for the training and test1 sets  in the case of the sponge?
-It would be better if the authors offer deeper insights regarding the interpretation of the figures shown in Table III.}
-
-Thanks for raising this point, we needed to explain this more clearly.  To enable this Table III has been shifted to section VIII, where the discussion has been expanded to cover these issues.
-
-\textbf{Reviewer 3}
-
-\emph{Reading the abstract and the introduction, I was expecting a framework in which the robot creates a representation of the environment (in this case the two objects) while it is interacting with it (looking at and pushing the object).}
-
-We are sorry that this set false expectations.
-We have altered the wording in several places to make the reader aware that we are not tackling this problem on-line, but off-line.
-We have also read and noted the references you mention in the introduction.
-
-\emph{While the authors recognize the issue of the on-line learning in the introduction (page 1, col 2, line 35), they do not address that in the paper.}
-
-From your comments, we believe that mentioning on-line learning, while important, was misleading in terms of our paper, so we have withdrawn mention of it.
-
-\emph{Also they refer the problem of getting access to partial views of the objects but it seems that in this paper the object is viewed from a convenient point of view.}
-
-We have added text to make clear the restriction in our study of using a single view chosen by the user.
-
-Finally, the word “learning from data” suggests to me that the model acquires a representation of the environment based on some kind on non-parametric model, such as neural network.
-However, the models seem carefully designed to represent the object deformation and the stress-gain diagram.
-Even if the difference between learning and calibration is subtle, in my opinion, the use of the word “learning” creates an expectation that is not fulfilled in the rest of the paper.
-
-We have changed the wording in the abstract to note that we calibrate the models from data.
-We hope this makes it clearer where on the spectrum of parameter tuning to full autonomous learning we lie.
-We now emphasise that we perform off-line learning, we sometimes refer to it as calibration, and we mention that we don’t address the issues of on-line learning.
-We hope this changes the tone of the introduction enough so as not to mislead.
-
-\emph{What happen if we add a new material? Do we need to add a new feature space?}
-
-If the object is plastic or elastic then we don’t need to add a new feature space, but we would be wise to calibrate a new model. We have added a comment to this effect in the conclusion.
-
-\emph{I would add a figure like Fig.
-9a at this point.
-In this way the reader can have an idea of the setup, how vision acquire information and how the mass-spring model depends on the visual information itself.
-For example, it might be interesting to combine Fig 9a and Fig 2 in single sketch of the model.}
-
-Thank you for this suggestion. We have prepared a new version of Figure 2, as you suggest. This incorporates elements from the old Figure 2 and Figure 9.
-
-\emph{A consideration about the organisation of the following sections.
-Given that the framework consists of two models, force prediction and shape prediction, I would expect section III.1 to be Force prediction and III.2 to the shape prediction (or vice-versa).
-However, the organization of Sections IV, V, VI, in my opinion is misleading.}
-
-We have reorganised the paper as you request.
-
-\emph{Finally, I would start describing the algorithm from the visual system. It seems that the visual system has a very important role in this paper but it is not highlighted.}
-
-We have slightly expanded and moved the section on the visual system as part of the paper reorganization. We also spend a little more space on vision in the model overview section.
-
-\emph{I would like to know how the particles are placed inside the object.
-What happens if we change the size of the object?
-Do we need to change the number of particles?
-What happen if the object is not rectangular?}
-
-We describe the way that the particles are placed inside the object in detail in Section VII.A on p.10.
-We specify a total number of cells in the mass-spring model, and then the particles (which are the nodes of the mass-spring model) are then spaced evenly in a grid.
-If we change the size of the object the method will allocate particles more widely spaced for a given grid size.
-No version of the algorithm has been specified for non-rectangular, we suggest that a Voronoi triangulation would be the way to proceed.
-We mention this in the conclusion.
-
-\emph{Eq.
-19.
-I really do not understand how the force could be proportional to the velocity.
-This point should be clarified.
-My understanding is that the authors assumed the acceleration (a) to be constant during the interval h, so that h*F = h*a*m => F = v*m/h.
-Given that h is a constant, the force (F), the velocity (v) or the mass (m) are recovered with a scale factor.
-Is that correct?
-If so, what is the difference between this approach and the quasi-static approach proposed by Frank.
-Please, explain.}
-
-The essential difference is that Frank models the system as being quasi-static, i.e.
-that acceleration is not only constant, but is zero.
-In contrast we model the acceleration as non-zero and non-constant.
-One point of possible confusion is that we are using a numerical method for approximate solution of the differential equations that model each mass particle.
-This numerical method simply assumes that between two successive points in time that acceleration is constant.
-Over multiple time steps, however, the acceleration does change.
-
-Regarding the physics, the corresponding math would be F = c m v  =>  a = v  =>  (a = dv/dt) = v which can only be solved with an exponential.
-Thus the constraint a=v forces us to solve with an exponential, and has the benefit of preventing oscillations of the springs in the model.
-This explanation has been added to the “Integration scheme” section.
-
-\emph{This section is quite long and I think it could be simplified notably.
-The authors could simply state the energy $E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2]$, where N is the number of constraints.
-There are several constraints to preserve the length, the area, etc.
-and the force is given as negative gradient of the potential energy with respect to the position of the particle.
-Details about all the equation could be reported in an appendix to make the paper more readable.
-About the equation, personally I don’t like the symbol composed of multiple letters such as Area, forcemag, forcedir, etc.}
-
-We have shortened the section as requested.
-
-\emph{Do the authors use just one video for the training and two video for the testing?
-It seems that the dataset is quite small to have a systematic test of the algorithm.
-Even if there are only two objects, the authors should test different motions of the finger (i.e.
-pushing velocity, penetration, …)}
-
-We accept that the test data set is rather small.
-We believe however that it is sufficient to show through a comparison with the GNGN method that our approach is considerably better than the state of the art when it comes to generalising deformation predictions with respect to push position.
-Also, we should note that the data needed for successful training is also very small: just one movie of each object.
-We will gather additional data, but our current plan is to use this for future work.
-
-\emph{I don’t see the results of the classification.
-I would expect a confusion matrix to show how many times the sponge was classified as sponge and as plasticine and vice-versa.
-Table III seems to show the results about the classification of the pixels.}
-
-Table III was moved to the discussion section were this point was clarified in the caption.
-The results show the classification not of the pixel, but of the frame.
-The classification of a whole movie is described in the text.
-
-\emph{Moreover, I would like to see a comparison against a baseline approach.
-They could some of the models proposed in the literature or some simple models.
-For example, the authors could assume that the material “disappears” when the finger penetrates the object and report the F-Score of this method.
-Alternatively, they could use a simple model with only one spring between the finger and the base of the object.}
-
-We have implemented a Growing Neural Gas Network (GNGN) previously published by Cretu et al.
-We have added additional results in a comparison of this method with our own.
-This is now described in section VII.D.
-The very poor generalization performance of the GNGN on the prediction problem is shown in a new Fig 17.
-
-\emph{Page 12.
-Col 2.
-Line 3.
-Please, can you clarify the hand tuning?
-It seems that the genetic algorithm is not the best way to find the parameters?
-Have you tried to use other method like gradient descent?
-Is it possible to use the model to mass-spring model to predict the force exerted on the finger?
-In that case, the authors could use this prediction error as fitness function and use the gradient descent to find the parameters.}
-
-The hand tuned solutions we tried were obtained by slightly modifying some of the parameters of the solutions proposed by the algorithm.
-However, the f-score of the solution chosen by the search algorithm is better than the f-score of the solution that was hand tuned, that is why the algorithm does not select it.
-This explanation was added to section IX.A.
-
-Gradient descent was not used because the space of parameters for a mass-spring model can offer good solutions in very separated regions.
-Therefore, gradient descent would not make a proper exploration of the space and would get stuck on local minima quite quickly.
-
-It could be possible to estimate the reaction force by adding the total forces on the particles in touch with the finger.
-At some point we did check if they were similar to the measured forces and found out that for some selected sets they are similar, but we didn’t try to use this criterion as evaluation function.
-The suggestion was added in the discussion section (IX.A).
-
-Other comments:
-
-\emph{Fig. 1 is not referenced in the paper.}
-
-We have decided that this figure added little, hence it was not referenced naturally anywhere in the text.
-We have thus decided to remove it to save space, since other requested changes have added considerable length.
-
-\emph{Page 2, Col 1, Line 17. Section VI is not mentioned.}
-
-We have fixed this.
-
-\emph{Page 2. Col 2, Line 5: FEM is not defined. Finite-element models?}
-
-We have defined this.
-
-\emph{Fig 2. Instead of use symbols, I would write Vision, strain, force, etc.}
-
-Thank you, this chimes with the comments of Reviewer 2, we have fixed this in Figures 1, 2 and 3.
-
-\emph{Page 3, Col 1, line 57. Cretu et al. -> cite [10]?}
-
-We have added this reference.
-
-\emph{Page 8. Col 2, Line 27. Linear snake = linear spline?}
-
-The term “linear spline” is associated with the geometric curve, while “linear snake” to the tracking algorithm that makes use of the linear spline.
-A clarification was added in section IV.A and the use of both terms was reviewed to make sure it was consistent with their definition.
-Thank you for noticing.
-
-\emph{Algorithm 5.
-I would specify that the F-Score is the fitness function of the algorithm.
-Moreover, it would be nice to know the values used for the parameters of the algorithm (perhaps using a table).}
-
-This information has been added to section VIII. C., the values for the parameters are on Table III.
-
-\emph{Page 10. Col 1. Line 45. Please specify that there are two objects, a sponge and plasticine, with a rectangular shape and so on.}
-
-We have added this information.
 \end{document}

### 2017-01-25 twostagedo-rev-final.tex
Added cover letter for final review.

@@
 %\fbox{
-\includegraphics[width=3.5in]{figures/learning.png}%}
+\includegraphics[width=3.5in]{figures/learning}%}
 \caption{The visual tracking and force measurement system subserves the offline learning system.
@@
 \centering
-\includegraphics[width=3.5in]{sponge__result.jpg}
+\includegraphics[width=3.5in]{sponge__result}
 \caption{Neural network approach fits quite well the training data set, but can not generalise to new pushing actions.
@@
 \centering
-\includegraphics[width=3.5in]{plasticine__result.jpg}
+\includegraphics[width=3.5in]{plasticine__result}
 \caption{GNG segmentation didn't work with a shadowed material (video of plasticine).
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.2in,clip,keepaspectratio]{arrio}}]{Veronica E. Arriola-Rios}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.2in,clip,keepaspectratio]{arrio.jpeg}}]{Veronica E. Arriola-Rios}
   has recently obtained her PhD in Computer Science from the University of Birmingham.
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.2in,clip,keepaspectratio]{wyatt}}]{Jeremy L. Wyatt} is Professor of Robotics
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.2in,clip,keepaspectratio]{wyatt.jpeg}}]{Jeremy L. Wyatt} is Professor of Robotics
   and Artificial Intelligence at the University of Birmingham. He

### 2017-01-25 figures/original/convert_ieee.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+for f in *.png; do convert -units PixelsPerInch "$f" -resample 300 "${f%%.*}.tiff"; done
+for f in *.jpg; do convert -units PixelsPerInch "$f" -resample 300 "${f%%.*}.tiff"; done
+

### 2017-02-01 AnswersToComments-finalreview.tex
changes to letter. Still need to check against paper.

@@
 
-We would like to thank you and all reviewers for timely and informative reviews, and for the comments, which have helped us to substantially improve our paper.
-In this final revised version we address the comments from the two reviewers who still made observations.
+We would like to thank you and all reviewers for timely and informative reviews, and for the minor changes requested.
+In this final revised version we address these final comments.
 We detail below how we have addressed each reviewer point one by one.
 
-We look forward to hearing from you.
+We hope that the paper is now in a satisfactory state for publication, and look forward to hearing from you.
 
@@
 
-We would like to thank all the reviewers most sincerely for their effort in reading and commenting on our paper.  We address now their final remarks.
+We would like to thank all the reviewers most sincerely for their effort in reading and commenting on our paper.  We now address their final remarks.
 
@@
 
-{\calligra\large The authors have successfully modified the work according to the reviewers comments.
+{\em The authors have successfully modified the work according to the reviewers comments.
 In my opinion the work is ready to be published under one condition: equation 11 should be replaced with equation 23 from the appendix.}
 
-Thanks to this observation we noticed an inconsistency between the text and the equations, which suggested we should include equation 23 were we had 11.
-However, since all lengthy equations were moved to the appendix, instead of bringing equation 23 back, we arranged the text in section VI.A.
-so that only equations for the constraints would be shown here, and all derived forces could be found in the appendix.
-We hope this will make the presentation clearer.
+Thanks to this observation we noticed an inconsistency between the text and the equations.
+We see that this inconsistency made it seem to the reader that equation 23 should appear where we used equation 11.
+However, rather than fixing the inconsistency by swapping the equations we have chosen to address it by changing the text in the paragraph above equation 11.
+We believe that only equations for the constraints should be shown in this section, as it is from the constraint equations that all elese arises.
+So we show a constraint equation for each of the length, area and angle.
+To this end we also change Equation 13 (from concerning energy to the constraint that underpins the energy).
+We show how both energy and all derived forces are obtained from these constraint equations in the appendix.
+We hope this makes the presentation clearer in the manner intended by the reviewer.
+Thanks for spotting this.
 
@@
 
-{\calligra\large In this revision, the authors solved the major problems in previous manuscript, but there are several small problems:}
+{\em In this revision, the authors solved the major problems in previous manuscript, but there are several small problems:}
 
-{\calligra\large 1, Equation (5) shows $F=m\frac{\partial p}{\partial t}$, but it seems that $F=m\frac{\partial^2 p}{\partial t^2}$ is more reasonable, since force equals mass multiplying acceleration.
+{\em 1, Equation (5) shows $F=m\frac{\partial p}{\partial t}$, but it seems that $F=m\frac{\partial^2 p}{\partial t^2}$ is more reasonable, since force equals mass multiplying acceleration.
 This problem remains in Equation (6) and the paragraph between Equation (5) and Equation (6).}
@@
 
-{\calligra\large 2, Equation (13) and Equation (14) are confusing. Is it a symbol conflict? Why does the derivative of $\vec{p}$ equal $\vec{p}$?}
+{\em 2, Equation (13) and Equation (14) are confusing. Is it a symbol conflict? Why does the derivative of $\vec{p}$ equal $\vec{p}$?}
 
